### 2015-02-15 Backup/Old versions/OldManuscript/MS_LDD.tex
Initial commit of directory structure.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% MS LDD - Pedro - 19 July 2006
% Revised for 2008
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Headers
\documentclass[a4paper,12pt]{article}
\usepackage{amsmath,amssymb}
%\usepackage[utf8]{inputenc}
%\usepackage[applemac]{inputenc}
\usepackage{geometry}
\usepackage{lscape}
\usepackage{setspace}
\usepackage{framed}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage{longtable}
\usepackage{rotating}                                                          
\newcommand{\tab}{\hspace{5mm}}
\usepackage{tabularx} 
\usepackage[margin=10pt,font=small,labelfont=bf]{caption}
\usepackage[left,pagewise]{lineno}
\usepackage{caption}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy}     % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{{\tiny Jordano - LDD events}}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title page
\begin{document}
\title{Manuscript Draft\\
\vspace{2cm}
What is long-distance dispersal? and a taxonomy of dispersal events}

\author{Pedro Jordano$^{\dag}$}

\date{Sevilla, \today}
\maketitle

\begin{spacing}{1.0}
$^{\dag}$ {\small Integrative Ecology Group, Estaci\'on Biol\'ogica de 
Do\~nana, CSIC, Pabell\'on del Per\'u, Avda. Mar\'ia Luisa, s/n, 
E-41013 Sevilla, Spain.}\\

{\small \textit{Corresponding author:} Pedro Jordano. Integrative Ecology Group, Estaci\'on Biol\'ogica de Do\~nana, CSIC, Pabell\'on del Per\'u, Avda. Mar\'ia Luisa, s/n, E-41013 Sevilla, Spain. Fax number: +34 95 4621125. Email address: jordano@ebd.csic.es}\\

\textbf{Key words}: dispersal kernel, fragmentation, frugivores, gene-flow, genetic neighborhood, long-distance dispersal, seed dispersal\\

{\small \textbf{Manuscript information: }** Words; ** Chars; ** Pages, * Figures; * Tables.}
\maketitle
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Abstract
\begin{linenumbers}
\section*{Abstract}

Long-distance dispersal (LDD) is a major component of the population dynamics, genetic structure, and biogeographic history of plant species. It determines the colonization ability of new habitats and the possibilities for fragmented populations to sustain a cohesive metapopulation by inmigration-emigration dynamics that rely on LDD events. Yet our current understanding of the extent, frequency, and consequences of LDD is very limited. On one hand, theoretical models fail to predict accurately the behavior of the tail of the dispersal functions, and thus fail to predict very basic properties of LDD. On the other hand, we still have very limited documentation of actual LDD events in natural populations and consider LDD as a sporadic, rarely far-reaching process marked with the stamp of natural history cursiosity. Here we propose an explicit definition of LDD and what constitutes a LDD event. We propose a combination of ecological and genetic evidences to define LDD events. LDD events should be characterized in terms of distance relative to the ecological limits of the population (in a metapopulation context) and to the limits of the genetic neighborhood.

\end{linenumbers}

\end{spacing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Introduction
\begin{flushleft}
\begin{spacing}{1.5}
\newpage 
\begin{linenumbers}
\section*{Introduction}

- General overview of LDD events.

- LDD events no so uncommon, especially for animal-mediated dispersal of pollen and seeds, but also for abiotic dispersal when singular events (strong wind uplift, tornadoes, water runoff) occur.

- Despite extensive evidence for the occurrence of LDD events and recents evidences of far-reaching consequences, no unambiguous definition of LDD still available.

- Adequate definition probably requires transdisciplinary concepts, from landscape ecology, metapopulation ecology and population genetics. Key ingredients to the definition would be landscape connectivity, genetic neighborhood size.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Materials and Methods
\begin{comment}
\section*{Material and Methods}
\subsection*{ Species and study site characteristics}

The study species, \textit{Prunus mahaleb} (L.) (Rosaceae), is a tree  producing fleshy-fruits ingested by frugivores that disperse  their seeds after regurgitating or defecating them.

\subsection*{*}

\subsubsection*{Sampling of dispersed seeds}
\end{comment}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Results  and Discussion

\section*{Results and Discussion}

\subsection*{Individual and population neighborhoods as reference marks}

- NOTE. Take evidence from Prunus for reliable estimates of percent pollen and seed coming from outside the population, compare estimates of neighbourhood size with e.g. work on Ficus in BCI \citep{E3.2543,Nason:1996,Stacy:1996vo} and Fraxinus \citep{Bacles:2006kq}.

- Population graph approach \citep{Dyer04}.

\subsection*{Evidences of neighborhood sizes and LDD events}

\subsection*{Challenges and promising avenues for research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}

I am indebted to Jos\'e A. Godoy, Manolo Carri\'on, Juan Luis Garc\'ia-Casta\~no, Jes\'us Rodr\'iguez, Cristina Garc\'ia and, especially, Juan Miguel Arroyo for generous help with field and laboratory work and making possible this study. We appreciate the help and advice of Cristina Garc\'ia, Jos\'e A. Godoy and Jordi Bascompte during the final stages of the manuscript. The study was supported by the Spanish Ministerio de Ciencia y Tecnolog\'ia (REN2003-0273 and CGL2006-00373 to PJ). The Agencia de Medio Ambiente, Junta de Andaluc\'ia, provided generous facilities that made possible this study in the Sierra de Cazorla and authorized my work there.

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% References
%\section{References}
\bibliographystyle{ecology_letters} % Choose EcolLett style for bibliography
%\bibliography{/Users/pedro/Documents/Databases/Bib_database}
\bibliography{MS_LDD}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{linenumbers}

\end{spacing}

\end{flushleft}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%:----MS NOTES
\begin{comment}

Breeding unit size. Donor number is related to the size of the breeding unit (N individuals) as d= Np, where d is the number of different pollen parent genotypes represented in a maternal treeÕs fruit crop and p is the expected proportion of trees in staminate phase at any given time. Assuming,conservatively, that individual trees reproduce asynchronously twice per year 6 and that staminate and pistillate flowering phases each last seven days 20, p is calculated to be 0.0712. Although these estimates could be effected by mosaicism 21 (the fusion of genetically distinct individuals), its frequency in the species examined here is low.

Breeding unit area and radii. These parameters were calculated from paternity-analysis-based estimates of breeding unit size (Nödö; Table2) and the censused densities of adult, reproductively mature trees over 15 km2 BCI. Based on the long-term censuses of C. Handley and E. Kalko, 6, 108 and 20 adults of F. dugandii, F. obtusifolia and F. popenoei, respectively, are known to occur on BCI. Because these species exhibit little spatial aggregation over the area censused, these densities are assumed to be representative of forested areas surrounding BCI, a conservative assumption given that approximately one-third of the area within 10 km of BCI is occupied by Lake Gatun where figs are absent. Breeding unit radii estimate the distances pollen-bearing, female fig wasps routinely disperse in search of receptive host trees. Although actual breeding populations of figs may deviate substantially from the assumed circular distribution, alternative structures (elliptical,for example)increase estimated wasp dispersal distances.

- LevinÕs paternity pool concept.  Levin, D. A. The paternity pools of plants. Am. Nat. 132, 309Ð317 (1988).

[Burnham, K.P. & Overton, W.S. Robust estimation of population size when capture probabilities vary among animals. Ecology 60, 927Ð936 (1979).]

\end{comment}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TABLES
\newpage
%-------------------------------------------------------------------- Table 1
\begin{table}
%\captionsetup{width=15cm}%.75\textwidth}
\caption{Types of dispersal as a function of population area limits and genetic neighborhood limits. LDD events can take place in a combination of circumstances, not necessarily involving dispersal outside the population.}
\vspace{0.5cm}

\begin{tabular}{llcccc}
\addlinespace
\toprule
      & &	\multicolumn{2}{c}{Population limit }\\
      &	&	Within	&	Outside \\
\toprule

Genetic	& Within	& Local, short-distance 		&Within neighborhood, long-\\
neighborhood	&	&dispersal	&	distance dispersal\\
limit	&	&	$SDD_{loc}$	&	$LDD_{neigh}$\\
	&	Outside &	Local, long-distance	&	Strict sense long-distance\\
	&	&	dispersal	&	dispersal\\
	&	&	$LDD_{loc}$	&	$LDD_{SS}$\\

\toprule

\end{tabular}
\end{table}
%-------------------------------------------------------------------- Table 1

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURES

\newpage
%\section*{Figures}
%------------------------------------------------------------------- Figure 1
\begin{figure}[htbp]
\centerline{\includegraphics[height=20cm]{Fig1.pdf}}
\caption{Types of dispersal events related to population limits and genetic neigborhood size. The grey area depicts the breeding unit area. Depending on its size it may overlap with neighbour populations (bottom) or be much smaller than the population limits (top). Consequently, strcit sense LDD events $LDD_{SS}$ should be considered those beyond both the ecological and breeding unit area limits (top). LDD events within a local population ($LDD_{loc}$) can occur when propagules disperse beyond the local limits of the reeding unit area. Propagules exiting the population ($LDD_{neigh}$ but moving within the breeding unit area should not be considered strict sense LDD events. Short distance dispersal events are those occurring within the limits of both the population and breeding unit area $SDD_{loc}$.}
\end{figure}
%----------------------------------------------------------------------------

\end{document}

### 2015-02-15 Backup/Old versions/OldManuscript/MS_LongDist.tex
Initial commit of directory structure.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% MS LAYOUT - Pedro - 20 August 2005.
%% Revised version -   17 Dec    2006.
%% MS text moved to new file EcolLett. 22 May 2011.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[a4paper,12pt]{article}
\usepackage{amsmath,amssymb}
%\usepackage[utf8]{inputenc}
%\usepackage[applemac]{inputenc}
\usepackage{geometry}
\usepackage{lscape}
\usepackage{setspace}
\usepackage{framed}
\usepackage{verbatim}
\usepackage{graphicx}                         
\usepackage{epstopdf}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage{longtable}
\usepackage{rotating}                                                                                    
\newcommand{\tab}{\hspace{5mm}}
\usepackage{tabularx} 
\usepackage[margin=10pt,font=small,labelfont=bf]{caption}
\usepackage[left,pagewise]{lineno}
\usepackage{caption}
\usepackage{bibunits}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy}     % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{{\tiny Jordano \& Godoy - LDD events}}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}
\setlength{\parskip}{\baselineskip} % increase paragraph separation
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title page
\begin{document}
\begin{bibunit}[science]
\title{Manuscript Draft\\
\vspace{2cm}
Extending the seed shadow: Long-distance dispersal of tree seeds by frugivorous animals}

\author{Pedro Jordano$^{\dag}$ and Jos\'e A. Godoy$^{\dag}$}

\date{Sevilla, \today}
\maketitle

\begin{spacing}{1.0}
$^{\dag}$ {\small Integrative Ecology Group, Estaci\'on Biol\'ogica de 
Do\~nana, CSIC, Pabell\'on del Per\'u, Avda. Mar\'ia Luisa, s/n, 
E-41013 Sevilla, Spain.}\\

{\small \textit{Corresponding author:} Pedro Jordano. Integrative Ecology Group, Estaci\'on Biol\'ogica de Do\~nana, CSIC, Pabell\'on del Per\'u, Avda. Mar\'ia Luisa, s/n, E-41013 Sevilla, Spain. Fax number: +34 95 4621125. Email address: jordano@ebd.csic.es}\\

\textbf{Key words}: dispersal kernel, fragmentation, frugivores, gene-flow, genetic neighborhood, long-distance dispersal, seed dispersal\\

{\small \textbf{Manuscript information: }** Words; ** Chars; ** Pages, * Figures; * Tables.}
\maketitle
\end{spacing}
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Abstract
\begin{linenumbers}
\begin{spacing}{1.5}
\begin{abstract}
Documenting long-distance dispersal events by frugivorous animals has been a challenge in seed dispersal studies \citep{Nathan:2003qe,Nathan:2006qy}. The difficulty to track movement patterns of animals and seed processing in the gut \citep{Mack:1995,Holbrook:2002fk,Holbrook:2006uq,Westcott:2005} has precluded attempts to directly estimate dispersal distances and made necessary resorting to estimates based on inverse modeling \citep{Clark:1999sv,Katul:2005yq,Nathan:2006}. Here we use multilocus microsatellite genotypes obtained from seed endocarps \citep{Godoy:2001} in combination with bayesian inference methods \citep{Wilson:2003gf} to assign the source population of dispersed seeds in a St Lucie's cherry population. We document long-distance dispersal by animals by showing that inter-population dispersal distances within a single reproductive episode can be as long as 17 km and represent up to 20 \% of the seeds. However, seed-mediated gene flow among these fragmented populations remains highly restricted, with most events of long-distance dispersal occurring among nearby populations. These are the first reliable long-distance estimates of seed dispersal among populations and are relevant to our understanding of connectedness among forest fragments and for predicting responses to habitat loss.
\end{abstract}
\end{spacing}
\end{linenumbers}
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Main text
\begin{linenumbers}
\begin{spacing}{2}
\label{sect:intro}
Dispersal events of seeds and pollen assisted by animals are central ecological processes in many plant species, contributing to genetic structure, connectivity of metapopulations, and gene flow. While recent research on gene-flow patterns has emphasized pollen-mediated dispersal, especially for anemophilous species \citep{Ennos:2001ii,Sork:1999}, seed dispersal can be a major influence in plant population genetic structure because of its lasting effects on recruitment \citep{Ouborg:1999}. When frugivorous animals mediate in seed dispersal, unfrequent long-distance dispersal events (LDD) can occur consistently, i.e., repeatedly in successive reproductive episodes. Yet assessing the source and the dispersal vector for dispersed seeds sampled directly in the field is a lasting challenge in seed dispersal studies \citep{Nathan:2006qy} despite the fact that these are key parameters to understand animal-mediated dispersal. Both the results from inverse modeling \citep{Clark:1999sv} of the seed shadows (i.e., the spatial, landscape pattern of seed distribution following seed dispersal, \citep{Sork:1999}) and the scarce evidence from direct field observation of frugivore movements contributed to the widely admitted tenet of limited dispersal by frugivorous animals when compared to dispersal by wind \citep{Oddou-Muratorio:2001}. The results might appear contradictory with indirect inferences of long-distance seed dispersal events, generally mediated by  large frugivorous animals \citep{Fragoso:1997,Holbrook:2002fk,Holbrook:2006uq,Westcott:2000}, involving some times successful long-distance dispersal and dispersal to oceanic islands. The frequency distributions of animal-mediated seed dispersal distances are characterized by fat tails that render rigorous modeling deceptively limited \citep{Jordano:2007,Jordano:2007a,Nathan:2006qy}. Despite our ability and recent advances \citep{Godoy:2001,Garcia:2006,Jordano:2007,Hardesty:2006lr,Jones:2005fk_E31.7321} to document such sporadic events of successful long-distance dispersal, a serious limitation for further developments stems in our inability to estimate their frequency. Thus, predicting the frequency of dispersal \texttt{>}100 m, and specially the detection of among-population dispersal, is seriously  constrained by current methods. The inferences from seed trap  data alone \citep{Clark:1999sv} or seed marking studies \citep{Levey_etal_E31_7300}, rely on spatially limited sampling designs, and are limited by the difficulties in fitting the dipersal functions for the tail of the distribution \citep{Nathan:2003qe,Jordano:2007a}. Therefore, a serious methodological and conceptual challenge in seed dispersal studies has been the robust characterization of the frequency and extent of long-distance dispersal events.

Assessing who moves seeds where and which are the sources of these seeds (e.g., the population and maternal tree sources) is a key question in dispersal studies. When this information is available, reliable reconstructions of the seed dispersal kernels can be achieved (i.e., the total dispersal kernel) \citep{Nathan:2003qe,Jordano:2007a}. The connectedness among population fragments can be determined by combining these type of direct estimates of seed-mediated gene flow and spatially-explicit analysis of landscape connectivity \citep{Urban:2001}.

By using hypervariable microsatellite markers with DNA amplified from the maternally-inherited endocarp seed tissue \citep{Godoy:2001} we have been able to ascribe the tree of origin for animal dispersed seeds within a population of \textit{Prunus mahaleb} (Rosaceae), an endangered species in SE Spanish mountains. Dispersal of this tree is mediated by a diverse coterie of frugivores, among which small-sized passerines (\textit{Sylvia} warblers, \textit{Turdus} thrush species, and redstarts \textit{Phoenicurus} spp.) predominate, together with carnivorous mammals \citep{Jordano:2000EcolMon}. Recently we showed that seed dispersal from other populations into a focal population can be substantial, and contributed differentially by a distinct subset of frugivore species, namely the largest sized ones \citep{Jordano:2007}.
% Need to re-write this par. with new updated data on samples...
We sampled seeds in seed traps laid out in a network of 20 sampling points including the major types of microhabitats in the forest (\textit{20}), throughout a ca. 21 ha plot including a large \textit{P. mahaleb} population (NCH, Fig. 1A) with 196 adult trees. Traps passively sampled seeds defecated or regurgitated by frugivores. All the adult trees in the population were genotyped for 11 microsatellite (simple sequence repeat, SSR) loci from leaf-extracted DNA. In addition, we sampled 192 trees from other 8 populations in the study area (Fig. 1A), totaling 388 trees genotyped. We assigned the maternal tree of origin for each of 95 seeds sampled in 1996 based on a direct comparison (identity check) of the multilocus genotype of each seed endocarp with the genotype of each of the adult trees, obtained from leaf tissue (\textit{10}). This method resulted in unambiguous direct estimates of the within-population seed dispersal distances for \textit{N}= 85 (89.47\%) of the seeds sampled in NCH that had the maternal source tree in this population. An ambiguous assignment of 2 seeds was due to failed amplification of 2 loci for these endocarps and was resolved by assigning the source to the nearest tree (see Suppl. Material).

Most of the dispersal events involving the within-population movement of seeds were within 25 m from the source maternal trees (Fig. 1B, \textit{10}). The remaining \textit{N}= 10 seeds (10.53 \%) are attributable to long-distance dispersal from other populations. We used assignment tests (\textit{21}) applied to the sample of 10 seeds unassigned to NCH trees to estimate the expected frequency of the seed endocarp multilocus genotype in each of 8 neighboring populations over an area of 30 x 25 km (Fig. 1A). We assigned each seed to the population with the highest log-likelihood of occurrence (Fig. 1A). Most seeds (6 out 10) came from populations CL and PCL, located at 3.2 and 1.2 km, respectively, from NCH (Fig. 1A). The remaining 4 seeds came from two relatively large populations (TC, with 100-150 trees, and NN having \texttt{} 1000 trees) located in a different watershed (Fig. 1B). Accounting for the area covered by \textit{P. mahaleb} trees in each population, we estimated the distances from the center of each population to be * km for TC and 6.4 km for NN. In contrast, we did not detect seed input from other nearby populations (e.g., CT, RH, Fig. 1B).

We are not aware of previous direct estimates of long-distance seed movement mediated by frugivorous animals and incorporating among-population dispersal. Our data extend the seed shadow of \textit{P. mahaleb} in a single reproductive episode up to a minimum of 6.4 km from the focal population and demonstrate that long-distance dispersal events are not infrequent yet very restricted to close populations. If frugivores track the temporal availability of fruits at different elevations (\textit{22, 23}), this might cause restricted possibilities for seed exchange among fragmented populations, even at close distance, that occupy different elevations (\textit{10}) and explains limited dispersal despite thorough removal of fruits by frugivores. Our data also suggest that the interchange of seeds is not symmetric among each pair of populations. The probability of NCH seeds being dispersed to CAL or PCAL should be much higher than the inverse, primarily because of the large difference in population size. Thus, we had evidence that seeds from the large NN and TC populations arrived in NCH in this particular reproductive event, despite being located at long distance and in a different watershed.

Animal frugivores are efficient in removing \textit{P. mahaleb} seeds and should contribute significantly to the connectedness of the fragmented populations via seed-mediated gene flow. The within- and among-population seed dispersal distances we document here are considerably larger than those reported previously for animal-dispersed species (\textit{5, 18, 24, 25}) and for breeding unit areas derived from pollen dispersal studies (\textit{14}). Previous analyses of the pollen and seed components of gene flow (\textit{12, 26-28}) have emphasized the former, probably due to the fact that most species examined had not endozoochorous dispersal and to the difficulty of reliably assessing seed dispersal events unambiguously, especially the long-distance ones. Using hypervariable molecular markers to jointly track seed dispersal patterns and to analyze open-pollinated progenies will help to clarify the relative contributions of seed and pollen movement to total gene flow. This will facilitate realistic modeling of dispersal functions incorporating the ubiquity and relative temporal frequency of long-distance dispersal events. In addition, our results emphasize the importance of considering the cascading, pervasive, effects of frugivore extirpartion from natural communities, where they play an important role in keeping the connectedness among population fragments and thus have a pivotal role for population recovery following perturbations.

\end{spacing}
\end{linenumbers}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Acknowledgments
\begin{linenumbers}
\begin{spacing}{1.5}
\section*{Acknowledgments}
\label{sect:ackn} We are indebted to Manolo Carri\'{o}n, Juan Luis Garc\'{\i}a-Casta\~{n}o, Jes\'{u}s Rodr\'{\i}guez, Cristina Garc\'{\i}a and, especially, Juan Miguel Arroyo for generous help with field and laboratory work and making possible this study. We appreciate the help and advice of Albert Abbott, Graham King and Amy Iezzoni, who generously provided \textit{Prunus} microsatellites before publication. Discussions with Cristina Garc\'{\i}a, Juan Miguel Arroyo, Eugene W. Schupp, Janis Boettinger, Jordi Bascompte, R\'{e}my Petit, Arndt Hampe, Alfredo Valido, and Carlos Meli\'{a}n were very helpful during the final stages of the manuscript. The study was supported by the Spanish Ministerio de Ciencia y Tecnolog\'{\i}a (BOS2000-1366-C02-01, REN2003-0273, and CGL2006-00373 to PJ). The Agencia de Medio Ambiente, Junta de Andaluc\'{\i}a, provided generous facilities that made possible this study in the Sierra de Cazorla and authorized our work there.
\end{spacing}
%\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% References
%\section{References}
%%% Literature Cited for main text
%\bibliographystyle{science} % Choose Science style for bibliography
%\bibliography{MS_Longdist}
\putbib[bu1]
\end{linenumbers}
\end{bibunit}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Tables
%\newpage
%section*{Tables}
\label{sect:tabl}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Figures
\newpage
%%%% Figure LEGENDS

\textbf{Figure legends}\\

\textbf{Figure 1}. A. Map of the study area (located in the SE limit of Ja\'{e}n province, SE Spain- thick black line; thin lines, main streams and rivers) showing the 9 populations studied. Seeds sampled in the Nava de las Correhuelas focal population (NCH) and not assigned to the local trees growing there (N= 117 seeds, 1996 and 1997 seed crops) were tested for provenance from 8 populations in the region located within 1.5-20 km distance from the focal population. Numbers refer to the number of seeds assigned to each of the surrounding populations. B. The seed dispersal curve of \textit{Prunus mahaleb}, i.e., the frequency distribution of dispersal distances in a tree population (NCH, Nava de las Correhuelas) including the long-distance, among-population, dispersal estimates. The abcissa shows the actual distance values between the location of a dispersed seed (defecated or regurgitated by a frugivore) and its maternal tree (log-scale, actual distance + 1 m). At distance 0 m we represent seeds falling beneath the canopy of the mother tree. The inset shows the frequency distribution of dispersal distances within the population (i.e., not including seeds originating from other populations).\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURES
%------------------------------------------------------------------- Figure 1
\begin{sidewaysfigure} 
\centering
%\centerline{\includegraphics[height=11cm]{Fig1.pdf}} 
\includegraphics[height=15.5cm]{Fig1.pdf}
\end{sidewaysfigure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% ONLINE SUPPORT MATERIAL
\newpage
\begin{bibunit}[science]
\section*{Online Support Material}
\begin{spacing}{2.0}
\textit{Methods}
A total of 9 local \textit{Prunus mahaleb} populations were selected  for analysis, all them in Parque Natural de las Sierras de Cazorla,  Segura y Las Villas (Ja\'{e}n province, SE Spain). In this area \textit{P.  mahaleb} naturally occurs as isolated small (10 trees) to medium-size  (150 trees) distinct populations. The main study population was  located in Nava de las Correhuelas (NCH), at 1615 m elevation. Detailed  descriptions of the area and general methods can be found in  \citep{Jordano:1994,Jordano:1995,Jordano:2000EcolMon,Jordano:2002}. We genotyped 388 trees, including all the 196 adult  trees from the main study site, NCH, increasing our previous  sample for this population (\citep{Godoy:2001}). Details of the extraction  and amplification protocols can be found in \citep{Godoy:2001,Jordano:2002}.

Seeds dispersed by animals were sampled by setting 20 replicates  of 2 seed traps each at different randomly-chosen locations in NCH population: beneath \textit{P.  mahaleb} trees (5 trees), beneath high shrubs (10 locations),  and benetah pine trees (5 locations) either with or without juniper  understory \citep{Godoy:2001,Jordano:2002}.

To assign the source tree for each dispersed seed we carried  out an identity check by matching the multilocus genotype of  the endocarp at 9 loci with those of the adult trees; for the  adult trees we assessed 11 SSR loci \citep{Cipriani:1999,Sosinski:2000}. We used CERVUS (\textit{9}) to identify the mismatches and multiple-matches among  endocarps and putative source trees. All the seed endocarps except  2 (among those matching NCH tree genotypes) were assigned to  a single tree in NCH; ten endocarps were not assignable to any  tree in NCH. The two seeds with double matching have failed amplifications  for two loci and resulted in ambiguous matchig with two putative  maternal trees. We assigned the seeds to the tree nearest to  the sampling location, due to the fact that this procedure would  minimize the estimation errors of dispersal distances when these  are distributed with high skew, i.e., a missasignment of a rare long-distance event would seriously bias the estimate of the dispersal function.

To assing the population of origin for the seeds not matched  by any tree in NCH we used a maximum-likelihood test based on  the probability that each particular endocarp multilocus genotype  would be obtained in each of the 9 populations (excluding NCH).  The assignement proceeds as follows \citep{Wilson:2003gf,Beaumont:2004lr}: for each population  and locus, the relative frequencies of alleles are estimated  (\textit{p}$_{\mathit{il}}$\textit{, p}$_{\mathit{jl}}$\dots ); the expected probability of  a seed endocarp genotype is determined for each locus in that  population and these expected frequencies are muliplied across  loci. The endocarp genotype is assigned to the population with  highest log-likelihood of occurrence.\\

\end{spacing}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% References for Online Material
\putbib[bu2]
\end{bibunit}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2015-02-15 Backup/Old versions/OldManuscript/_region_.tex
Initial commit of directory structure.

--- 2015-02-15 Backup/Old versions/OldManuscript/MS_LDD.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\message{ !name(\string~/Documents/Working/MS_LDD/Manuscript/_region_.tex)}\message{ !name(MS_LDD.tex)}%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% MS LDD - Pedro - 19 July 2006
-% Revised for 2008
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{document}
-\title{Manuscript Draft\\
-\vspace{2cm}
-What is long-distance dispersal? and a taxonomy of dispersal events}
 
-\author{Pedro Jordano$^{\dag}$}
-
-\date{Sevilla, \today}
-\maketitle
-
-\begin{spacing}{1.0}
-$^{\dag}$ {\small Integrative Ecology Group, Estaci\'on Biol\'ogica de 
-Do\~nana, CSIC, Pabell\'on del Per\'u, Avda. Mar\'ia Luisa, s/n, 
-E-41013 Sevilla, Spain.}\\
-
-{\small \textit{Corresponding author:} Pedro Jordano.
-Integrative Ecology Group, Estaci\'on Biol\'ogica de Do\~nana, CSIC, Pabell\'on del Per\'u, Avda.
-Mar\'ia Luisa, s/n, E-41013 Sevilla, Spain.
-Fax number: +34 95 4621125.
-Email address: jordano@ebd.csic.es}\\
-
-\textbf{Key words}: dispersal kernel, fragmentation, frugivores, gene-flow, genetic neighborhood, long-distance dispersal, seed dispersal\\
-
-{\small \textbf{Manuscript information: }** Words; ** Chars; ** Pages, * Figures; * Tables.}
-\maketitle
-\newpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Abstract
-\begin{linenumbers}
-\section*{Abstract}
-
-Long-distance dispersal (LDD) is a major component of the population dynamics, genetic structure, and biogeographic history of plant species.
-It determines the colonization ability of new habitats and the possibilities for fragmented populations to sustain a cohesive metapopulation by inmigration-emigration dynamics that rely on LDD events.
-Yet our current understanding of the extent, frequency, and consequences of LDD is very limited.
-On one hand, theoretical models fail to predict accurately the behavior of the tail of the dispersal functions, and thus fail to predict very basic properties of LDD.
-On the other hand, we still have very limited documentation of actual LDD events in natural populations and consider LDD as a sporadic, rarely far-reaching process marked with the stamp of natural history cursiosity.
-Here we propose an explicit definition of LDD and what constitutes a LDD event.
-We propose a combination of ecological and genetic evidences to define LDD events.
-LDD events should be characterized in terms of distance relative to the ecological limits of the population (in a metapopulation context) and to the limits of the genetic neighborhood.
-
-\end{linenumbers}
-
-\end{spacing}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Introduction
-\begin{flushleft}
-\begin{spacing}{1.5}
-\newpage 
-\begin{linenumbers}
-\section*{Introduction}
-
-- General overview of LDD events.
-
-- LDD events no so uncommon, especially for animal-mediated dispersal of pollen and seeds, but also for abiotic dispersal when singular events (strong wind uplift, tornadoes, water runoff) occur.
-
-- Despite extensive evidence for the occurrence of LDD events and recents evidences of far-reaching consequences, no unambiguous definition of LDD still available.
-
-- Adequate definition probably requires transdisciplinary concepts, from landscape ecology, metapopulation ecology and population genetics.
-Key ingredients to the definition would be landscape connectivity, genetic neighborhood size.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Materials and Methods
-\begin{comment}
-\section*{Material and Methods}
-\subsection*{ Species and study site characteristics}
-
-The study species, \textit{Prunus mahaleb} (L.) (Rosaceae), is a tree  producing fleshy-fruits ingested by frugivores that disperse  their seeds after regurgitating or defecating them.
-
-\subsection*{*}
-
-\subsubsection*{Sampling of dispersed seeds}
-\end{comment}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Results  and Discussion
-
-\section*{Results and Discussion}
-
-\subsection*{Individual and population neighborhoods as reference marks}
-
-- NOTE.
-Take evidence from Prunus for reliable estimates of percent pollen and seed coming from outside the population, compare estimates of neighbourhood size with e.g.
-work on Ficus in BCI \citep{E3.2543,Nason:1996,Stacy:1996vo} and Fraxinus \citep{Bacles:2006kq}.
-
-- Population graph approach \citep{Dyer04}.
-
-\subsection*{Evidences of neighborhood sizes and LDD events}
-
-\subsection*{Challenges and promising avenues for research}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Acknowledgements}
-
-I am indebted to Jos\'e A.
-Godoy, Manolo Carri\'on, Juan Luis Garc\'ia-Casta\~no, Jes\'us Rodr\'iguez, Cristina Garc\'ia and, especially, Juan Miguel Arroyo for generous help with field and laboratory work and making possible this study.
-We appreciate the help and advice of Cristina Garc\'ia, Jos\'e A.
-Godoy and Jordi Bascompte during the final stages of the manuscript.
-The study was supported by the Spanish Ministerio de Ciencia y Tecnolog\'ia (REN2003-0273 and CGL2006-00373 to PJ).
-The Agencia de Medio Ambiente, Junta de Andaluc\'ia, provided generous facilities that made possible this study in the Sierra de Cazorla and authorized my work there.
-
-\newpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% References
-%\section{References}
-\bibliographystyle{ecology_letters} % Choose EcolLett style for bibliography
-%\bibliography{/Users/pedro/Documents/Databases/Bib_database}
-\bibliography{MS_LDD}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\end{linenumbers}
-
-\end{spacing}
-
-\end{flushleft}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%:----MS NOTES
-\begin{comment}
-
-Breeding unit size.
-Donor number is related to the size of the breeding unit (N individuals) as d= Np, where d is the number of different pollen parent genotypes represented in a maternal treeÕs fruit crop and p is the expected proportion of trees in staminate phase at any given time.
-Assuming,conservatively, that individual trees reproduce asynchronously twice per year 6 and that staminate and pistillate flowering phases each last seven days 20, p is calculated to be 0.0712.
-Although these estimates could be effected by mosaicism 21 (the fusion of genetically distinct individuals), its frequency in the species examined here is low.
-
-Breeding unit area and radii.
-These parameters were calculated from paternity-analysis-based estimates of breeding unit size (Nödö; Table2) and the censused densities of adult, reproductively mature trees over 15 km2 BCI.
-Based on the long-term censuses of C.
-Handley and E.
-Kalko, 6, 108 and 20 adults of F.
-dugandii, F.
-obtusifolia and F.
-popenoei, respectively, are known to occur on BCI.
-Because these species exhibit little spatial aggregation over the area censused, these densities are assumed to be representative of forested areas surrounding BCI, a conservative assumption given that approximately one-third of the area within 10 km of BCI is occupied by Lake Gatun where figs are absent.
-Breeding unit radii estimate the distances pollen-bearing, female fig wasps routinely disperse in search of receptive host trees.
-Although actual breeding populations of figs may deviate substantially from the assumed circular distribution, alternative structures (elliptical,for example)increase estimated wasp dispersal distances.
-
-- LevinÕs paternity pool concept.  Levin, D. A. The paternity pools of plants. Am. Nat. 132, 309Ð317 (1988).
-
-[Burnham, K.P. & Overton, W.S. Robust estimation of population size when capture probabilities vary among animals. Ecology 60, 927Ð936 (1979).]
-
-\end{comment}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TABLES
-\newpage
-%-------------------------------------------------------------------- Table 1
-\begin{table}
-%\captionsetup{width=15cm}%.75\textwidth}
-\caption{Types of dispersal as a function of population area limits and genetic neighborhood limits.
-LDD events can take place in a combination of circumstances, not necessarily involving dispersal outside the population.}
-\vspace{0.5cm}
-
-\begin{tabular}{llcccc}
-\addlinespace
-\toprule
-      & &	\multicolumn{2}{c}{Population limit }\\
-      &	&	Within	&	Outside \\
-\toprule
-
-Genetic	& Within	& Local, short-distance 		&Within neighborhood, long-\\
-neighborhood	&	&dispersal	&	distance dispersal\\
-limit	&	&	$SDD_{loc}$	&	$LDD_{neigh}$\\
-	&	Outside &	Local, long-distance	&	Strict sense long-distance\\
-	&	&	dispersal	&	dispersal\\
-	&	&	$LDD_{loc}$	&	$LDD_{SS}$\\
-
-\toprule
-
-\end{tabular}
-\end{table}
-%-------------------------------------------------------------------- Table 1
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURES
-
-\newpage
-%\section*{Figures}
-%------------------------------------------------------------------- Figure 1
-\begin{figure}[htbp]
-\centerline{\includegraphics[height=20cm]{Fig1.pdf}}
-\caption{Types of dispersal events related to population limits and genetic neigborhood size.
-The grey area depicts the breeding unit area.
-Depending on its size it may overlap with neighbour populations (bottom) or be much smaller than the population limits (top).
-Consequently, strcit sense LDD events $LDD_{SS}$ should be considered those beyond both the ecological and breeding unit area limits (top).
-LDD events within a local population ($LDD_{loc}$) can occur when propagules disperse beyond the local limits of the reeding unit area.
-Propagules exiting the population ($LDD_{neigh}$ but moving within the breeding unit area should not be considered strict sense LDD events.
-Short distance dispersal events are those occurring within the limits of both the population and breeding unit area $SDD_{loc}$.}
-\end{figure}
-%----------------------------------------------------------------------------
+\message{ !name(MS_LDD.tex) !offset(0) }
+dispersal
+\message{ !name(\string~/Documents/Working/MS_LDD/Manuscript/_region_.tex) !offset(153) }
 
 \end{document}
+
+\message{ !name(MS_LDD.tex) !offset(-194) }

### 2015-02-15 Data/kernels.Rmd
Initial commit of directory structure.

---
title: "Density estimations for dispersal kernels"
author: "Pedro Jordano"
date: "enero 29, 2015"
output: html_document
bibliography: refs.bib
---

Code for ploting diffreent forms of dispersal kernels and testing dispersal 
model functions.

Data refer to seed dispersal distances estimated by direct assignement from
seed endocarp genotypes

#### Plotting the kernels

```{r plain, fig.height=5, fig.width=8}
# Plain frequency distribution of dispersal distances.
require(MASS)
x<-assdist$dist
truehist(x,xlim=c(0,20000),prob=F,h=100,col=1,xlab="Distance (m)",ylab="Probability",ymax=350)
rug(x,side=1,col="red")

```

#### Fitting models

```{r, echo=FALSE}

```

----------------------------------
```{r}
sessionInfo()
```
----------------------------------

### References

### 2015-02-15 Data/code/assdist.R

@@ -1,0 +1,211 @@
+################################################################################
+#### Long distance dispersal analysis
+#### 8 Oct 2006. Modified 28 Jan 2015.
+################################################################################
+source("code/fit.disp.R")
+source("code/fitModels.R")
+
+assdist <-read.table("datasets/distances.txt",header=TRUE,sep="\t",dec=",",na.strings="NA")
+
+str(assdist)
+`data.frame:	557 obs. of  15 variables:
+ $ cohort     : int  1997 1997 1997 1997 1997 1997 1997 1997 1997 1997 ...
+ $ seedcode   : Factor w/ 557 levels "S_001","S_002",..: 1 2 3 4 5 6 7 9 10 11 ...
+ $ endocarp   : Factor w/ 557 levels "1843_C_T1","1843_C_T2",..: 23 24 25 26 38 27 28 30 31 32 ...
+ $ tree_source: Factor w/ 115 levels "1065","1069",..: 55 54 54 110 75 22 84 50 101 82 ...
+ $ x1.trap    : num  175.7 130.6  75.5 270.4 200.7 ...
+ $ y1.trap    : num  465 196 176 783 329 ...
+ $ x2.tree    : num  132 143 143 374 233 ...
+ $ y2.tree    : num   146  143  143 1016  413 ...
+ $ dist       : num  321.5  55.0  75.1 254.5  89.9 ...
+ $ censor     : Factor w/ 2 levels "c","e": 2 2 2 2 2 2 2 2 2 2 ...
+ $ loc        : Factor w/ 3 levels "DCHA","ESQ","IN": 1 1 1 2 3 3 3 3 3 3 ...
+ $ point      : Factor w/ 137 levels "","A1","A10",..: 42 78 20 126 132 43 65 67 73 88 ...
+ $ code       : Factor w/ 147 levels "","DCHAA10","DCHAA3",..: 6 12 13 22 142 60 79 81 87 100 ...
+ $ mhab       : Factor w/ 7 levels "A","AA","N","P",..: 1 4 4 7 7 1 3 3 3 4 ...
+ $ popsource  : Factor w/ 9 levels "CL","CM","CT",..: 4 4 4 4 4 4 4 4 4 4 ...
+summary(assdist$dist)
+   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+    0.0    13.0    62.2   574.0   273.0 16400.0 
+
[…]

### 2015-02-15 Data/code/bootkernelplot.R

@@ -1,0 +1,61 @@
+#------------------------------------------------------
+# Plot randomized kernel with specific mean and SD values.
+
+# Dispersal Kernel
+dd<-abs(rnorm(1000, mean = 20, sd = 415))
+
+# We can omit the truehist and step directly to the Kernel.
+truehist(dd, xlim=c(0,max(dd)),
+         # ylim=c(0,0.012),
+         prob=T,h=10,xlab="Distance (m)",
+         ylab="Probability",col=rgb(0, 0, 1, 0.2),
+         lty=0)
+rug(x,side=1,col="red")
+
+# Kernel
+d<-density(kern$dist,bw=20,from=0,to=max(dd)) # add density estimate
+lines(d,xlim=c(0,max(dd)),col="blue")
+
+# These are resampled (randomized) values to visualize a confidence envelope.
+for (k in 1: 100) {
+	di<- sample(dd, length(dd), replace = T, prob = NULL)
+	ddi<-density(di, bw=20, from=0, to=max(dd)) # add density estimate
+	lines(ddi,xlim=c(0, max(dd)), col="grey", lwd=0.2)
+
+# TDKs
+# Model species-specific kernels with different means and variances.
+
+# Dispersal Kernel
+dd1<- abs(rnorm(1000, mean = 5, sd = 10))
+dd2<- abs(rnorm(1000, mean = 20, sd = 40))
+dd3<- abs(rnorm(1000, mean = 20, sd = 120))
+dd4<- abs(rnorm(1000, mean = 40, sd = 445))
+
+totd<- c(dd1,dd2,dd3,dd4)
+
+# Kernels
+# Total kernel (histogram and rug)
+truehist(totd, xlim=c(0,max(totd)),
+         # ylim=c(0,0.012),
+         prob=T,h=10,xlab="Distance (m)",
+         ylab="Probability",col=rgb(0, 0, 1, 0.2),
+         lty=0)
+rug(x,side=1,col="red")
+
+# Species-specific kernels
+d1<-density(dd1,bw=20,from=0,to=max(dd1)) # add density estimate
+lines(d1,xlim=c(0,max(dd1)),col="blue")
+d2<-density(dd2,bw=20,from=0,to=max(dd2)) # add density estimate
+lines(d2,xlim=c(0,max(dd2)),col="orange")
+d3<-density(dd3,bw=20,from=0,to=max(dd3)) # add density estimate
[…]

### 2015-02-15 Data/code/dist_function.R

@@ -1,0 +1,14 @@
+dist<-function(x1, y1, x2, y2)  sqrt((x2 - x1)^2 + (y2 - y1)^2)
+
+r<-numeric(258)
+nn<-numeric(258)
+d<-numeric(258)
+for (i in 1:258) {
+d<-0
+for (k in 1:258) d[k]<-dist(x[i],y[i],x[k],y[k])
+r[i]<-min(d[-i])
+nn[i]<-which(d==min(d[-i]))
+}
+
+
+for (i in 1:258) lines(c(x[i],x[nn[i]]),c(y[i],y[nn[i]]))

### 2015-02-15 Data/code/jlgc.R

@@ -1,0 +1,277 @@
+### New dataset created with data 28 OCT 2004 ###
+R> jlgc <-read.table("seedrain9799.txt",header=TRUE,sep="\t",dec=",",na.strings="NA")
+R> pmah.nch <-read.table("pmah_nch9799seedrain.txt",header=TRUE,sep="\t",dec=",",na.strings="NA")
+R> jlgc.nch <-read.table("seedrain_nch9799.txt",header=TRUE,sep="\t",dec=",",na.strings="NA")
+R> ls()
+[1] "jlgc"      "jlgc.nch"  "pmah.nch"  "pmahtraps" "sdbvul"    "sdcmon"   
+R> names(jlgc.nch)
+  [1] "ord"              "site"             "site2"            "mhab"            
+  [5] "mhab2"            "trap"             "name"             "ox"              
+  [9] "oy"               "x"                "y"                "spp.pral"        
+ [13] "pl.agra97"        "seed.agra9798"    "pl.agra98"        "seed.agra9899"   
+ [17] "pl.agra99"        "pl.aita97"        "seed.aita9798"    "pl.aita98"       
+ [21] "seed.aita9899"    "pl.aita99"        "pl.aova97"        "seed.aova9798"   
+ [25] "pl.aova98"        "seed.aova9899"    "pl.aova99"        "pl.bvul97"       
+ [29] "seed.bvul9798"    "pl.bvul98"        "seed.bvul9899"    "pl.bvul99"       
+ [33] "pl.cgra97"        "seed.cgra9798"    "pl.cgra98"        "seed.cgra9899"   
+ [37] "pl.cgra99"        "pl.cmon97"        "seed.cmon9798"    "pl.cmon98"       
+ [41] "seed.cmon9899"    "pl.cmon99"        "pl.dlau97"        "seed.dlau9798"   
+ [45] "pl.dlau98"        "seed.dlau9899"    "pl.dlau99"        "pl.hhel97"       
[…]

### 2015-02-15 Data/code/levelplot.R

@@ -1,0 +1,300 @@
+### New dataset created with data 28 OCT 2004 ###
+R> jlgc <-read.table("seedrain9799.txt",header=TRUE,sep="\t",dec=",",na.strings="NA")
+R> pmah.nch <-read.table("pmah_nch9799seedrain.txt",header=TRUE,sep="\t",dec=",",na.strings="NA")
+R> jlgc.nch <-read.table("seedrain_nch9799.txt",header=TRUE,sep="\t",dec=",",na.strings="NA")
+R> ls()
+[1] "jlgc"      "jlgc.nch"  "pmah.nch"  "pmahtraps" "sdbvul"    "sdcmon"   
+R> names(jlgc.nch)
+  [1] "ord"              "site"             "site2"            "mhab"            
+  [5] "mhab2"            "trap"             "name"             "ox"              
+  [9] "oy"               "x"                "y"                "spp.pral"        
+ [13] "pl.agra97"        "seed.agra9798"    "pl.agra98"        "seed.agra9899"   
+ [17] "pl.agra99"        "pl.aita97"        "seed.aita9798"    "pl.aita98"       
+ [21] "seed.aita9899"    "pl.aita99"        "pl.aova97"        "seed.aova9798"   
+ [25] "pl.aova98"        "seed.aova9899"    "pl.aova99"        "pl.bvul97"       
+ [29] "seed.bvul9798"    "pl.bvul98"        "seed.bvul9899"    "pl.bvul99"       
+ [33] "pl.cgra97"        "seed.cgra9798"    "pl.cgra98"        "seed.cgra9899"   
+ [37] "pl.cgra99"        "pl.cmon97"        "seed.cmon9798"    "pl.cmon98"       
+ [41] "seed.cmon9899"    "pl.cmon99"        "pl.dlau97"        "seed.dlau9798"   
+ [45] "pl.dlau98"        "seed.dlau9899"    "pl.dlau99"        "pl.hhel97"       
[…]

### 2015-02-15 Data/code/newspatial.R

@@ -1,0 +1,162 @@
+####################################################################
+# New analyses. Zahara, Mar 2010
+# Pedro Jordano.
+####################################################################
+# Setting ppp objects for tree maps
+prumap <-read.table("prumap.dat",header=TRUE,sep="\t",dec=".",na.strings=".")
+attach(prumap)		# All trees, N= 258, incl. CSABINAS
+names(prumap)
+[1] "MICROHAB" "POINT"    "NAME"     "TREE"     "x"        "y"       
+prumap.ppp <- ppp(prumap$x, prumap$y, 
+                  window=owin(c(0,500),c(0,1100)),
+                  unitname=c("metre","metres"))
+plot(prumap.ppp,pch=16)
+
+attach(prumap196)		# Cris trees, N=196, incl. CSABINAS
+tree196.ppp <- ppp(x, y,                   
+                   window=owin(c(0,500),c(0,1100)),
+                   unitname=c("metre","metres")),
+                   marks=prumap196$SEX)
+plot(tree196.ppp)
+? F H 
+1 2 3 
+
+summary(tree196.ppp)
+Marked planar point pattern: 196 points
+Average intensity 0.000349 points per unit area
+Marks:
+  frequency proportion intensity
+?        32      0.163  0.000057
+F        62      0.316  0.000111
+H       102      0.520  0.000182
+
+Window: rectangle = [ -10 , 500 ] x [ 0 , 1100 ]
+Window area =  561000 
+
+# ppp objects for trap data 1997-2000-JLGC
+attach(traps.9700)
+sd9798 <- ppp(x, y, c(0,500), c(0,1200), marks=sd9798)
+sd9899 <- ppp(x, y, c(0,500), c(0,1200), marks=sd9899)
+sd9900 <- ppp(x, y, c(0,500), c(0,1200), marks=sd9900)
+sdl97 <- ppp(x, y, c(0,500), c(0,1200), marks=sdl97)
+sdl98 <- ppp(x, y, c(0,500), c(0,1200), marks=sdl98)
+sdl99 <- ppp(x, y, c(0,500), c(0,1200), marks=sdl99)
[…]

### 2015-02-15 Data/code/pmahtraps.R

@@ -1,0 +1,114 @@
+R> pmahtraps <-read.table("pmahtraps9700.txt",header=TRUE,sep="\t",dec=".",na.strings="NA")
+R> save.image()
+R> ls()
+ [1] "R_LIBS"       "dist"         "jlgc"         "last.warning" "mmhab"       
+ [6] "pm.pp"        "pmahtraps"    "prugeo"       "prumap"       "prumap.pp"   
+[11] "sd9798"       "sd9899"       "sd9900"       "sdl97"        "sdl98"       
+[16] "sdl99"       
+R> str(pmahtraps)
+`data.frame':	615 obs. of  21 variables:
+ $ site      : Factor w/ 3 levels "DRCHA","ESQ",..: 1 1 1 1 1 1 1 1 1 1 ...
+ $ mhab      : Factor w/ 9 levels "A","N","P","Pos..",..: 1 1 1 1 1 1 1 1 1 1 ...
+ $ punto     : Factor w/ 83 levels "","1","10","11",..: 2 13 24 35 46 59 70 81 83 3 ...
+ $ label     : Factor w/ 614 levels "DRCHAA1","DRCHA..",..: 1 3 4 5 6 7 8 9 10 2 ...
+ $ ox        : num  -142.4  -87.8 -110.7 -129.4 -123.7 ...
+ $ oy        : num  -174 -139 -193 -146 -142 ...
+ $ x         : num   57.6 112.2  89.3  70.6  76.3 ...
+ $ y         : num  126 161 107 154 158 ...
+ $ site.1    : Factor w/ 2 levels "IN","OUT": 2 2 2 2 2 2 2 2 2 2 ...
+ $ subsite   : Factor w/ 6 levels "DRCHA","ESQ",..: 1 1 1 1 1 1 1 1 1 1 ...
+ $ micro     : Factor w/ 7 levels "","A","N","P",..: 2 2 2 2 2 2 2 2 2 2 ...
+ $ microcampo: Factor w/ 8 levels "","A","N","P",..: 2 2 2 2 2 2 2 2 2 2 ...
+ $ punto2    : Factor w/ 83 levels "","1","10","11",..: 2 13 24 35 46 59 70 81 83 3 ...
+ $ seedl.97  : num  0 0 1 0 0 0 0 3 0 0 ...
+ $ seed.9798 : num  0 0 13 0 9 0 44 29 5 1 ...
+ $ seedl.98  : num  0 0 14 0 0 0 0 9 0 0 ...
[…]

### 2015-02-15 Data/code/seedlings.R

@@ -1,0 +1,4 @@
+##############################################################################
+# Seedlings genotypes. 2006 sampling.
+##############################################################################
+seedl <-read.table("seedlings.txt",header=TRUE,sep="\t",dec=",",na.strings="NA")

### 2015-02-15 Data/code/seedshad.R

@@ -1,0 +1,79 @@
+R> ls()
+ [1] "R_LIBS"       "dist"       "last.warning" "mmhab"   "pm.pp"   "prugeo"
+ [7] "prumap"       "prumap.pp"  "sd9798"       "sd9899"  "sd9900"  "sdl97"
+[13] "sdl98"        "sdl99"      "tt"          
+R> names(sd9899)
+[1] "window" "n"      "x"      "y"      "marks" 
+R> data(julliot)
+R> names (julliot)
+[1] "tab"  "xy"   "area"
+
+R> s.image(julliot$xy, log(julliot$tab[,k]+1), kgrid = 3, span = 0.25)
+
+R> str(julliot$tab)
+`data.frame':	160 obs. of  7 variables:
+ $ Pouteria_torta             : int  0 0 0 0 0 0 0 0 0 0 ...
+ $ Minquartia_guianensis      : int  0 0 0 0 0 0 0 0 0 0 ...
+ $ Quiina_obovata             : int  0 0 0 0 0 0 0 0 0 0 ...
+ $ Chrysophyllum_lucentifolium: int  0 0 0 0 1 0 1 0 0 0 ...
+ $ Parahancornia_fasciculata  : int  0 0 0 0 0 0 0 0 0 0 ...
+ $ Virola_michelii            : int  0 1 0 6 1 2 1 0 0 0 ...
+ $ Pourouma_spp               : int  0 0 0 0 0 0 0 0 0 0 ...
+
+R> names(sd9899)
+[1] "window" "n"      "x"      "y"      "marks" 
+
+R> str(sd9899)
+List of 5
+ $ window:List of 3
+  ..$ type  : chr "rectangle"
+  ..$ xrange: num [1:2] 0 470
+  ..$ yrange: num [1:2] 0 1150
+  ..- attr(*, "class")= chr "owin"
+ $ n     : int 615
+ $ x     : num [1:615]  57.6 112.2  89.3  70.6  76.3 ...
+ $ y     : num [1:615] 126 161 107 154 158 ...
+ $ marks : num [1:615] 0 0 4 0 6 0 6 11 2 0 ...
+ - attr(*, "class")= chr "ppp"
+
+ R> prunus9899<- new.env()
+R> prunus9899$tab<- sd9899$marks
+R> names(julliot$xy)
+[1] "x" "y"
+
[…]

### 2015-02-15 Data/code/spatial.R

@@ -1,0 +1,265 @@
+5-15 Feb 2004
+
+ ls()
+ [1] "dist"         "last.warning" "mmhab"        "pm.pp"        "prugeo"      
+ [6] "prumap"       "prumap.pp"    "sd9798"       "sd9899"       "sd9900"      
+[11] "sdl97"        "sdl98"        "sdl99"        "tt"          
+
+--- 
+dev.copy2eps("distplot&map.eps",width=6,height=6)
+dev.print(device=pdf,"plotout2.pdf")
+---
+
+#----------
+# distance between two points
+dist<-function(x1, y1, x2, y2)  sqrt((x2 - x1)^2 + (y2 - y1)^2)
+
+#  to estimate nearest neighbor for n points and draw lines
+# after plotting. Substitue 100 by no. of points 
+
+r<-numeric(258)
+nn<-numeric(258)
+d<-numeric(258)
+for (i in 1: 258) {
+d<-0
+for (k in 1: 258) d[k]<-dist(x[i],y[i],x[k],y[k])
+r[i]<-min(d[-i])
+nn[i]<-which(d==min(d[-i]))						
+}
+# plot the locations of seed traps and the lines joining to each source tree
+for (i in 1: 258) lines(c(x[i],x[nn[i]]),c(y[i],y[nn[i]]))   # draw the lines
+#-------------
+
+# Setting the ppp objects for seeds and seedlings
+attach(traps.9700)
+R > names(traps.9700)
+ [1] "SITE"       "MHAB"       "POINT"      "LABEL"      "OX"        
+ [6] "OY"         "x"          "y"          "AREA"       "SITE2"     
+[11] "MICRO"      "MICROCAMPO" "sdl97"      "sd9798"     "sdl98"     
+[16] "sd9899"     "sdl99"      "sd9900"     "sd"         "sdl"       
+plot(x,y)
+sd9798 <- ppp(x, y, c(0,500), c(0,1200), marks=sd9798)
+sd9899 <- ppp(x, y, c(0,500), c(0,1200), marks=sd9899)
+sd9900 <- ppp(x, y, c(0,500), c(0,1200), marks=sd9900)
+sdl97 <- ppp(x, y, c(0,500), c(0,1200), marks=sdl97)
+sdl98 <- ppp(x, y, c(0,500), c(0,1200), marks=sdl98)
+sdl99 <- ppp(x, y, c(0,500), c(0,1200), marks=sdl99)
[…]

### 2015-02-15 Data/kernels.Rmd
Edited .Rmd file; added GEV functions.

@@
 
-Code for ploting diffreent forms of dispersal kernels and testing dispersal 
-model functions.
+Code for ploting different forms of dispersal kernels and testing dispersal model functions.
 
-Data refer to seed dispersal distances estimated by direct assignement from
-seed endocarp genotypes
+Data refer to seed dispersal distances estimated by direct assignement from seed endocarp genotypes for *Prunus mahaleb*.
 
-#### Plotting the kernels
+#### Data input
 
-```{r plain, fig.height=5, fig.width=8}
-# Plain frequency distribution of dispersal distances.
+```{r data, message=FALSE,warning=FALSE}
+# Input data.
 require(MASS)
+require(downloader)
+require(ismev)
+require(adehabitatHS)               # For simulations of Levy walks
+
+# Prunus mahaleb dispersal distances data. -------------------------------
+# Get the data from my GitHub repository.
+link = "https://raw.githubusercontent.com/pedroj/dispkernels/master/distances.txt"
+file = "distances.txt"
+if(!file.exists(file)) download(link, file, mode = "wb")
+assdist <- read.table(file, sep = "\t", dec = ",", 
+    header = TRUE, na.strings="NA")
+str(assdist)
+
+```
+
+#### Raw kernel
+
+```{r plot_hist}
+# Plain frequency distribution of dispersal distances. -------------------
 x<-assdist$dist
@@
 
-#### Fitting models
+#### Simulation of Lévy pattern
 
-```{r, echo=FALSE}
+```{r levy, fig.height=5, fig.width=8}
+# Lévy distribution. -----------------------------------------------------
+# dd<- abs(rnorm(1000, mean = 32.7, sd = 415)) # Normal distrib.
+# dd<- rweibull(1000, shape= 1, scale = 3.5)   # Weibull distrib.
+# dd<- rgamma(1000, scale= 100, shape= 10)     # Gamma distrib.
+                                               # req. adehabitatHS
+dd<- as.data.frame(simm.levy(1:1000, mu = 1.5, burst = "mu = 1.5")) 
+dd<- na.omit(dd$dist)
+
+# Sourcing the kernel function code from GitHub.
+link = "https://raw.githubusercontent.com/pedroj/dispkernels/master/functions/mykernel.R"
+file = "./code/mykernel.R"
+if(!file.exists(file)) download(link, file, mode = "wb")
+source(file)
+
+# With function mykernel.
+mykernel(dd, bw= 20, h= 200)     # Extract the distances vector
+ddtrunc=dd[dd<3000]              # Truncate below a distance
+mykernel(ddtrunc, bw= 10, h= 20) # Extract the distances vector
+
+```
+
+#### Analysis of generalized extreme value (GEV) distributions
+
+```{r GEV}
+# Prunus mahaleb dispersal distances data. -------------------------------
+source("./code/GEV_functions.R")
+
+dd<- assdist$dist
+# With function mykernel.
+mykernel(dd, bw= 100, h= 50) # Extract the distances vector
+gevdd<-gev.fit(dd)
+
+# Notice the output:
+# ∗ $conv gives a value of zero (in row[1] of the output), which 
+#   indicates success- full convergence, i.e. no errors in fitting;
+# In the case of the full dispersal kernel I get convergence errors!!
+# ∗ $nllh shows the negative (maximised) log–likelihood;
+# ∗ $mle shows the maximum likelihood estimates for μ, σ and ξ respectively;
+# ∗ $se gives the associated standard errors for these parameters.
+
+gev.diag(gevdd)
+
+# Truncate events after 1200 m to get the within-stand dispersal events
+ddtrunc=dd[dd<1200]
+mykernel(ddtrunc, bw= 10, h= 10) # Extract the distances vector
+gevdd<-gev.fit(ddtrunc)
+gev.diag(gevdd)
+gev.ret(ddtrunc,100)
+# If we want to construct a confidence interval for q100, we are better off using the method of profile–likelihood as described in Section 1.2.8.
+We can use the function gev.prof(fit,period,lower-bound, upper-bound).
+This com- mand is slightly unstable, and relies on an appropriate choice of the bounds for the profile–likelihood.
+For the Boston annual maxima, the following works well for the 100–year level:
+gev.prof(gevdd,100,60,150)
+
+# Note that this enables us to read off the 95% confidence interval (the default) for q100. Suppose we wanted a 99% interval we would use:
+gev.prof(gevdd,100,60,150,conf=0.99)
 

### 2016-06-17 Manuscript/MS- What is LDD.tex
Updated edited version of MS and bib file

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% MS Draft Journal of Ecology Special Issue, Dispersal - 
%% Pedro - 07 April 2016
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Headers

\documentclass[a4paper, 12pt]{article}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[a4paper]{geometry}
\usepackage{hyperref}
\pagestyle{plain}
\usepackage{amsmath,amssymb}
\usepackage{geometry}
\usepackage{lscape}
\usepackage{setspace}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage{longtable}
\usepackage{rotating}                                                                                    
\newcommand{\tab}{\hspace{5mm}}
\usepackage{tabularx} 
\usepackage[margin=10pt,font=small,labelfont=bf]{caption}
\usepackage[left]{lineno}
\usepackage{caption}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy}     % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{{\tiny Jordano - What is long-distance dispersal?}}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}
\usepackage{parskip}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title page
\begin{document}

\title{Manuscript Draft\\
\vspace{2cm}
What is long-distance dispersal? And a taxonomy of dispersal events}

\author{Pedro Jordano$^{\dag}$}

\date{Sevilla, \today}
\maketitle

\begin{spacing}{1.0}
$^{\dag}$ {\small Integrative Ecology Group, Estaci\'on Biol\'ogica de 
Do\~nana, CSIC, Avda. Americo Vespucio, s/n, Isla de La Cartuja
E-41092 Sevilla, Spain.}\\

{\small \textit{Corresponding author:} Pedro Jordano. Integrative Ecology Group, Estaci\'on Biol\'ogica de Do\~nana, CSIC, Avda. americo Vespucio, s/n, E-41092 Sevilla, Spain. Email address: jordano@ebd.csic.es}\\

\textbf{Key words}: ***\\

{\small \textbf{Manuscript information: }** Words; ** Chars; ** Pages, * Figures; * Tables.}
\end{spacing}

\maketitle
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Abstract
\section*{Abstract}
\begin{linenumbers}

Dispersal is a key individual-based process influencing many life-history attributes, scaling up to population-level properties (e.g., metapopulation connectivity). A persistent challenge in dispersal ecology has been the robust characterization of dispersal functions (kernels), a fundamental tool to predict how dispersal processes respond under global change scenarios. Especially the rightmost tail of these functions, i.e. the long-distance dispersal (LDD) events, are difficult to characterize empirically and to model in realistic ways. But, when is it a LDD event? In the specific case of plants, dispersal has three basic components: 1) a distinct (sessile) source, the maternal plant producing the fruits or the paternal tree acting as a source of pollen; 2) a distance component between source and target locations; and 3) a vector actually performing the movement entailing the dispersal event. Here we discuss operative definitions of LDD based on their intrinsic properties: 1) events crossing geographic boundaries among stands; and 2) events contributing to effective gene flow and propagule migration. LDD is a characteristically extreme event of propagule movement in any plant population, typically occurring with an extremely low probability but potentially reaching an extremely long distance. Strict-sense long distance involves movement both outside the stand geographic limits and outside the genetic neighborhood area of individuals. Combinations of propagule movements within/outside these two spatial reference frames results in distinct modes of LDD. Beyond traditional statistical approaches to characterize distributions, Extreme Value Analysis (EVA) can be used to properly and explicitly evaluate the properties of frequency and extent of LDD events. We discuss conditions where global change scenarios truncate dispersal processes, leading to the loss of key dispersal services in natural populations. Proper characterization of the LDD events helps to assess, for example, how the ongoing defaunation of large-bodied frugivores pervasively entails the loss of crucial LDD functions.

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Introduction
\section*{Introduction}

Dispersal is a key individual-based process influencing many life-history attributes and scaling up to population-level properties (e.g., metapopulation connectivity, \citealt{Cousens:2008aa}). In the specific case of plants, largely sessile organisms, dispersal has three basic components: 1) a distinct (sessile) source, the maternal plant producing the fruits or the paternal tree acting as a source of pollen; 2) a distance component between source and target locations; and 3) a vector actually performing the movement entailing the dispersal event. While realized dispersal also depends upon stages subsequent to dissemination (e.g., successful germination and seedling establishment) \cite{Schupp:1995}, the three previous components fully characterize the dispersal process per se. Therefore, plant movement differs in important natural history details from animal dispersal, yet both can be assessed within a common conceptual framework \citep[e.g., ][]{Nathan:2006aa}. Characteristically, animal-assisted plant dispersal has three distinct, highly integrated, components missing in the process of animal dispersal: the properties of the source (parental) plant, that mediate in the foraging of the animal vector (pollinator or frugivore), the intrinsic properties of the propagule, and the functional characteristics of the animal vector who performs the movement \citep{Nathan:2008fx}.

The movement of pollen and seeds by animals and its consequences have intrigued population geneticists and field ecologists since the infancy of both research disciplines. Each has generated an impressive body of theoretical and empirical research through the past decades, yet advances have long been co-existing in ‘parallel worlds’ and the great synergistic potential of population genetics and demography for the study of plant dispersal by animals remains little explored. Knowledge gaps still having the imprint of this conceptual disconnection include the idea of long distance dispersal, and the paradoxes of forest fragmentation effects on genetic diversity \citep{Kramer:2008kg}, survival and persistence of relict tree species \citep{Hampe:2011bv}, rapid post-Pleistocene recolonization of vast continental areas in response to climate modification \citep{Clark:1998aa,Clark:1998vi}, among other persisting issues. This conceptual isolation has been exacerbated by technical difficulties for the robust characterization of dispersal events, especially those involving movement over long-distances (long-distance dispersal, LDD). Some progress has recently been made through the fast-paced implementation of molecular tools in ecological research labs and the availability of cutting-edge technology for biotelemetry applications . But much of the population geneticist and ecologist communities remains unaware of the state of the art in each other and likely under-appreciates their potential to validate and enrich dispersal studies \citep{Jones:2008il}. In particular, LDD events remain difficult to assess, both technically- with serious methodological problems for its reliable estimation- and conceptually. Our aim here is to review the LDD concept with a specific emphasis on dispersal of plant propagules (seeds and pollen), providing an extended definition that might be helpful in the robust quantification of LDD events.

Two main conceptual approaches have been used to assess dispersal (Fig. 1). The “forward” approach attempts to track the dispersal events away from the known sources, e.g., by tracking the movement patterns of frugivores as they leave fruiting plants after feeding (Fig. 1A). This is the main approach used in the movement ecology framework \citep{Nathan:2008fx}, with extensive application to animal movement based on the use of advanced biotelemetry. The “backward” approach attempts to reconstruct the most likely source of a dispersed propagule by inferring the sources given the propagule delivery pattern, the fecundity of potential sources, and the dispersal function (Fig. 1B), i.e., by using an inverse modeling approach. The main technical challenge in Fig. 1A is to sample enough dispersal events away from the source to be able to fully characterize the tail (LDD events) of the dispersal function. In Fig. 1B, the main challenge is to have a robust sampling scheme with propagule collectors (e.g., seed traps) and a good characterization of the potential sources to derive robust estimates of the actual sources. Both approaches are limited logistically by the difficulties to sample the vast areas required to assess LDD events from the focal source population.

No explicit definition of what constitutes an LDD event exists. Previous approaches \citep[e.g., ][]{Nathan:2006aa,schurr2009long} include both absolute and proportional definitions to characterize LDD events. This means providing information about the absolute distances moved by a given percentile of the events and/or providing data on the proportion of events exceeding a given distance threshold \citep{Nathan:2008is}. The exact proportional or absolute thresholds selected remain arbitrary, as no reference spatial frame is provided within the definition of LDD. This leaves the consideration of LDD as an extreme form of context-dependent phenomenon, strongly dependent upon the scale of the biological process studied \citep{Kinlan:2005fb}. For example, \cite{Kinlan:2005fb} used a spatial reference frame to characterize LDD events of marine organisms, where sedentary adults and larvae differ enormously in the spatial scales of their dispersal \citep{DAloia:2013fc}. Therefore, any measure of extent and reach of LDD events requires reference to an explicit spatial frame or "local" scale \citep{Kinlan:2005fb}.

We aim at providing a general framework for the quantitative analysis of LDD events so that estimates of its frequency and extent could be comparable across different study systems. We argue that both demographic and genetic elements are needed for this framework, most likely requiring a combination of field-based movement data and genetic analyses. These elements can be overlaid on previous definitions based on absolute and proportional characterizations of LDD. We start with a definition of LDD events within a spatially-explicit mechanistic framework allowing an unambiguous meaning for setting long-distance thresholds. We then use a case study to assess differential contributions of animal frugivores performing LDD.

Long-distance dispersal is currently one of the most debated topics in dispersal ecology; it defines the connectedness within the network of local populations and the possibilities for range expansion and successful colonization events. We propose a first demogenetically-based, operational definition of what a long-distance dispersal event actually is, and review existing empirical literature on distance thresholds from population and genetic perspectives. We also show how molecular tools have been used to identify the respective contributions of different animal species to the LDD portion of dispersal kernels of pollen and seeds by setting empirically-derived distance thresholds. Finally, we highlight potential applications of molecular markers beyond the quantification of just the dispersal distances that prevails in current studies, e.g., experimental approaches to assess dispersal limitation and Janzen-Connell effects.

\subsection*{LDD within a demo-genetic perspective: a taxonomy of dispersal events}

Here we propose an explicit definition of LDD and what constitutes a LDD event. Previous definitions of dispersal patterns emphasized only their distance components and characterized LDD events basically in terms of geographic distance between a dispersed propagule (or an established early seedling) and its most likely maternal or paternal (in case of pollen) source. Absolute and proportional definitions for the LDD events have been proposed depending on arbitrary thresholds of either the distance beyond which a dispersal event is LDD or the proportion of events occurring beyond a specific distance \citep{Nathan:2005jc, Nathan:2008is}. Thus, two key biological aspects of LDD events involve the transport of propagules outside a reference area: moving away from the source stand or population, and moving away from the area where relatives stand \citep{Kinlan:2005fb}. These two movements do not necessarily concur: a propagule may move over a very long distance yet still be disseminated within the reach of the neighborhood where parental individuals mate. Within a demogenetic framework it is easy to envision a combination of situations concerning the spatial scale of the dispersal processes (Table 1) and unambiguously define different types of LDD events. The idea that dispersal occurs in reference to these two spatial reference frames, i.e., the population or stand and the genetic neighborhood area, is motivated by the fact that dispersal entails the movement of both an individual propagule (i.e., a pollen grain or a seed) and a distinct set of genes (i.e., the male genotype in case of pollen, or a seed genotype). Thus, dispersal entails simultaneous demographic and genetic effects through recruitment of new individuals in the population and through contributions to gene flow \citep{Harper:1977aa}.

Two important components of plant dispersal ecology concern the movement of propagules away from the source population, a type of dispersal relevant to colonization ability and range expansion, and the movement away from the location of close relatives, i.e., a movement away from the genetic neighborhood. If we classify dispersal events according to these two spatial frameworks (Table 1) we end up with four distinct types of events depending on whether or not dispersed propagules are disseminated within these reference areas. Setting the limits of a population can be problematic yet we can identify with relative ease the geographical limits of plant stands, patches, habitat spots or other types of habitat or microhabitat discontinuities \citep[see][for further discussion of boundaries for dispersal]{Kinlan:2005fb}. These "frontiers" set biological limits to what a LDD event is in relation to the geographic limits of the source population. Most plants are distributed as clumped patches, discrete stands, or relatively isolated populations, so we may distinguish between short-distance and long-distance dispersal events that end up with dissemination within or beyond, respectively, the stand or population geographic boundaries (Table 1, $SSD_{loc}$ or $LDD_{loc}$) (Figure 2).

\subsection*{Individual and Population Neighborhoods as Reference}

Continuous populations can be modeled with the concepts of isolation by distance and neighborhood size\citep{Wright:1943aa,Wright:1946aa}. The former refers to the case that limited gene dispersal in continuous populations produces demes that are panmictic internally, but are isolated to some extent, from adjacent demes. Each group of reproducing individuals is the neighborhood, defined as the population of a region in a continuum, from which the parents of individuals born near the center may be treated as if drawn at random \citep{Wright:1969mb}. The importance and influence of the dispersal process in determining the size of the neighborhood is given by this equation, which shows how the spatial dispersion (pattern of spatial distribution) of the population influences $N_e$. This influence on the effective size is given by:

\begin{equation}
					N_e= 4 \pi \sigma^2 \delta
\end{equation}

where $\sigma^2$ is the variance of the dispersal distance and $\delta$ is the density of individuals. This formulation is often called the neighborhood size and assumes a normal distribution of distances  between parents and offspring (out in a perfect circular shape from the source). Thus, changes in the variance of dispersal size can affect $N_e$ (highly clumped populations will have reduced $N_e$).

\begin{equation}
					N_b= 4p \sigma^{2d}
\end{equation}

where $d$ is the density of adults per unit area and $\sigma$ is the variance in distance between birth and breeding sites. This is the basic model of ‘Isolation by Distance’ proposed by   \citet{Wright:1943aa,Wright:1946aa}. Under this type of model, migration (gene flow) is given by the variance in dispersal, and not by the proportion of the population that is composed of migrants (denoted $m$), as is the case with island models \citep{Slatkin:1985qb}.

For plants, gene flow may be accomplished by both seeds and pollen, so the variance may be decomposed to account for different patterns of seed and pollen dispersal, and to take into account the mating system (outcrossing rate, $t$). Thus, neighborhood size can be defined with the following equation (Crawford 1984) :

\begin{equation}
  					N_b = 4p (\sigma^2_s + \frac{t \sigma^2_p}{2}) d (1 + t)
\end{equation}

where $\sigma^2_s$ is the variance in seed dispersal, $\sigma^2_s$ is the variance in pollen dispersal and $d$ is the density of potential parents.

Neighborhood size in plants can be estimated by marking pollen and seeds with fluorescent dyes, tags, or  marks. However, these methods do not measure effective pollen or seed movement, but they may be combined with genetic analysis to do so.  Individuals with a unique allele in a stand can provide valuable insight on seed movement \citep{Eguiarte:1993aa}.

\paragraph*{Empirical Studies of Seed Dispersal}

\subsection*{Long-Distance Dispersal: the ecology of extreme events}

Long-distance dispersal (LDD) is a major component of the population dynamics, genetic structure, and biogeographic history of plant species. It determines the colonization ability of new habitats and the possibilities for fragmented populations to sustain a cohesive metapopulation by immigration-emigration dynamics that rely on LDD events. Yet our current understanding of the extent, frequency, and consequences of LDD is very limited. On one hand, theoretical models fail to predict accurately the behavior of the tail of the dispersal functions, and thus fail to predict very basic properties of LDD. On the other hand, we still have very limited documentation of actual LDD events in natural populations and we still see LDD as a sporadic, rarely far-reaching process still marked with the stamp of natural history curiosity.

As defined in our framework (Table 1), LDD, and in particular $LDD_{ss}$ events are a particular case of extreme events \citep{Garcia:2017aa}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Discussion
%\section*{Discussion}

\subsection*{Challenges and Promising Avenues for Research}

\emph{Acknowledgements}. I am indebted to Cristina García, José A. Godoy, Manolo Carrión, Juan Luis García-Castaño, Jesús Rodríguez and, especially, Juan Miguel Arroyo for generous help with field and laboratory work and making possible this study. I appreciate the help and advice of Cristina García and Etienne Klein during the final stages of the manuscript. The study was supported by a Junta de Andalucía Excellence Grant (RNM-5731), as well as a Severo Ochoa Excellence Award from the Ministerio de Economía y Competitividad (SEV-2012-0262). The Agencia de Medio Ambiente, Junta de Andalucía, provided generous facilities that made possible this study in the Andalusian natural parks (Sierra de Cazorla, Alcornocales) and authorized my work there.

\end{linenumbers}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TABLES
%-------------------------------------------------------------------- Table 1
\pagestyle{empty}
\begin{landscape}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	\oddsidemargin  0.0in
%	\evensidemargin 0.0in
%	\textwidth      8.5in
	\headheight     3cm
%	\topmargin      0.0in
%	\textheight=8.0in
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\captionsetup{width=20cm}%.75\textwidth}
\caption{Types of dispersal as a function of population area limits and genetic neighborhood limits.}
\vspace{0.5cm}
  \begin{tabular}{lll}
    \hline
\\&Population geographic limit    &  \\\\
Genetic neighborhood limit     &Within &Outside \\\\
    \hline
\\Within    &  Local, short-distance \linebreak dispersal, $SDD_{loc}$  
          &  Within neighborhood, \linebreak long-distance dispersal, $LDD_{neigh}$ \\\\
Outside   &     Local, long-distance \linebreak dispersal, $LDD_{loc}$    
          &  Strict sense \linebreak long-distance dispersal, $LDD_{ss}$\\\\ 
    \hline
  \end{tabular}
\end{table}
\end{landscape}

%----------------------------------------------------------------------------

\section*{References}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% References
%%% Unnumbered Literature Cited
\bibliographystyle{bes}        %Compile with bes.bst style file
\bibliography{MS_BES}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURES

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LANDSCAPE FIGURES
%%%% Figure LEGENDS
%%%% This is for sideways figures (landscape), with legend in a separate page.
\section*{Figures}
\begin{linenumbers}

\textbf{Figure 1}. The two approaches used in analyses of dispersal processes in plants. A, the “forward” approach attempts to track the dispersal events away from the known sources, e.g., by tracking the movement patterns of frugivores as they leave fruiting plants after feeding. B, the “backward” approach  attempts to reconstruct the most likely source of a dispersed propagule by inferring the sources given the propagule delivery pattern, the fecundity of potential sources, and the dispersal function. The main technical challenge in A is to sample enough dispersal events away from the source to be able to fully characterize the tail (LDD events) of the dispersal function. In B, the main challenge is to have a robust sampling scheme with propagule collectors (e.g., seed traps) and a good characterization of the potential sources to derive robust estimates of the actual sources.\\

\textbf{Figure 2}. Schematic representation of different types of long-distance dispersal events in relation to the geographical limits of local populations (dashed lines) and the genetic neighborhood size (grey area) of specific individual plants. Dispersal events (arrows) can be classified depending on their actual incidence on propagule movement outside these reference areas (Table 1).\\

\textbf{Figure 3}. LDD events depend on a combination of disperser characteristics and landscape features. While different ***.\\

\textbf{Figure 4}. Differential contributions of functional groups of frugivores to the seed dispersal of \textit{Prunus mahaleb}.\\ 
\newpage

%------------------------------------------------------------------- Figure 1
\begin{figure}[htbp]
\centerline{\includegraphics[height=25cm]{Fig1.pdf}}
%
%\caption{***}
\end{figure}

\newpage 
%------------------------------------------------------------------- Figure 2
\begin{figure}[htbp]
\centerline{\includegraphics[height=25cm]{Fig2.pdf}}
%
%\caption{***}
\end{figure}

\newpage 
%------------------------------------------------------------------- Figure 3
\begin{figure}[htbp]
\centerline{\includegraphics[height=26cm]{Fig3.pdf}}
%
%\caption{***}
\end{figure}
%----------------------------------------------------------------------------

\newpage 
%------------------------------------------------------------------- Figure 4
\begin{figure}[htbp]
\centerline{\includegraphics[height=25cm]{Fig4.pdf}}
%
%\caption{***}
\end{figure}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Online Support Material and data accessiblity}

This review does not use new raw data, but includes some re-analyses of previously published material. All the original data supporting the paper, R code, supplementary figures, and summaries of analytical protocols is available at the author's GitHub repository (\texttt{https://github.com/pedroj/MS\_LDD}), with DOI: \texttt{\#/zenodo.\#}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{linenumbers}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% MS NOTES
\begin{comment}
NOTES
Breeding unit size. Donor number is related to the size of the breeding unit (N individuals) as d= Np, where d is the number of different pollen parent genotypes represented in a maternal tree’s fruit crop and p is the expected proportion of trees in staminate phase at any given time. Assuming, conservatively, that individual trees reproduce asynchronously twice per year 6 and that staminate and pistillate flowering phases each last seven days 20, p is calculated to be 0.0712. Although these estimates could be effected by mosaicism 21 (the fusion of genetically distinct individuals), its frequency in the species examined here is low.

Breeding unit area and radii. These parameters were calculated from paternity-analysis-based estimates of breeding unit size (Nˆdˆ; Table2) and the censused densities of adult, reproductively mature trees over 15 km2 BCI. Based on the long-term censuses of C. Handley and E. Kalko, 6, 108 and 20 adults of F. dugandii, F. obtusifolia and F. popenoei, respectively, are known to occur on BCI. Because these species exhibit little spatial aggregation over the area censused, these densities are assumed to be representative of forested areas surrounding BCI, a conservative assumption given that approximately one-third of the area within 10 km of BCI is occupied by Lake Gatun where figs are absent. Breeding unit radii estimate the distances pollen-bearing, female fig wasps routinely disperse in search of receptive host trees. Although actual breeding populations of figs may deviate substantially from the assumed circular distribution, alternative structures (elliptical,for example)increase estimated wasp dispersal distances.

\begin{itemize}
\item Levin’s paternity pool concept.  Levin, D. A. The paternity pools of plants. Am. Nat. 132, 309–317 (1988).
\end{itemize}

[Burnham, K.P. \& Overton, W.S. Robust estimation of population size when capture probabilities vary among animals. Ecology 60, 927–936 (1979).]

\paragraph*{Genetic ‘neighborhood’}

For continuously distributed populations, Wright (1946) introduced the concept of a genetic ‘neighborhood’ that describes the local area within which most matings occur. In a two-dimensional land- scape, Wright’s neighborhood size (NS) is

\$NS = 4pisigma\textasciicircum{}2 D\$

where D is density (number of individuals per unit area) and \$sigma\$ (mean squared distance along one axis between birthplaces of parents and their offspring) is a measure of dispersal. \emph{NS} can be thought of as the number of reproducing individuals in a circle of radius \$2sigma\$. Assuming that dispersal is Gaussian, a circle of this size would include \textasciitilde{}87 \% of the potential parents of individuals at the center (Wright, 1946).

If the dispersal of an individual between place of birth and breeding site is essentially random, it resembles a "drunkards walk". It has the same distribution as passive diffusion, a two-dimensional normal distribution.

If this is true, dispersal distance can simply measured as the standard deviation, \$sigma\$, of the dispersal distribution.  A population "neighbourhood" can be defined approximately as a group of individuals who come from an area \$2sigma\$ wide.

[Strictly, \$sigma\$ is a valid measure of dispersal only if dispersal is exactly normally distributed. Many field studies have shown that dispersal is actually leptokurtic, i.e. most offspring breed very close to their parents, but some breed an enormous distance away. In practice, it doesn't much matter if dispersal is non-normal, provided it is not too extreme.]

%%%%%%%%%%%
https://www.nceas.ucsb.edu/nceas-web/projects/2057/nceas-paper3/data/GFBoxB.html
%%%%%%%%%%%
Continuous populations can be modeled with the concepts of isolation by distance and neighborhood size\citep{Wright:1943aa,Wright:1946aa}. The former refers to the case that limited gene dispersal in continuous populations produces demes that are panmictic internally, but are isolated to some extent, from adjacent demes. Each group of reproducing individuals is the neighborhood, defined as the population of a region in a continuum, from which the parents of individuals born near the center may be treated as if drawn at random \citep{Wright:1969mb}. If distances between parents and offspring follow a normal distribution, the effective population size of the neighborhood will be:

$N_b= 4p \sigma^{2d}$

where $d$ is the density of adults per unit area and $\sigma$ is the variance in distance between birth and breeding sites. This is the basic model of ‘Isolation by Distance’ proposed by Wright  (1943, 1946). Under this type of model, migration (gene flow) is given by the variance in dispersal, and not by the proportion of the population that is composed of migrants (denoted $m$), as is the case with island models (Slatkin 1989).

For plants, gene flow may be accomplished by both seeds and pollen, so the variance may be decomposed to account for different patterns of seed and pollen dispersal, and to take into account the mating system (outcrossing rate, $t$). Thus, neighborhood size can be defined with the following equation (Crawford 1984) :

$N_b = 4p (\sigma^2_s + \frac{t \sigma^2_p}{2}) d (1 + t)$

where $\sigma^2_s$ is the variance in seed dispersal, $\sigma^2_s$ is the variance in pollen dispersal and $d$ is the density of potential parents.

Neighborhood size in plants can be estimated by marking pollen and seeds with fluorescent dyes, tags, or  marks. However, these methods do not measure effective pollen or seed movement, but they may be combined with genetic analysis to do so.  Individuals with a unique allele in a stand can provide valuable insight on seed movement \citep{Eguiarte:1993aa}.

\end{comment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2016-06-17 Manuscript/MS- What is LDD.tex
Minor edits to Table and Figure legends.

@@
 %% Pedro - 07 April 2016
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%----------------------------------------------------------------------------
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Headers
-
 \documentclass[a4paper, 12pt]{article}
@@
 
+%\section*{References}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% References
+%%% Unnumbered Literature Cited
+\bibliographystyle{bes}        %Compile with bes.bst style file
+\bibliography{MS_BES}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURES
+
+\newpage
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TABLES
@@
 \captionsetup{width=20cm}%.75\textwidth}
-\caption{Types of dispersal as a function of population area limits and genetic neighborhood limits.}
+\caption{Types of dispersal as a function of population area limits and genetic neighborhood limits. See Fig. 2 for a graphical representation of the four scenarios.}
 \vspace{0.5cm}
@@
 %----------------------------------------------------------------------------
-
-\section*{References}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% References
-%%% Unnumbered Literature Cited
-\bibliographystyle{bes}        %Compile with bes.bst style file
-\bibliography{MS_BES}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURES
 
@@
 Schematic representation of different types of long-distance dispersal events in relation to the geographical limits of local populations (dashed lines) and the genetic neighborhood size (grey area) of specific individual plants.
-Dispersal events (arrows) can be classified depending on their actual incidence on propagule movement outside these reference areas (Table 1).\\
+Dispersal events (arrows) can be classified depending on their actual incidence on propagule movement outside these spatially-explicit reference areas (Table 1).\\
 
@@
 
-\textbf{Figure 4}. Differential contributions of functional groups of frugivores to the seed dispersal of \textit{Prunus mahaleb}.\\ 
+\textbf{Figure 4}.
+Differential contributions of functional groups of frugivores to the short- and long-distance seed dispersal events for \textit{Prunus mahaleb} according to the classification in Table 1.\\
 \newpage

### 2016-06-17 Manuscript/MS- What is LDD.tex
Minor edits; added bib references.

@@
 \usepackage{parskip}
+\linespread{1.25}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title page
@@
 These two movements do not necessarily concur: a propagule may move over a very long distance yet still be disseminated within the reach of the neighborhood where parental individuals mate.
-Within a demogenetic framework it is easy to envision a combination of situations concerning the spatial scale of the dispersal processes (Table 1) and unambiguously define different types of LDD events.
+Within a demo-genetic framework it is easy to envision a combination of situations concerning the spatial scale of the dispersal processes (Table 1) and unambiguously define different types of LDD events.
 The idea that dispersal occurs in reference to these two spatial reference frames, i.e., the population or stand and the genetic neighborhood area, is motivated by the fact that dispersal entails the movement of both an individual propagule (i.e., a pollen grain or a seed) and a distinct set of genes (i.e., the male genotype in case of pollen, or a seed genotype).
@@
 If we classify dispersal events according to these two spatial frameworks (Table 1) we end up with four distinct types of events depending on whether or not dispersed propagules are disseminated within these reference areas.
-Setting the limits of a population can be problematic yet we can identify with relative ease the geographical limits of plant stands, patches, habitat spots or other types of habitat or microhabitat discontinuities \citep[see][for further discussion of boundaries for dispersal]{Kinlan:2005fb}.
+Setting the limits of a population can be problematic yet we can identify with relative ease the geographical limits of plant stands, patches, habitat spots or other types of habitat or microhabitat discontinuities that determine landmark boundaries of biological significance \citep[see][for further discussion of boundaries for dispersal]{Kinlan:2005fb}.
 These "frontiers" set biological limits to what a LDD event is in relation to the geographic limits of the source population.
 Most plants are distributed as clumped patches, discrete stands, or relatively isolated populations, so we may distinguish between short-distance and long-distance dispersal events that end up with dissemination within or beyond, respectively, the stand or population geographic boundaries (Table 1, $SSD_{loc}$ or $LDD_{loc}$) (Figure 2).
+
+A second consideration in terms of spatial boundaries that has effects on dispersal patterns is the genetic neighborhood area.
+The neighborhood area $N_b$ is the spatial extent including a subset of panmictic individuals within a population.
+Thus, the $N_b$ area can be equal to the whole extent of the population whenever the population is unstructured and there is evidence for random mating events among all th individuals.
+However, most populations and stands of long-lived trees show highly aggregated and clumped distributions, where relatively long distances may separate groups of individuals within the same population.
+In these cases we might expect $N_b$ area to be substantially smaller than the total population area.
 

### 2016-06-19 Manuscript/SupplMat.tex
Added Suppl Mat file and figure S1.

--- 2016-06-17 Manuscript/MS- What is LDD.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% MS Draft Journal of Ecology Special Issue, Dispersal - 
-%% Pedro - 07 April 2016
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% MS Draft Journal of Ecology Special Issue, Dispersal - SUPPL MATERIAL
+%% Pedro - 07 June 2016
+%----------------------------------------------------------------------------
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Headers
-
 \documentclass[a4paper, 12pt]{article}
@@
 \vspace{2cm}
-What is long-distance dispersal? And a taxonomy of dispersal events}
+What is long-distance dispersal? And a taxonomy of dispersal events \\
+\textbf{Supplementary Material}}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Abstract
-\section*{Abstract}
-\begin{linenumbers}
-
-Dispersal is a key individual-based process influencing many life-history attributes, scaling up to population-level properties (e.g., metapopulation connectivity).
-A persistent challenge in dispersal ecology has been the robust characterization of dispersal functions (kernels), a fundamental tool to predict how dispersal processes respond under global change scenarios.
-Especially the rightmost tail of these functions, i.e.
-the long-distance dispersal (LDD) events, are difficult to characterize empirically and to model in realistic ways.
-But, when is it a LDD event?
-In the specific case of plants, dispersal has three basic components: 1) a distinct (sessile) source, the maternal plant producing the fruits or the paternal tree acting as a source of pollen; 2) a distance component between source and target locations; and 3) a vector actually performing the movement entailing the dispersal event.
-Here we discuss operative definitions of LDD based on their intrinsic properties: 1) events crossing geographic boundaries among stands; and 2) events contributing to effective gene flow and propagule migration.
-LDD is a characteristically extreme event of propagule movement in any plant population, typically occurring with an extremely low probability but potentially reaching an extremely long distance.
-Strict-sense long distance involves movement both outside the stand geographic limits and outside the genetic neighborhood area of individuals.
-Combinations of propagule movements within/outside these two spatial reference frames results in distinct modes of LDD.
-Beyond traditional statistical approaches to characterize distributions, Extreme Value Analysis (EVA) can be used to properly and explicitly evaluate the properties of frequency and extent of LDD events.
-We discuss conditions where global change scenarios truncate dispersal processes, leading to the loss of key dispersal services in natural populations.
-Proper characterization of the LDD events helps to assess, for example, how the ongoing defaunation of large-bodied frugivores pervasively entails the loss of crucial LDD functions.
-
-\newpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Introduction
-\section*{Introduction}
-
-Dispersal is a key individual-based process influencing many life-history attributes and scaling up to population-level properties (e.g., metapopulation connectivity, \citealt{Cousens:2008aa}).
-In the specific case of plants, largely sessile organisms, dispersal has three basic components: 1) a distinct (sessile) source, the maternal plant producing the fruits or the paternal tree acting as a source of pollen; 2) a distance component between source and target locations; and 3) a vector actually performing the movement entailing the dispersal event.
-While realized dispersal also depends upon stages subsequent to dissemination (e.g., successful germination and seedling establishment) \cite{Schupp:1995}, the three previous components fully characterize the dispersal process per se.
-Therefore, plant movement differs in important natural history details from animal dispersal, yet both can be assessed within a common conceptual framework \citep[e.g., ][]{Nathan:2006aa}.
-Characteristically, animal-assisted plant dispersal has three distinct, highly integrated, components missing in the process of animal dispersal: the properties of the source (parental) plant, that mediate in the foraging of the animal vector (pollinator or frugivore), the intrinsic properties of the propagule, and the functional characteristics of the animal vector who performs the movement \citep{Nathan:2008fx}.
-
-The movement of pollen and seeds by animals and its consequences have intrigued population geneticists and field ecologists since the infancy of both research disciplines.
-Each has generated an impressive body of theoretical and empirical research through the past decades, yet advances have long been co-existing in ‘parallel worlds’ and the great synergistic potential of population genetics and demography for the study of plant dispersal by animals remains little explored.
-Knowledge gaps still having the imprint of this conceptual disconnection include the idea of long distance dispersal, and the paradoxes of forest fragmentation effects on genetic diversity \citep{Kramer:2008kg}, survival and persistence of relict tree species \citep{Hampe:2011bv}, rapid post-Pleistocene recolonization of vast continental areas in response to climate modification \citep{Clark:1998aa,Clark:1998vi}, among other persisting issues.
-This conceptual isolation has been exacerbated by technical difficulties for the robust characterization of dispersal events, especially those involving movement over long-distances (long-distance dispersal, LDD).
-Some progress has recently been made through the fast-paced implementation of molecular tools in ecological research labs and the availability of cutting-edge technology for biotelemetry applications .
-But much of the population geneticist and ecologist communities remains unaware of the state of the art in each other and likely under-appreciates their potential to validate and enrich dispersal studies \citep{Jones:2008il}.
-In particular, LDD events remain difficult to assess, both technically- with serious methodological problems for its reliable estimation- and conceptually.
-Our aim here is to review the LDD concept with a specific emphasis on dispersal of plant propagules (seeds and pollen), providing an extended definition that might be helpful in the robust quantification of LDD events.
-
-Two main conceptual approaches have been used to assess dispersal (Fig.
-1).
-The “forward” approach attempts to track the dispersal events away from the known sources, e.g., by tracking the movement patterns of frugivores as they leave fruiting plants after feeding (Fig.
-1A).
-This is the main approach used in the movement ecology framework \citep{Nathan:2008fx}, with extensive application to animal movement based on the use of advanced biotelemetry.
-The “backward” approach attempts to reconstruct the most likely source of a dispersed propagule by inferring the sources given the propagule delivery pattern, the fecundity of potential sources, and the dispersal function (Fig.
-1B), i.e., by using an inverse modeling approach.
-The main technical challenge in Fig.
-1A is to sample enough dispersal events away from the source to be able to fully characterize the tail (LDD events) of the dispersal function.
-In Fig.
-1B, the main challenge is to have a robust sampling scheme with propagule collectors (e.g., seed traps) and a good characterization of the potential sources to derive robust estimates of the actual sources.
-Both approaches are limited logistically by the difficulties to sample the vast areas required to assess LDD events from the focal source population.
-
-No explicit definition of what constitutes an LDD event exists.
-Previous approaches \citep[e.g., ][]{Nathan:2006aa,schurr2009long} include both absolute and proportional definitions to characterize LDD events.
-This means providing information about the absolute distances moved by a given percentile of the events and/or providing data on the proportion of events exceeding a given distance threshold \citep{Nathan:2008is}.
-The exact proportional or absolute thresholds selected remain arbitrary, as no reference spatial frame is provided within the definition of LDD.
-This leaves the consideration of LDD as an extreme form of context-dependent phenomenon, strongly dependent upon the scale of the biological process studied \citep{Kinlan:2005fb}.
-For example, \cite{Kinlan:2005fb} used a spatial reference frame to characterize LDD events of marine organisms, where sedentary adults and larvae differ enormously in the spatial scales of their dispersal \citep{DAloia:2013fc}.
-Therefore, any measure of extent and reach of LDD events requires reference to an explicit spatial frame or "local" scale \citep{Kinlan:2005fb}.
-
-We aim at providing a general framework for the quantitative analysis of LDD events so that estimates of its frequency and extent could be comparable across different study systems.
-We argue that both demographic and genetic elements are needed for this framework, most likely requiring a combination of field-based movement data and genetic analyses.
-These elements can be overlaid on previous definitions based on absolute and proportional characterizations of LDD.
-We start with a definition of LDD events within a spatially-explicit mechanistic framework allowing an unambiguous meaning for setting long-distance thresholds.
-We then use a case study to assess differential contributions of animal frugivores performing LDD.
-
-Long-distance dispersal is currently one of the most debated topics in dispersal ecology; it defines the connectedness within the network of local populations and the possibilities for range expansion and successful colonization events.
-We propose a first demogenetically-based, operational definition of what a long-distance dispersal event actually is, and review existing empirical literature on distance thresholds from population and genetic perspectives.
-We also show how molecular tools have been used to identify the respective contributions of different animal species to the LDD portion of dispersal kernels of pollen and seeds by setting empirically-derived distance thresholds.
-Finally, we highlight potential applications of molecular markers beyond the quantification of just the dispersal distances that prevails in current studies, e.g., experimental approaches to assess dispersal limitation and Janzen-Connell effects.
-
-\subsection*{LDD within a demo-genetic perspective: a taxonomy of dispersal events}
-
-Here we propose an explicit definition of LDD and what constitutes a LDD event.
-Previous definitions of dispersal patterns emphasized only their distance components and characterized LDD events basically in terms of geographic distance between a dispersed propagule (or an established early seedling) and its most likely maternal or paternal (in case of pollen) source.
-Absolute and proportional definitions for the LDD events have been proposed depending on arbitrary thresholds of either the distance beyond which a dispersal event is LDD or the proportion of events occurring beyond a specific distance \citep{Nathan:2005jc, Nathan:2008is}.
-Thus, two key biological aspects of LDD events involve the transport of propagules outside a reference area: moving away from the source stand or population, and moving away from the area where relatives stand \citep{Kinlan:2005fb}.
-These two movements do not necessarily concur: a propagule may move over a very long distance yet still be disseminated within the reach of the neighborhood where parental individuals mate.
-Within a demogenetic framework it is easy to envision a combination of situations concerning the spatial scale of the dispersal processes (Table 1) and unambiguously define different types of LDD events.
-The idea that dispersal occurs in reference to these two spatial reference frames, i.e., the population or stand and the genetic neighborhood area, is motivated by the fact that dispersal entails the movement of both an individual propagule (i.e., a pollen grain or a seed) and a distinct set of genes (i.e., the male genotype in case of pollen, or a seed genotype).
-Thus, dispersal entails simultaneous demographic and genetic effects through recruitment of new individuals in the population and through contributions to gene flow \citep{Harper:1977aa}.
-
-Two important components of plant dispersal ecology concern the movement of propagules away from the source population, a type of dispersal relevant to colonization ability and range expansion, and the movement away from the location of close relatives, i.e., a movement away from the genetic neighborhood.
-If we classify dispersal events according to these two spatial frameworks (Table 1) we end up with four distinct types of events depending on whether or not dispersed propagules are disseminated within these reference areas.
-Setting the limits of a population can be problematic yet we can identify with relative ease the geographical limits of plant stands, patches, habitat spots or other types of habitat or microhabitat discontinuities \citep[see][for further discussion of boundaries for dispersal]{Kinlan:2005fb}.
-These "frontiers" set biological limits to what a LDD event is in relation to the geographic limits of the source population.
-Most plants are distributed as clumped patches, discrete stands, or relatively isolated populations, so we may distinguish between short-distance and long-distance dispersal events that end up with dissemination within or beyond, respectively, the stand or population geographic boundaries (Table 1, $SSD_{loc}$ or $LDD_{loc}$) (Figure 2).
-
-\subsection*{Individual and Population Neighborhoods as Reference}
-
-Continuous populations can be modeled with the concepts of isolation by distance and neighborhood size\citep{Wright:1943aa,Wright:1946aa}.
-The former refers to the case that limited gene dispersal in continuous populations produces demes that are panmictic internally, but are isolated to some extent, from adjacent demes.
-Each group of reproducing individuals is the neighborhood, defined as the population of a region in a continuum, from which the parents of individuals born near the center may be treated as if drawn at random \citep{Wright:1969mb}.
-The importance and influence of the dispersal process in determining the size of the neighborhood is given by this equation, which shows how the spatial dispersion (pattern of spatial distribution) of the population influences $N_e$.
-This influence on the effective size is given by:
-
-\begin{equation}
-					N_e= 4 \pi \sigma^2 \delta
-\end{equation}
-
-where $\sigma^2$ is the variance of the dispersal distance and $\delta$ is the density of individuals.
-This formulation is often called the neighborhood size and assumes a normal distribution of distances  between parents and offspring (out in a perfect circular shape from the source).
-Thus, changes in the variance of dispersal size can affect $N_e$ (highly clumped populations will have reduced $N_e$).
-
-\begin{equation}
-					N_b= 4p \sigma^{2d}
-\end{equation}
-
-where $d$ is the density of adults per unit area and $\sigma$ is the variance in distance between birth and breeding sites.
-This is the basic model of ‘Isolation by Distance’ proposed by   \citet{Wright:1943aa,Wright:1946aa}.
-Under this type of model, migration (gene flow) is given by the variance in dispersal, and not by the proportion of the population that is composed of migrants (denoted $m$), as is the case with island models \citep{Slatkin:1985qb}.
-
-For plants, gene flow may be accomplished by both seeds and pollen, so the variance may be decomposed to account for different patterns of seed and pollen dispersal, and to take into account the mating system (outcrossing rate, $t$).
-Thus, neighborhood size can be defined with the following equation (Crawford 1984) :
-
-\begin{equation}
-  					N_b = 4p (\sigma^2_s + \frac{t \sigma^2_p}{2}) d (1 + t)
-\end{equation}
-
-where $\sigma^2_s$ is the variance in seed dispersal, $\sigma^2_s$ is the variance in pollen dispersal and $d$ is the density of potential parents.
-
-Neighborhood size in plants can be estimated by marking pollen and seeds with fluorescent dyes, tags, or  marks.
-However, these methods do not measure effective pollen or seed movement, but they may be combined with genetic analysis to do so.
-Individuals with a unique allele in a stand can provide valuable insight on seed movement \citep{Eguiarte:1993aa}.
-
-\paragraph*{Empirical Studies of Seed Dispersal}
-
-\subsection*{Long-Distance Dispersal: the ecology of extreme events}
-
-Long-distance dispersal (LDD) is a major component of the population dynamics, genetic structure, and biogeographic history of plant species.
-It determines the colonization ability of new habitats and the possibilities for fragmented populations to sustain a cohesive metapopulation by immigration-emigration dynamics that rely on LDD events.
-Yet our current understanding of the extent, frequency, and consequences of LDD is very limited.
-On one hand, theoretical models fail to predict accurately the behavior of the tail of the dispersal functions, and thus fail to predict very basic properties of LDD.
-On the other hand, we still have very limited documentation of actual LDD events in natural populations and we still see LDD as a sporadic, rarely far-reaching process still marked with the stamp of natural history curiosity.
-
-As defined in our framework (Table 1), LDD, and in particular $LDD_{ss}$ events are a particular case of extreme events \citep{Garcia:2017aa}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Discussion
-%\section*{Discussion}
-
-\subsection*{Challenges and Promising Avenues for Research}
-
-\emph{Acknowledgements}.
-I am indebted to Cristina García, José A.
-Godoy, Manolo Carrión, Juan Luis García-Castaño, Jesús Rodríguez and, especially, Juan Miguel Arroyo for generous help with field and laboratory work and making possible this study.
-I appreciate the help and advice of Cristina García and Etienne Klein during the final stages of the manuscript.
-The study was supported by a Junta de Andalucía Excellence Grant (RNM-5731), as well as a Severo Ochoa Excellence Award from the Ministerio de Economía y Competitividad (SEV-2012-0262).
-The Agencia de Medio Ambiente, Junta de Andalucía, provided generous facilities that made possible this study in the Andalusian natural parks (Sierra de Cazorla, Alcornocales) and authorized my work there.
-
-\end{linenumbers}
-
-\newpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TABLES
-%-------------------------------------------------------------------- Table 1
-\pagestyle{empty}
-\begin{landscape}
-	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%	\oddsidemargin  0.0in
-%	\evensidemargin 0.0in
-%	\textwidth      8.5in
-	\headheight     3cm
-%	\topmargin      0.0in
-%	\textheight=8.0in
-	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}
-\captionsetup{width=20cm}%.75\textwidth}
-\caption{Types of dispersal as a function of population area limits and genetic neighborhood limits.}
-\vspace{0.5cm}
-  \begin{tabular}{lll}
-    \hline
-\\&Population geographic limit    &  \\\\
-Genetic neighborhood limit     &Within &Outside \\\\
-    \hline
-\\Within    &  Local, short-distance \linebreak dispersal, $SDD_{loc}$  
-          &  Within neighborhood, \linebreak long-distance dispersal, $LDD_{neigh}$ \\\\
-Outside   &     Local, long-distance \linebreak dispersal, $LDD_{loc}$    
-          &  Strict sense \linebreak long-distance dispersal, $LDD_{ss}$\\\\ 
-    \hline
-  \end{tabular}
-\end{table}
-\end{landscape}
-
-%----------------------------------------------------------------------------
-
-\section*{References}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% References
-%%% Unnumbered Literature Cited
-\bibliographystyle{bes}        %Compile with bes.bst style file
-\bibliography{MS_BES}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURES
-
-\newpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LANDSCAPE FIGURES
-%%%% Figure LEGENDS
-%%%% This is for sideways figures (landscape), with legend in a separate page.
-\section*{Figures}
-\begin{linenumbers}
-
-\textbf{Figure 1}.
-The two approaches used in analyses of dispersal processes in plants.
-A, the “forward” approach attempts to track the dispersal events away from the known sources, e.g., by tracking the movement patterns of frugivores as they leave fruiting plants after feeding.
-B, the “backward” approach  attempts to reconstruct the most likely source of a dispersed propagule by inferring the sources given the propagule delivery pattern, the fecundity of potential sources, and the dispersal function.
-The main technical challenge in A is to sample enough dispersal events away from the source to be able to fully characterize the tail (LDD events) of the dispersal function.
-In B, the main challenge is to have a robust sampling scheme with propagule collectors (e.g., seed traps) and a good characterization of the potential sources to derive robust estimates of the actual sources.\\
-
-\textbf{Figure 2}.
-Schematic representation of different types of long-distance dispersal events in relation to the geographical limits of local populations (dashed lines) and the genetic neighborhood size (grey area) of specific individual plants.
-Dispersal events (arrows) can be classified depending on their actual incidence on propagule movement outside these reference areas (Table 1).\\
-
-\textbf{Figure 3}. LDD events depend on a combination of disperser characteristics and landscape features. While different ***.\\
-
-\textbf{Figure 4}. Differential contributions of functional groups of frugivores to the seed dispersal of \textit{Prunus mahaleb}.\\ 
-\newpage
-
-%------------------------------------------------------------------- Figure 1
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Supplementary Material
+%------------------------------------------------------------------ Figure S1
 \begin{figure}[htbp]
-\centerline{\includegraphics[height=25cm]{Fig1.pdf}}
+\centerline{\includegraphics[height=21cm]{FigS1.pdf}}
 %
-%\caption{***}
-\end{figure}
-
-\newpage 
-%------------------------------------------------------------------- Figure 2
-\begin{figure}[htbp]
-\centerline{\includegraphics[height=25cm]{Fig2.pdf}}
-%
-%\caption{***}
-\end{figure}
-
-\newpage 
-%------------------------------------------------------------------- Figure 3
-\begin{figure}[htbp]
-\centerline{\includegraphics[height=26cm]{Fig3.pdf}}
-%
-%\caption{***}
-\end{figure}
-%----------------------------------------------------------------------------
-
-\newpage 
-%------------------------------------------------------------------- Figure 4
-\begin{figure}[htbp]
-\centerline{\includegraphics[height=25cm]{Fig4.pdf}}
-%
-%\caption{***}
+\caption{Dispersal events for pollen (left) and seeds (right) traced for \textit{Prunus mahaleb} trees (white dots).
+All the adult, reproductive, trees in the population are mapped.
+Lines indicate mating events of pollen dispersal among trees (left) or seed dissemination events from source fruiting trees to seed traps (suqares; right).
+Line thickness is proportional to the number of events recorded.}
 \end{figure}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Online Support Material and data accessiblity}
-
-This review does not use new raw data, but includes some re-analyses of previously published material.
-All the original data supporting the paper, R code, supplementary figures, and summaries of analytical protocols is available at the author's GitHub repository (\texttt{https://github.com/pedroj/MS\_LDD}), with DOI: \texttt{\#/zenodo.\#}.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\end{linenumbers}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% MS NOTES
-\begin{comment}
-NOTES
-Breeding unit size.
-Donor number is related to the size of the breeding unit (N individuals) as d= Np, where d is the number of different pollen parent genotypes represented in a maternal tree’s fruit crop and p is the expected proportion of trees in staminate phase at any given time.
-Assuming, conservatively, that individual trees reproduce asynchronously twice per year 6 and that staminate and pistillate flowering phases each last seven days 20, p is calculated to be 0.0712.
-Although these estimates could be effected by mosaicism 21 (the fusion of genetically distinct individuals), its frequency in the species examined here is low.
-
-Breeding unit area and radii.
-These parameters were calculated from paternity-analysis-based estimates of breeding unit size (Nˆdˆ; Table2) and the censused densities of adult, reproductively mature trees over 15 km2 BCI.
-Based on the long-term censuses of C.
-Handley and E.
-Kalko, 6, 108 and 20 adults of F.
-dugandii, F.
-obtusifolia and F.
-popenoei, respectively, are known to occur on BCI.
-Because these species exhibit little spatial aggregation over the area censused, these densities are assumed to be representative of forested areas surrounding BCI, a conservative assumption given that approximately one-third of the area within 10 km of BCI is occupied by Lake Gatun where figs are absent.
-Breeding unit radii estimate the distances pollen-bearing, female fig wasps routinely disperse in search of receptive host trees.
-Although actual breeding populations of figs may deviate substantially from the assumed circular distribution, alternative structures (elliptical,for example)increase estimated wasp dispersal distances.
-
-\begin{itemize}
-\item Levin’s paternity pool concept.  Levin, D. A. The paternity pools of plants. Am. Nat. 132, 309–317 (1988).
-\end{itemize}
-
-[Burnham, K.P. \& Overton, W.S. Robust estimation of population size when capture probabilities vary among animals. Ecology 60, 927–936 (1979).]
-
-\paragraph*{Genetic ‘neighborhood’}
-
-For continuously distributed populations, Wright (1946) introduced the concept of a genetic ‘neighborhood’ that describes the local area within which most matings occur.
-In a two-dimensional land- scape, Wright’s neighborhood size (NS) is
-
-\$NS = 4pisigma\textasciicircum{}2 D\$
-
-where D is density (number of individuals per unit area) and \$sigma\$ (mean squared distance along one axis between birthplaces of parents and their offspring) is a measure of dispersal.
-\emph{NS} can be thought of as the number of reproducing individuals in a circle of radius \$2sigma\$.
-Assuming that dispersal is Gaussian, a circle of this size would include \textasciitilde{}87 \% of the potential parents of individuals at the center (Wright, 1946).
-
-If the dispersal of an individual between place of birth and breeding site is essentially random, it resembles a "drunkards walk".
-It has the same distribution as passive diffusion, a two-dimensional normal distribution.
-
-If this is true, dispersal distance can simply measured as the standard deviation, \$sigma\$, of the dispersal distribution.
-A population "neighbourhood" can be defined approximately as a group of individuals who come from an area \$2sigma\$ wide.
-
-[Strictly, \$sigma\$ is a valid measure of dispersal only if dispersal is exactly normally distributed.
-Many field studies have shown that dispersal is actually leptokurtic, i.e.
-most offspring breed very close to their parents, but some breed an enormous distance away.
-In practice, it doesn't much matter if dispersal is non-normal, provided it is not too extreme.]
-
-%%%%%%%%%%%
-https://www.nceas.ucsb.edu/nceas-web/projects/2057/nceas-paper3/data/GFBoxB.html
-%%%%%%%%%%%
-Continuous populations can be modeled with the concepts of isolation by distance and neighborhood size\citep{Wright:1943aa,Wright:1946aa}.
-The former refers to the case that limited gene dispersal in continuous populations produces demes that are panmictic internally, but are isolated to some extent, from adjacent demes.
-Each group of reproducing individuals is the neighborhood, defined as the population of a region in a continuum, from which the parents of individuals born near the center may be treated as if drawn at random \citep{Wright:1969mb}.
-If distances between parents and offspring follow a normal distribution, the effective population size of the neighborhood will be:
-
-$N_b= 4p \sigma^{2d}$
-
-where $d$ is the density of adults per unit area and $\sigma$ is the variance in distance between birth and breeding sites.
-This is the basic model of ‘Isolation by Distance’ proposed by Wright  (1943, 1946).
-Under this type of model, migration (gene flow) is given by the variance in dispersal, and not by the proportion of the population that is composed of migrants (denoted $m$), as is the case with island models (Slatkin 1989).
-
-For plants, gene flow may be accomplished by both seeds and pollen, so the variance may be decomposed to account for different patterns of seed and pollen dispersal, and to take into account the mating system (outcrossing rate, $t$).
-Thus, neighborhood size can be defined with the following equation (Crawford 1984) :
-
-$N_b = 4p (\sigma^2_s + \frac{t \sigma^2_p}{2}) d (1 + t)$
-
-where $\sigma^2_s$ is the variance in seed dispersal, $\sigma^2_s$ is the variance in pollen dispersal and $d$ is the density of potential parents.
-
-Neighborhood size in plants can be estimated by marking pollen and seeds with fluorescent dyes, tags, or  marks.
-However, these methods do not measure effective pollen or seed movement, but they may be combined with genetic analysis to do so.
-Individuals with a unique allele in a stand can provide valuable insight on seed movement \citep{Eguiarte:1993aa}.
-
-\end{comment}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-06-20 Manuscript/MS- What is LDD.tex
Replaced figures 3 and 4.
Added legends and minor edits.

@@
 Here we discuss operative definitions of LDD based on their intrinsic properties: 1) events crossing geographic boundaries among stands; and 2) events contributing to effective gene flow and propagule migration.
-LDD is a characteristically extreme event of propagule movement in any plant population, typically occurring with an extremely low probability but potentially reaching an extremely long distance.
-Strict-sense long distance involves movement both outside the stand geographic limits and outside the genetic neighborhood area of individuals.
-Combinations of propagule movements within/outside these two spatial reference frames results in distinct modes of LDD.
+Strict-sense long distance dispersal involves movement both outside the stand geographic limits and outside the genetic neighborhood area of individuals.
+Combinations of propagule movements within/outside these two spatial reference frames results in four distinct modes of LDD.
 Beyond traditional statistical approaches to characterize distributions, Extreme Value Analysis (EVA) can be used to properly and explicitly evaluate the properties of frequency and extent of LDD events.
@@
 This conceptual isolation has been exacerbated by technical difficulties for the robust characterization of dispersal events, especially those involving movement over long-distances (long-distance dispersal, LDD).
-Some progress has recently been made through the fast-paced implementation of molecular tools in ecological research labs and the availability of cutting-edge technology for biotelemetry applications .
+LDD is a characteristically extreme event of propagule movement in any plant population, typically occurring with an extremely low probability but potentially reaching an extremely long distance.
+Some progress has recently been made through the fast-paced implementation of molecular tools in ecological research labs and the availability of cutting-edge technology for biotelemetry applications [REF].
 But much of the population geneticist and ecologist communities remains unaware of the state of the art in each other and likely under-appreciates their potential to validate and enrich dispersal studies \citep{Jones:2008il}.
@@
 
-A second consideration in terms of spatial boundaries that has effects on dispersal patterns is the genetic neighborhood area.
-The neighborhood area $N_b$ is the spatial extent including a subset of panmictic individuals within a population.
-Thus, the $N_b$ area can be equal to the whole extent of the population whenever the population is unstructured and there is evidence for random mating events among all th individuals.
-However, most populations and stands of long-lived trees show highly aggregated and clumped distributions, where relatively long distances may separate groups of individuals within the same population.
+A second consideration in terms of spatial boundaries, with effects on dispersal patterns, is the genetic neighborhood area $N_b$, i.e., the spatial extent including a subset of panmictic individuals within a population \citep{Wright:1943aa,Wright:1946aa}.
+Thus, the $N_b$ area can be equal to the whole extent of the population whenever the population is unstructured and there is evidence for random mating events among all the individuals.
+However, most populations and stands of long-lived trees show highly aggregated and clumped distributions \citep{Seidler:2006hx}, where relatively long distances may separate groups of individuals within the same population.
 In these cases we might expect $N_b$ area to be substantially smaller than the total population area.
+Therefore, at least four possible scenarios exist with distinct implications in terms of consequences for dispersal (Table 1).
+In the case of dispersal events not extending beyond the geographic limits of the population or reference area actual LDD event may involve dissemination beyond a reduced neighborhood area that is smaller than the geographic extent of the population, originating local long-distance ($LDD_{loc}$) dispersal events (Table 1, Fig.
+2A).
+Actual short-distance dispersal would then involve those situations where the propagule is disseminated within \textit{both} the population limits and the genetic neighborhood boundary ($SDD_{loc}$).
+Along a similar reasoning, dispersal events outside the population limits will not necessarily convey LDD (Table 1, Fig.
+2B): this is expected in cases where the genetic neighborhoods are extensive, going beyond the geographic limits of local populations, as in fig trees \citep{Nason:1998aa}, generating LDD events within the genetic neighborhood ($LDD_{neigh}$).
+Finally, strict-sense LDD events would involve dissemination outside \textit{both} the population limits and the genetic neighborhood boundary ($LDD_{ss}$) (Table 1, Fig.
+2A).
 
@@
 \vspace{0.5cm}
-  \begin{tabular}{lll}
+  \begin{tabular}{lcc}
     \hline
@@
     \hline
-\\Within    &  Local, short-distance \linebreak dispersal, $SDD_{loc}$  
-          &  Within neighborhood, \linebreak long-distance dispersal, $LDD_{neigh}$ \\\\
-Outside   &     Local, long-distance \linebreak dispersal, $LDD_{loc}$    
-          &  Strict sense \linebreak long-distance dispersal, $LDD_{ss}$\\\\ 
+\\Within  &  Local, short-distance dispersal, $SDD_{loc}$  
+          &  Within neighborhood, long-distance dispersal, $LDD_{neigh}$ \\\\
+Outside   &     Local, long-distance dispersal, $LDD_{loc}$    
+          &  Strict sense long-distance dispersal, $LDD_{ss}$\\\\ 
     \hline
@@
 B, the “backward” approach  attempts to reconstruct the most likely source of a dispersed propagule by inferring the sources given the propagule delivery pattern, the fecundity of potential sources, and the dispersal function.
-The main technical challenge in A is to sample enough dispersal events away from the source to be able to fully characterize the tail (LDD events) of the dispersal function.
-In B, the main challenge is to have a robust sampling scheme with propagule collectors (e.g., seed traps) and a good characterization of the potential sources to derive robust estimates of the actual sources.\\
+The main technical challenge in A is to sample enough dispersal events away from the source to be able to fully characterize the tail (long-distance dispersal, LDD, events) of the dispersal function.
+In B, the main challenge is to have a robust sampling scheme with propagule collectors (e.g., seed traps) and a good characterization of the potential sources to derive robust estimates of the actual sources with inverse-modeling techniques.\\
 
 \textbf{Figure 2}.
-Schematic representation of different types of long-distance dispersal events in relation to the geographical limits of local populations (dashed lines) and the genetic neighborhood size (grey area) of specific individual plants.
-Dispersal events (arrows) can be classified depending on their actual incidence on propagule movement outside these spatially-explicit reference areas (Table 1).\\
+Schematic representation of different types of long-distance dispersal events in relation to the geographical limits of local populations (dashed lines) and the genetic neighborhood area $N_b$ (grey area) of specific individual plants (squares).
+Dispersal events (arrows) can be classified depending on their actual incidence on propagule movement outside these spatially-explicit reference areas (Table 1).
+Strict-sense long-distance dispersal events ($LDD_{ss}$) just include the LDD events that disseminate propagules out of \textit{both} the population and genetic neighborhood boundaries.
+A, the neighborhood area is included within the geographic limits of the population; B, the neighborhood area is much larger than the geographic limits of the population.\\
 
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[height=25cm]{Fig4.pdf}}
+\centerline{\includegraphics[height=27cm]{Fig4.pdf}}
 %
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% MS NOTES
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% MS NOTES
 \begin{comment}
 NOTES
+% -------------- Crawford 21986 Heredity ------------------------------------
+Sewal Wright's neighborhood model indicates that the area containing a panmictic unit within a continuous and uniform array of organisms can be 
+estimated by $4\pir^2$ where a-2 is the parent-offspring dispersal variance measured around a zero  mean and relative to a single reference axis passing through the population.
+
+% ---------------------------------------------------------------------------
 Breeding unit size.
@@
 
-%%%%%%%%%%%
+% --------------------------------------------------------------------------
 https://www.nceas.ucsb.edu/nceas-web/projects/2057/nceas-paper3/data/GFBoxB.html
-%%%%%%%%%%%
+
 Continuous populations can be modeled with the concepts of isolation by distance and neighborhood size\citep{Wright:1943aa,Wright:1946aa}.
@@
 
+%------ Levin & Kerster 1971 ------------------------------------------------
+It has been shown that neighborhood size contracts under autogamy,as it does under dioecy or heteromorphic incompatibility if these morph ratios are disproportionate.
+Neighborhood size may expand under heteromorphic incompatibility or dioecy through enhanced pollen carry-over if the sex ratio is 1:1 and if the plants are insect-pollinated.
+
 \end{comment}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%----------------------------------------------------- SUPPLEMENTARY MATERIAL
+%\include{SupplMat}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-06-20 Manuscript/SupplMat.tex
Added legends and minor edits.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Headers
-\documentclass[a4paper, 12pt]{article}
+\documentclass[a4paper, 12pt]{article}}
 \usepackage{graphicx}
@@
 \usepackage{rotating} 
+\usepackage{standalone}                                                                                    
 \newcommand{\tab}{\hspace{5mm}}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Supplementary Material
+%------------------------------------------------------------------- Table S1
+\begin{landscape}
+\begin{table}
+  \caption{Summary of neighborhood area sizes and estimated neighborhood radius for tree species with different combinations of dispersal modes. Data from Nason et al and present study. }
+    \begin{tabular}{lllccc}
+     Species                & Pollinator  & Seed disperser & Density ($ha^{-1}$) & Breeding unit ($km^2$) & Radius (km) \\
+    \textit{Ficus dugandii}          & Fig wasp    & Vertebrates    & 0.004          & 631.7               & 14.2        \\
+    \textit{Ficus obtusifolia}       & Fig wasp    & Vertebrates    & 0.072          & 105.9               & 5.8         \\
+    \textit{Prunus mahaleb}          & Bees, flies & Vertebrates    & 0.003          & 0.87                & 0.042       \\
+    \textit{Frangula alnus}          & Bees, flies & Vertebrates    & 0.0004         & 0.45                & 0.013       \\
+    \textit{Astrocaryum mexicanum}   & Beetle      & Vertebrates    & 1364.0         & 0.011               & 0.06        \\
+    \textit{Calophyllum longifolium} & Bees        & Vertebrates    & 0.28           & 1.241               & 0.629       \\
+    \textit{Platypodium elegans}     & Bees        & Wind           & 0.78           & 0.866               & 0.525       \\
+    \textit{Cedrus atlantica}        & Wind        & Wind           & 61.7           & 0.151               & 0.22        \\
+   \textit{Fraxinus americana}      & Wind        & Wind           & 24.7           & 0.008               & 0.05        \\
+    \textit{Pseudotsuga menziesii}   & Wind        & Wind           & 25.0           & 0.078               & 0.158       \\
+    \end{tabular}
+\end{table}
+\end{landscape}
+\newpage
+
 %------------------------------------------------------------------ Figure S1
@@
 All the adult, reproductive, trees in the population are mapped.
-Lines indicate mating events of pollen dispersal among trees (left) or seed dissemination events from source fruiting trees to seed traps (suqares; right).
+Lines indicate mating events of pollen dispersal among trees (left) or seed dissemination events from source fruiting trees to seed traps (squares; right).
 Line thickness is proportional to the number of events recorded.}

### 2016-06-20 Manuscript/MS- What is LDD.tex
Added Fig 4 and moved fig to Suppl Mat. Edited figure legends.

@@
 
-\textbf{Figure 3}. LDD events depend on a combination of disperser characteristics and landscape features. While different ***.\\
+\textbf{Figure 3}.
+Empirical frequency distributions of seed dispersal events as a function of dispersal distance by animal frugivores consuming \textit{Prunus mahaleb} fruits.
+In red, left (inset), frequencies of within-populations dispersal events inferred from direct assignment based on seed endocarp genotypes and maternal trees genotypes.
+Larger frame, left, contributions of four functional frugivore groups (small birds, medium- and large-sized birds, and mammals) to seed dissemination and proportional contributions (right bar) to dispersal of the inferred immigrant seeds (i.e., those not matching any maternal tree in the study population).\\
 
 \textbf{Figure 4}.
-Differential contributions of functional groups of frugivores to the short- and long-distance seed dispersal events for \textit{Prunus mahaleb} according to the classification in Table 1.\\
+Differential contributions of functional groups of frugivores to the four combinations of \textit{Prunus mahaleb} seed dispersal events outlined in Table 1.
+These result from dissemination within (yellow) and outside (blue) the population geographic limits ($SDD_{loc}$, $LDD_{loc}$, respectively) and/or the genetic neigborhood area limits ($SDD_{neigh}$, $LDD_{ss}$, respectively).\\
 \newpage

### 2016-06-20 Manuscript/MS- What is LDD.tex
Added Table 2 and edited figure legends,

@@
 \usepackage{tabularx} 
+%\usepackage{multirow}
 \usepackage[margin=10pt,font=small,labelfont=bf]{caption}
@@
 
+\newpage
+
+%-------------------------------------------------------------------- Table 2
+\pagestyle{empty}
+\begin{landscape}
+\begin{table}
+\captionsetup{width=21cm}%.75\textwidth}
+\caption{Frequencies of \textit{Prunus mahaleb} seed dispersal events for different frugivore groups according to population area limits and genetic neighborhood limits.
+See Fig.
+2 for a graphical representation of the four scenarios.}
+\vspace{0.5cm}
+    \begin{tabular}{lp{3cm}p{3cm}p{3cm}p{3cm}}
+    \hline \\
+    Frugivore group & Within-population, within-neighborhood $SSD_{loc}$ & Within-population, long-distance $LDD_{loc}$ & Outside-population, within-neighborhood $LDD_{neigh}$ & Strict-sense long-distance $LDD_{ss}$ \\\\ \hline \\
+    Small birds     & ?~           & ?~            & ?~           & ?~  \\
+    \textit{Turdus} & ?~           & ?~            & ?~           & ?~  \\
+
+    Large birds     & ?~           & ?~            & ?~           & ?~  \\
+
+    Mammals         & ?~           & ?~            & ?~           & ?~  \\\\
+\hline
+    \end{tabular}
+\end{table}
+\end{landscape}
+
 %----------------------------------------------------------------------------
 
+\newpage  
 \newpage
@@
 Differential contributions of functional groups of frugivores to the four combinations of \textit{Prunus mahaleb} seed dispersal events outlined in Table 1.
-These result from dissemination within (yellow) and outside (blue) the population geographic limits ($SDD_{loc}$, $LDD_{loc}$, respectively) and/or the genetic neigborhood area limits ($SDD_{neigh}$, $LDD_{ss}$, respectively).\\
+These result from dissemination within (yellow) or outside (blue) the population geographic limits ($SDD_{loc}$, $LDD_{loc}$, respectively) and within or outside the genetic neigborhood area limits ($SDD_{neigh}$, $LDD_{ss}$, respectively).\\
 \newpage

### 2016-06-23 Manuscript/MS- What is LDD.tex
Edits
Minor edits to first section.

@@
 No explicit definition of what constitutes an LDD event exists.
-Previous approaches \citep[e.g., ][]{Nathan:2006aa,schurr2009long} include both absolute and proportional definitions to characterize LDD events.
+Previous approaches \citep[e.g., ][]{Nathan:2006aa,Schurr2009long} include both absolute and proportional definitions to characterize LDD events.
 This means providing information about the absolute distances moved by a given percentile of the events and/or providing data on the proportion of events exceeding a given distance threshold \citep{Nathan:2008is}.
@@
 Thus, dispersal entails simultaneous demographic and genetic effects through recruitment of new individuals in the population and through contributions to gene flow \citep{Harper:1977aa}.
+When considered its combined influence on demography and population genetics, the concept of LDD nicely bridges these two paradigms embedded in the biological definition of population \citep{Waples:2006ev}.
 
-Two important components of plant dispersal ecology concern the movement of propagules away from the source population, a type of dispersal relevant to colonization ability and range expansion, and the movement away from the location of close relatives, i.e., a movement away from the genetic neighborhood.
+Two important components of plant dispersal ecology concern the movement of propagules away from the source population, a type of dispersal relevant to colonization ability and range expansion \citep{Howe:2004}, and the movement away from the location of close relatives, i.e., a movement away from the genetic neighborhood \citep{Hardesty:2006lr,Jones:2008il}.
 If we classify dispersal events according to these two spatial frameworks (Table 1) we end up with four distinct types of events depending on whether or not dispersed propagules are disseminated within these reference areas.
-Setting the limits of a population can be problematic yet we can identify with relative ease the geographical limits of plant stands, patches, habitat spots or other types of habitat or microhabitat discontinuities that determine landmark boundaries of biological significance \citep[see][for further discussion of boundaries for dispersal]{Kinlan:2005fb}.
+Setting the limits of a population can be problematic \citep{Waples:2006ev} yet we can identify with relative ease the geographical limits of plant stands, patches, habitat spots or other types of habitat or microhabitat discontinuities that determine landmark boundaries of biological significance \citep[see][for further discussion of boundaries for dispersal]{Kinlan:2005fb}.
 These "frontiers" set biological limits to what a LDD event is in relation to the geographic limits of the source population.
@@
 Therefore, at least four possible scenarios exist with distinct implications in terms of consequences for dispersal (Table 1).
-In the case of dispersal events not extending beyond the geographic limits of the population or reference area actual LDD event may involve dissemination beyond a reduced neighborhood area that is smaller than the geographic extent of the population, originating local long-distance ($LDD_{loc}$) dispersal events (Table 1, Fig.
+In the case of dispersal events not extending beyond the geographic limits of the population or reference area, actual LDD events may involve dissemination beyond a reduced neighborhood area that is smaller than the geographic extent of the population, originating local long-distance ($LDD_{loc}$) dispersal events (Table 1, Fig.
 2A).
@@
 2A).
+
+While both $SSD_{loc}$ and $LDD_{loc}$ can be crucial for assuring the local persistence of populations, $LDD_{neigh}$ and $LDD_{ss}$ would be extremely important contributors to the structuring of genetic pools, realized gene flow, and maintaining connectivity in metapopulation scenarios.
+We argue that both the demographic and the genetic references are relevant for a proper definition of LDD.
 
@@
 
-\newpage  
 \newpage

### 2016-06-23 Manuscript/MS- What is LDD.tex
Edits to section 2 and 3 and Tab 1 SupplMat.

@@
 Continuous populations can be modeled with the concepts of isolation by distance and neighborhood size\citep{Wright:1943aa,Wright:1946aa}.
-The former refers to the case that limited gene dispersal in continuous populations produces demes that are panmictic internally, but are isolated to some extent, from adjacent demes.
+The former refers to the case that limited gene dispersal in continuous populations produces demes that are panmictic internally, but are isolated to some extent from adjacent demes.
 Each group of reproducing individuals is the neighborhood, defined as the population of a region in a continuum, from which the parents of individuals born near the center may be treated as if drawn at random \citep{Wright:1969mb}.
-The importance and influence of the dispersal process in determining the size of the neighborhood is given by this equation, which shows how the spatial dispersion (pattern of spatial distribution) of the population influences $N_e$.
+The importance and influence of the dispersal process in determining the size of the neighborhood is given by this equation, which shows how the spatial dispersion (pattern of spatial distribution) of the population influences the effective population size.
 This influence on the effective size is given by:
@@
 \begin{equation}
-					N_e= 4 \pi \sigma^2 \delta
+					N_b= 4 \pi \sigma^2 \delta
 \end{equation}
 
-where $\sigma^2$ is the variance of the dispersal distance and $\delta$ is the density of individuals.
+where $\delta$ is the density of adults per unit area and $\sigma$ is the variance in distance between birth and breeding sites.
 This formulation is often called the neighborhood size and assumes a normal distribution of distances between parents and offspring (out in a perfect circular shape from the source).
-Thus, changes in the variance of dispersal size can affect $N_e$ (highly clumped populations will have reduced $N_e$).
-
-\begin{equation}
-					N_b= 4p \sigma^{2d}
-\end{equation}
-
-where $d$ is the density of adults per unit area and $\sigma$ is the variance in distance between birth and breeding sites.
-This is the basic model of ‘Isolation by Distance’ proposed by   \citet{Wright:1943aa,Wright:1946aa}.
+Thus, changes in the variance of dispersal distance can affect $N_e$ (highly clumped populations will have reduced $N_e$).
+This is the basic model of "Isolation by Distance" proposed by   \citet{Wright:1943aa,Wright:1946aa}.
 Under this type of model, migration (gene flow) is given by the variance in dispersal, and not by the proportion of the population that is composed of migrants (denoted $m$), as is the case with island models \citep{Slatkin:1985qb}.
@@
 \begin{equation}
-  					N_b = 4p (\sigma^2_s + \frac{t \sigma^2_p}{2}) d (1 + t)
+  			N_b = 4 \pi (\sigma^2_s + \frac{t \sigma^2_p}{2}) \delta (1 + t)
 \end{equation}
 
-where $\sigma^2_s$ is the variance in seed dispersal, $\sigma^2_s$ is the variance in pollen dispersal and $d$ is the density of potential parents.
+where $\sigma^2_s$ is the variance in seed dispersal, $\sigma^2_s$ is the variance in pollen dispersal and $\delta$ is the density of potential parents.
 
-Neighborhood size in plants can be estimated by marking pollen and seeds with fluorescent dyes, tags, or  marks.
-However, these methods do not measure effective pollen or seed movement, but they may be combined with genetic analysis to do so.
-Individuals with a unique allele in a stand can provide valuable insight on seed movement \citep{Eguiarte:1993aa}.
+Neighborhood size in plants can be estimated by marking pollen and seeds with fluorescent dyes, tags, or stable isotope enrichment \citep{Carlo:2009wa}.
+However, these methods do not measure effective pollen or seed movement, but they may be combined with genetic analysis to assess genetic identity and relatedness with hypervariable DNA markers \citep{Levin:1988fm,Nason:1998aa,Godoy:2001} to achieve reliable estimates of both effective population size and neighborhood area.
 
-\paragraph*{Empirical Studies of Seed Dispersal}
+The extent of neighborhood area in plants can be extremely variable, depending on life-history attributes such as life-span, spacing patterns, mating system, etc.
+Even a limited sample of available information (Table S1) highlights the fact that the size of neighborhood areas can in some cases exceed the geographic limits of local populations \citep{Nason:1998aa}.
+The size of neighborhood areas may encompass at least four orders of magnitude, $10{^-2}-10^2$ km in radius, and include many individuals.
+Therefore, reference to this "genetic/evolutionary" paradigm and reference \citep[sensu][]{Waples:2006ev} may be instrumental to understand the actual role of LDD events in shaping the structuring of genetic pools and contributing to gene dispersal.
+
+Whenever there is a large discrepancy between population area extent and $N_b$ we might expect the frequency of $LDD_{loc}$ and $LDD_{neigh}$ differ enormously.
+For example, relatively small $N_b$ may rise the importance of $LDD_{loc}$ in preserving scenarios of panmixia within a local population, as most distant dispersal events will disseminate seeds outside the neighborhood of maternal plants.
+
+\subsection*{Empirical analysis of contributions to LDD}
+
+Empirical evaluation of differential contributions to the different forms of LDD events outlined in Table 1 requires identification of source trees as well as assignment of the dispersed propagules to specific vectors or functional groups of vectors \citep{Jordano:2007}.
+Recently, DNA-barcoding techniques have been developed and successfully applied to the identification of frugivore species contributing to specific seed dispersal events whose source can be identified with genetic, direct assignment techniques \citep{GonzalezVaro:2014ij}.
+Otherwise, visual identification can reliably assign the genotyped seeds to frugivore species groups based on specific characteristics of scats and regurgitations \citep{Jordano:2007a}.
 
@@
 %------------------------------------------------------------------- Figure 1
+\pagestyle{empty}
 \begin{figure}[htbp]
@@
 %------------------------------------------------------------------- Figure 2
+\pagestyle{empty}
 \begin{figure}[htbp]
@@
 %------------------------------------------------------------------- Figure 3
+\pagestyle{empty}
 \begin{figure}[htbp]
@@
 %------------------------------------------------------------------- Figure 4
+\pagestyle{empty}
 \begin{figure}[htbp]
-\centerline{\includegraphics[height=27cm]{Fig4.pdf}}
+\centerline{\includegraphics[height=26cm]{Fig4.pdf}}
 %

### 2016-06-26 Manuscript/MS- What is LDD.tex
Edits to Table 2 and Discussion.

@@
 \begin{equation}
-					N_b= 4 \pi \sigma^2 \delta
+					N^b_e= 4 \pi \sigma \delta
 \end{equation}
 
-where $\delta$ is the density of adults per unit area and $\sigma$ is the variance in distance between birth and breeding sites.
+where $\delta$ is the density of adults per unit area and $\sigma$ is the standard deviation of the distance between birth and breeding sites.
 This formulation is often called the neighborhood size and assumes a normal distribution of distances between parents and offspring (out in a perfect circular shape from the source).
-Thus, changes in the variance of dispersal distance can affect $N_e$ (highly clumped populations will have reduced $N_e$).
+Thus, changes in the variance of dispersal distance can affect $N^b_e$ (highly clumped populations will have reduced $N^b_e$).
 This is the basic model of "Isolation by Distance" proposed by   \citet{Wright:1943aa,Wright:1946aa}.
 Under this type of model, migration (gene flow) is given by the variance in dispersal, and not by the proportion of the population that is composed of migrants (denoted $m$), as is the case with island models \citep{Slatkin:1985qb}.
+With enough distance separating them, two plant individuals have a low probability of mating and can be considered members of distinct genetic populations even if they are not located in geographically distinct populations.
 
@@
 \begin{equation}
-  			N_b = 4 \pi (\sigma^2_s + \frac{t \sigma^2_p}{2}) \delta (1 + t)
+  		N^b_e = 4 \pi (\sigma^2_s + \frac{t \sigma^2_p}{2}) \delta (1 + t)
 \end{equation}
 
-where $\sigma^2_s$ is the variance in seed dispersal, $\sigma^2_s$ is the variance in pollen dispersal and $\delta$ is the density of potential parents.
+where $\sigma_s$ is the standard deviation of seed dispersal distance, $\sigma_p$ is the standard deviation of pollen dispersal distance, and $\delta$ is the density of potential parents.
 
@@
 The size of neighborhood areas may encompass at least four orders of magnitude, $10{^-2}-10^2$ km in radius, and include many individuals.
-Therefore, reference to this "genetic/evolutionary" paradigm and reference \citep[sensu][]{Waples:2006ev} may be instrumental to understand the actual role of LDD events in shaping the structuring of genetic pools and contributing to gene dispersal.
+Therefore, reference to this "genetic/evolutionary" paradigm and reference to the geographic boundaries \citep[sensu][]{Waples:2006ev} may be instrumental to understand the actual role of LDD events in shaping the structuring of genetic pools and contributing to gene dispersal.
 
-Whenever there is a large discrepancy between population area extent and $N_b$ we might expect the frequency of $LDD_{loc}$ and $LDD_{neigh}$ differ enormously.
-For example, relatively small $N_b$ may rise the importance of $LDD_{loc}$ in preserving scenarios of panmixia within a local population, as most distant dispersal events will disseminate seeds outside the neighborhood of maternal plants.
+Whenever there is a large discrepancy between population area extent and $N^b_e$ we might expect the frequency of $LDD_{loc}$ and $LDD_{neigh}$ differ enormously.
+For example, relatively small $N^b_e$ may rise the importance of $LDD_{loc}$ in preserving scenarios of panmixia within a local population, as most distant dispersal events will disseminate seeds outside the neighborhood of maternal plants.
 
@@
 Recently, DNA-barcoding techniques have been developed and successfully applied to the identification of frugivore species contributing to specific seed dispersal events whose source can be identified with genetic, direct assignment techniques \citep{GonzalezVaro:2014ij}.
-Otherwise, visual identification can reliably assign the genotyped seeds to frugivore species groups based on specific characteristics of scats and regurgitations \citep{Jordano:2007a}.
+Otherwise, visual identification can reliably assign the genotyped seeds to frugivore species groups based on specific characteristics of scats and regurgitations \citep{Jordano:2007}.
+
+We inferred the frugivore groups contributing dispersal events by visually identifying scats and regurgitations in seed traps and line transects \citep[see ][ and Suppl.
+Mat.
+for additional details of methods]{Jordano:2007}.
+These frugivore functional groups include up to 38 bird and 4 mammal species feeding on \textit{P.
+mahaleb} fruits \citep{Jordano:2000ft}.
+Here, we differentiate four major frugivore groups: large carnivorous mammals (such as foxes, badgers, and stone martens); two species of medium-sized frugivorous birds, mistle thrushes (\textit{T.
+viscivorus}), and carrion crows (\textit{Corvus corone}); and a pool of small-sized frugivorous birds, including warblers, redstarts, and robins \citep{Jordano:2007}.
+
+To a large extent, short-distance dispersal events (strict-sense, $SDD_{loc}$ events) are contributed by small- and medium-sized (\textit{Turdus}) frugivorous birds (Table 2).
+Given the relatively reduced $N^b_e$ area of \textit{P.
+mahaleb} (Suppl.
+Mat.
+Table S1), $< 1km^2$, well below the extent of the local study population \citep{Garcia:2007he,Garcia:2005fu},  we cannot estimate $LDD_{neigh}$ events (Table 2), as all LDD events outside the reference population occur, by definition, outside the $N^b_e$ area.
+Larger frugivores such as corvids and the pigeon \textit{Columba palumbus} contribute most LDD events, and most immigrant seeds potentially dispersed from other populations.
+Notably, strict-sense long-distance dispersal ($LDD_{ss}$) appears consistently associated with large-bodied frugivores (Table 2), most likely associated with a greater frequency of movements outside the local population (Fig.
+4).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Discussion
+%\section*{Discussion}
 
@@
 
-As defined in our framework (Table 1), LDD, and in particular $LDD_{ss}$ events are a particular case of extreme events \citep{Garcia:2017aa}.
+Combining spatially-explicit references to the geographic population limits and the $N^b_e$ area extent helps avoiding some imprecision in setting distance thresholds to characterize LDD events.
+In addition the framework outlined in Table 1 bridges the combined demographic and genetic effects of LDD events.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Discussion
-%\section*{Discussion}
+As defined in our framework (Table 1), LDD, and in particular $LDD_{ss}$ events are a specific case of extreme events \citep{Garcia:2017aa} consistently associated with large-sized frugivores, yet including also medium-sized and highly efficient frugivorous bird species.
 
-\subsection*{Challenges and Promising Avenues for Research}
+- Large-bodied frugivores and LDD.  Discuss lb frugivores not always contributing LDD events: Pratt, tapirs
+- Loss of large-bodied frugivores, consequences.
 
+\subsection*{Challenges and future avenues for research}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Acknowledgements
 \emph{Acknowledgements}.
@@
 %\section*{References}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% References
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% References
 %%% Unnumbered Literature Cited
@@
 \captionsetup{width=21cm}%.75\textwidth}
-\caption{Frequencies of \textit{Prunus mahaleb} seed dispersal events for different frugivore groups according to population area limits and genetic neighborhood limits.
+\caption{Relative frequencies of \textit{Prunus mahaleb} seed dispersal events for different frugivore groups according to population area limits and genetic neighborhood limits.
 See Fig.
-2 for a graphical representation of the four scenarios.}
+2 for a graphical representation of the four scenarios.
+\textit{N}= 655 seeds \citep[see Table 1 in ][]{Jordano:2007}.
+Given that the estimated neighborhood size is smaller than population area, $LDD_{neigh}$ would be zero.}
 \vspace{0.5cm}
-    \begin{tabular}{lp{3cm}p{3cm}p{3cm}p{3cm}}
+    \begin{tabular}{lp{3cm}p{3cm}p{3cm}p{3cm}p{3cm}}
     \hline \\
-    Frugivore group & Within-population, within-neighborhood $SSD_{loc}$ & Within-population, long-distance $LDD_{loc}$ & Outside-population, within-neighborhood $LDD_{neigh}$ & Strict-sense long-distance $LDD_{ss}$ \\\\ \hline \\
-    Small birds     & ?~           & ?~            & ?~           & ?~  \\
-    \textit{Turdus} & ?~           & ?~            & ?~           & ?~  \\
-
-    Large birds     & ?~           & ?~            & ?~           & ?~  \\
-
-    Mammals         & ?~           & ?~            & ?~           & ?~  \\\\
+    Frugivore group & Within-population, within-neighborhood $SDD_{loc}$ & Within-population, long-distance $LDD_{loc}$ & Outside-population, within-neighborhood $LDD_{neigh}$ & Strict-sense long-distance $LDD_{ss}$ & \textit{N} seeds  \\\\ \hline \\
+	Small-birds	    & 0.7842 &  0.0171 & 0.00 & 0.1986 & 292  \\
+	\textit{Turdus} & 0.2370 &  0.5549 & 0.00 & 0.2081 & 173  \\
+	Large-birds     & 0.0435 &  0.3913 & 0.00 & 0.5652 &  23  \\
+	Mammals         & 0.0120 &  0.2455 & 0.00 & 0.7425 & 167  \\\\
 \hline
@@
 % -------------- Crawford 21986 Heredity ------------------------------------
-Sewal Wright's neighborhood model indicates that the area containing a panmictic unit within a continuous and uniform array of organisms can be 
+Sewall Wright's neighborhood model indicates that the area containing a panmictic unit within a continuous and uniform array of organisms can be 
 estimated by $4\pir^2$ where a-2 is the parent-offspring dispersal variance measured around a zero  mean and relative to a single reference axis passing through the population.

### 2016-06-26 Manuscript/SupplMat.tex
Added Fig 4 and moved fig to Suppl Mat. Edited figure legends.
Edits to Table 2 and Discussion.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Headers
-\documentclass[a4paper, 12pt]{article}}
+\documentclass[a4paper, 12pt]{article}
 \usepackage{graphicx}
@@
 \usepackage{rotating}                                                                                     
-\usepackage{standalone}                                                                                    
 \newcommand{\tab}{\hspace{5mm}}
@@
 \usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
-\pagestyle{fancy}     % options: empty , plain , fancy
+\pagestyle{empty}     % options: empty , plain , fancy
 \renewcommand{\headrulewidth}{0pt} % customise the layout...
@@
 \begin{table}
-  \caption{Summary of neighborhood area sizes and estimated neighborhood radius for tree species with different combinations of dispersal modes. Data from Nason et al and present study. }
+  \caption{Summary of neighborhood area sizes and estimated neighborhood radius for tree species with different combinations of dispersal modes.
+Data from \citet{Nason:1998aa,Garcia:2005fu,Garcia:2007he} and present study.}
     \begin{tabular}{lllccc}
@@
 
+%------------------------------------------------------------------ Figure S2
+\begin{figure}[htbp]
+\centerline{\includegraphics[height=22cm]{FigS2.pdf}}
+%
+\caption{Differential contributions of functional groups of frugivores to the short- ($SDD_{loc}$) and long-distance ($LDD_{loc}$) local seed dispersal events for \textit{Prunus mahaleb}.}
+\end{figure}
+
+\newpage
+
+%\section*{References}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% References
+%%% Unnumbered Literature Cited
+\bibliographystyle{bes}        %Compile with bes.bst style file
+\bibliography{MS_BES}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-06-26 Manuscript/MS- What is LDD.tex
Edits for Discussion.

@@
 Combining spatially-explicit references to the geographic population limits and the $N^b_e$ area extent helps avoiding some imprecision in setting distance thresholds to characterize LDD events.
-In addition the framework outlined in Table 1 bridges the combined demographic and genetic effects of LDD events.
+In addition, the framework outlined in Table 1 bridges the combined demographic and genetic effects of LDD events.
+When methods available to assign frugivore taxa to the analyzed dispersal events, as in the study case with \textit{P.
+mahaleb}, a classification in the four categories of events is possible.
+The frugivore assemblage of \textit{P.
+mahaleb} is composed by a diversified set of animal species spanning a wide rage of body sizes, ca.
+12-14000 g in weight.
+We might expect that this extreme variation translates in an ample pattern of foraging modes, movement distances, and fruit/seed processing \citep{Jordano:2000ft}.
+If the results for \textit{P.
+mahaleb} are generalizable to other disperser assemblages, it seems that the functional roles of frugivore species in terms of contributions to LDD events are structured in two distinct groups: small-bodied frugivores, with substantial contributions to SDD events and a crucial role for in-situ population recruitment, and large-bodied species with a disproportionate contribution to LDD events.
 

### 2016-06-27 Manuscript/SupplMat.tex
Major edits and finishing Suppl. Material.

@@
 
+%-------------------------------------------------------------------- METHODS
+\begin{linenumbers}
+
+\section*{Methods}
+
+\paragraph*{Species and Study Site.} The study species, \textit{Prunus mahaleb} (L.) (Rosaceae), is a shrub or small tree that produces fleshy fruits that are consumed by frugivores, who disperse seeds after regurgitating or defecating them.
+This species is frequently visited during July to mid-August by small- and medium-sized birds and carnivorous mammals that include fruits in their diets during late summer to winter \citep{Jordano:2000ft}.
+\textit{P.
+mahaleb} occurs in a patchy distribution at the regional scale, with relatively isolated populations consisting of dozens to hundreds of trees.
+Our study population included a total of 196 adult reproductive trees distributed over an area of 26 ha in patches of variable density.
+Other populations within 20 km exist as scattered patches of 10–150 trees, with some containing $\geqq$ 1,000 trees.
+The nearest population is 1.5 km away.
+Additional information on the study population is reported by \citet{Jordano:2007} and \citet{Garcia:2009do}.
+
+\paragraph*{Sampling dispersed seeds.} 
+To estimate the relative contribution of each dispersal vector to the different categories of dispersal events defined in Table 1, we first collected dispersed seeds, following different sampling schemes according to the functional group of dispersal vector.
+We used this grouping of frugivores giving the difficulties of resolving the identification of scats, pellets and regurgitated seeds down to species level just based on visual cues.
+We differentiated four major frugivore types: large carnivorous mammals (such as foxes, badgers, and stone martens); two species of medium- and large-sized frugivorous birds, mistle thrushes (\textit{T.
+viscivorus}), and carrion crows (\textit{C.
+corone}); and a pool of small-sized frugivorous birds, including warblers, redstarts, and robins \citep{Jordano:2007}.
+
+Seeds were collected in 1997–1999 and 2003–2005.
+The sampling schemes are described in detail elsewhere \citep{Jordano:2007,Garcia:2009do} and include a combination of seed traps and direct sampling of mammal feces along fixed transects.
+We haphazardly collected 130 samples of mammal feces during the \textit{P.
+mahaleb} fruit ripening period and recorded their location relative to potential source trees.
+Overall, we genotyped 167 seeds from 20 fecal samples.
+Most samples were from red fox (\textit{Vulpes vulpes}) and stone marten (\textit{Martes foina}); some ( 10 samples) were from badger (\textit{Meles meles}) \citep{Jordano:2007}.
+
+In addition we sampled directly the pellets of large corvids (\textit{Corvus corone}) and from \textit{Turdus viscivorus}, the latter by direct sampling beneath pine trees and scats from seed traps \citep[see ][ for details]{Jordano:2007}.
+Finally, a seed sample directly from seed traps included seeds dispersed by small- and medium-sized passerines, such as \textit{Phoenicurus ochruros}, \textit{Turdus merula}, \textit{Erithacus rubecula}, \textit{Sylvia communis}, \textit{Sylvia atricapilla}, etc.
+\citep{Jordano:2007}.
+The total seed sample thus consisted of seed endocarps collected from the seed traps (mostly small passerines) ($n= $465), mammal scats ($n= $167), and \textit{C.
+corone} pellets ($n= $23).
+
+\paragraph*{Seed genotyping.} 
+We used material described in \citet{Jordano:2007}, and genotyping methods described in detail in previous work \citep{Godoy:2001,Garcia:2007he,Garcia:2009do}.
+Briefly, we used a set of 10 polymorphic microsatellite markers (simple DNA sequence repeats) \citep{Godoy:2001} to obtain the multilocus genotypes of both of the adult trees (candidate source trees from the study population) and the sample of seed endocarps.
+Given that all adult trees in the population had a distinct multilocus genotype, an unambiguous assignment of each seed to its source tree could be made.
+When a full match between the endocarp genotype and any of the adult-tree genotypes in the population was not possible, we assumed that the seed came from another population.
+To assess the effect of genotyping errors, we reexamined the exclusion of genotypes due to a single locus mismatch, two loci mismatches, etc.
+At the analysis level, any exclusion of identity between a seed and a potential mother tree based on mismatches of only one or two loci was rechecked.
+We used GIMLET software \citep{Valiere:2002fo} to find the matching adult multilocus genotype for each endocarp with eight or more loci successfully typed.
+Because each seed belonged to one of the four groups of dispersers, we could thus derive the relative contribution of each frugivore group to different classes of seed dispersal events and to seed immigration.
+
+\paragraph*{Contribution of dispersal vectors to types of dispersal events.} 
+We considered each dispersed seed as an independent replicate, because each represented a dispersal event from the perspective of plant population genetics, i.e., an independent "arrival" event resulting from the dispersal process mediated by the frugivore.
+
+Once the maternal source tree of each individual seed was identified (or its provenance from outside the study population determined) we assessed the dispersal distance and grouped the seeds separately as coming from trees locate within or outside the population.
+In addition, for seeds originating from local trees we determined whether dispersal distances were $\geqq$ 45 m to sort out $LDD_{loc}$ dispersal events from $SDD_{loc}$ events.
+All the events involving immigrant seeds were considered $LDD_{ss}$ by definition, given that the neighborhood size was very reduced (radius= 0.045 km) relative to the geographic limits of the study population (maximum length for a within-population dispersal event: 1220 m)\citep{Garcia:2009do}.
+Along this reasoning, $LDD_{neigh}$ events were considered non-existent in this particular case study given that neighborhood size area was smaller than the population area.
+
+\end{linenumbers}
+\newpage
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Supplementary Material
@@
 \begin{table}
-  \caption{Summary of neighborhood area sizes and estimated neighborhood radius for tree species with different combinations of dispersal modes.
+  \caption*{Table S1.
+Summary of neighborhood area sizes and estimated neighborhood radius for tree species with different combinations of dispersal modes.
 Data from \citet{Nason:1998aa,Garcia:2005fu,Garcia:2007he} and present study.}
@@
 %
-\caption{Dispersal events for pollen (left) and seeds (right) traced for \textit{Prunus mahaleb} trees (white dots).
+\caption*{Figure S1.
+Dispersal events for pollen (left) and seeds (right) traced for \textit{Prunus mahaleb} trees (white dots).
 All the adult, reproductive, trees in the population are mapped.
@@
 %
-\caption{Differential contributions of functional groups of frugivores to the short- ($SDD_{loc}$) and long-distance ($LDD_{loc}$) local seed dispersal events for \textit{Prunus mahaleb}.}
+\caption*{Figure S2. Differential contributions of functional groups of frugivores to the short- ($SDD_{loc}$) and long-distance ($LDD_{loc}$) local seed dispersal events for \textit{Prunus mahaleb}.}
 \end{figure}

### 2016-06-28 Manuscript/MS- What is LDD.tex
Minor edits to SupplMat and figure legends.

@@
 \textbf{Figure 2}.
-Schematic representation of different types of long-distance dispersal events in relation to the geographical limits of local populations (dashed lines) and the genetic neighborhood area $N_b$ (grey area) of specific individual plants (squares).
+Schematic representation of different types of long-distance dispersal events in relation to the geographical limits of local populations (dashed lines) and the genetic neighborhood area $N^b_e$ (grey area) of specific individual plants (squares).
 Dispersal events (arrows) can be classified depending on their actual incidence on propagule movement outside these spatially-explicit reference areas (Table 1).
 Strict-sense long-distance dispersal events ($LDD_{ss}$) just include the LDD events that disseminate propagules out of \textit{both} the population and genetic neighborhood boundaries.
-A, the neighborhood area is included within the geographic limits of the population; B, the neighborhood area is much larger than the geographic limits of the population.\\
+A, the neighborhood area is included within the geographic limits of the population, with some dispersal events potentially contributing local LDD; B, the neighborhood area is much larger than the geographic limits of the population.\\
 
@@
 Empirical frequency distributions of seed dispersal events as a function of dispersal distance by animal frugivores consuming \textit{Prunus mahaleb} fruits.
-In red, left (inset), frequencies of within-populations dispersal events inferred from direct assignment based on seed endocarp genotypes and maternal trees genotypes.
-Larger frame, left, contributions of four functional frugivore groups (small birds, medium- and large-sized birds, and mammals) to seed dissemination and proportional contributions (right bar) to dispersal of the inferred immigrant seeds (i.e., those not matching any maternal tree in the study population).\\
+In red, left (inset), frequencies of within-population dispersal events inferred from direct assignment based on seed endocarp genotypes and maternal trees genotypes.
+Larger frame, left, contributions of four functional frugivore groups (small birds, medium- and large-sized birds, and mammals) to seed dissemination and proportional contributions (right bar) to dispersal of the inferred immigrant seeds (i.e., those not matching any maternal tree in the study population) \citep{Jordano:2007}.\\
 
@@
 Differential contributions of functional groups of frugivores to the four combinations of \textit{Prunus mahaleb} seed dispersal events outlined in Table 1.
-These result from dissemination within (yellow) or outside (blue) the population geographic limits ($SDD_{loc}$, $LDD_{loc}$, respectively) and within or outside the genetic neigborhood area limits ($SDD_{neigh}$, $LDD_{ss}$, respectively).\\
+These result from dissemination within (yellow) or outside (blue) the population geographic limits; within-population dispersal events can either be short-distance ($SDD_{loc}$) or local LDD ($LDD_{loc}$) depending on the size of the genetic neigborhood.
+Dispersal outside the local population can entail short-distance dispersal, if within the genetic neighborhood area limits ($SDD_{neigh}$) (yellow) or represent strict-sense LDD ($LDD_{ss}$) (blue).\\
 \newpage

### 2016-06-28 Manuscript/SupplMat.tex
Minor edits to SupplMat and figure legends.

@@
 
-\paragraph*{Species and Study Site.} The study species, \textit{Prunus mahaleb} (L.) (Rosaceae), is a shrub or small tree that produces fleshy fruits that are consumed by frugivores, who disperse seeds after regurgitating or defecating them.
+\paragraph*{Species and Study Site.} The tree species we use as a case study to illustrate different types of dispersal events, \textit{Prunus mahaleb} (L.) (Rosaceae), is a shrub or small tree that produces fleshy fruits that are consumed by frugivores, who disperse seeds after regurgitating or defecating them.
 This species is frequently visited during July to mid-August by small- and medium-sized birds and carnivorous mammals that include fruits in their diets during late summer to winter \citep{Jordano:2000ft}.
@@
 The nearest population is 1.5 km away.
-Additional information on the study population is reported by \citet{Jordano:2007} and \citet{Garcia:2009do}.
+Additional information on the study population and description of methodological apporaches is reported by \citet{Jordano:2007} and \citet{Garcia:2009do} and references therein.
 
@@
 The total seed sample thus consisted of seed endocarps collected from the seed traps (mostly small passerines) ($n= $465), mammal scats ($n= $167), and \textit{C.
-corone} pellets ($n= $23).
+corone} pellets ($n= $23) \citep[see Table 1 in ][]{Jordano:2007}.
 
@@
 
-Once the maternal source tree of each individual seed was identified (or its provenance from outside the study population determined) we assessed the dispersal distance and grouped the seeds separately as coming from trees locate within or outside the population.
+Once the maternal source tree of each individual seed was identified (or its provenance from outside the study population determined) we assessed the dispersal distance and grouped the seeds separately as coming from trees located within or outside the population.
 In addition, for seeds originating from local trees we determined whether dispersal distances were $\geqq$ 45 m to sort out $LDD_{loc}$ dispersal events from $SDD_{loc}$ events.

### 2016-06-28 Manuscript/MS- What is LDD.tex
Edits to Discussion.

@@
 \usepackage{parskip}
-\linespread{1.25}
+\linespread{1.75}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title page
@@
 Long-distance dispersal (LDD) is a major component of the population dynamics, genetic structure, and biogeographic history of plant species.
-It determines the colonization ability of new habitats and the possibilities for fragmented populations to sustain a cohesive metapopulation by immigration-emigration dynamics that rely on LDD events.
+It determines the colonization ability of new habitats and the possibilities for fragmented populations to sustain a cohesive metapopulation by immigration-emigration dynamics that rely on LDD events \citep{Nathan:2008is,Schurr2009long}.
 Yet our current understanding of the extent, frequency, and consequences of LDD is very limited.
@@
 
-Combining spatially-explicit references to the geographic population limits and the $N^b_e$ area extent helps avoiding some imprecision in setting distance thresholds to characterize LDD events.
+Combining spatially-explicit references to the geographic population limits and the genetic neighborhood area extent ($N^b_e$) helps avoiding some imprecision in setting distance thresholds to characterize LDD events \citep{Jones:2008il}.
 In addition, the framework outlined in Table 1 bridges the combined demographic and genetic effects of LDD events.
@@
 mahaleb}, a classification in the four categories of events is possible.
+
 The frugivore assemblage of \textit{P.
-mahaleb} is composed by a diversified set of animal species spanning a wide rage of body sizes, ca.
-12-14000 g in weight.
+mahaleb} is composed by a diversified set of animal species spanning a wide size range, ca.
+12-14000 g in body mass.
 We might expect that this extreme variation translates in an ample pattern of foraging modes, movement distances, and fruit/seed processing \citep{Jordano:2000ft}.
 If the results for \textit{P.
-mahaleb} are generalizable to other disperser assemblages, it seems that the functional roles of frugivore species in terms of contributions to LDD events are structured in two distinct groups: small-bodied frugivores, with substantial contributions to SDD events and a crucial role for in-situ population recruitment, and large-bodied species with a disproportionate contribution to LDD events.
+mahaleb} are generalizable to other disperser assemblages, it seems that the functional roles of frugivore species in terms of contributions to LDD events are structured in two distinct groups: small-bodied frugivores, with substantial contributions to SDD events, and large-bodied species with a disproportionate contribution to LDD events.
+Both components of this sort of diplochorous \citep{vanderWall:2004hv} dispersal system are very frequent in fleshy-fruited plants with diversified frugivore assemblages \citep{Galetti:2013jv}.
+In such cases, small-bodied frugivores largely contribute the short-distance dispersal key to support \textit{in situ} recruitment and population persistence.
+Yet the large-bodied frugivores distinctly contribute LDD events that sustain the connectivity of metapopulation scenarios \citep{Urban:2001}.
+Yet, as shown in Table 1, SDD and LDD events can be more complex when we consider the contributions to gene flow via seed and the consequences in terms of structure and spatial distributions of the genetic pools.
+
+A number of classic studies have demonstrated that the activity of large furgivores may also significantly contribute to SDD events and inefficient dispersal: i.e., territorial defence, short gut retention times relative to on-tree foraging, frequent revisititation of same trees and perches, etc., resulting in substantial SDD events \citep{Pratt:1983gf,Pratt:1984ly,Snow:1984ul,Snow:1988ve,Wheelwright:1991qf}.
+Yet these large-bodied frugivores are crucial for both $LDD_{loc}$ and $LDD_{ss}$, given that extensive movement patterns and extremely large foraging ranges may frequently contribute dissemination beyond distance thresholds defined with either spatial landscape or genetic references.
+Recent analyses of the movement ecology of large frugivores, coupled with results of their seed dispersal services emphasize that LDD are by no means exceptional, either in terms of frequency and extent \citep{Westcott:2005,Bueno:2013cg}.
 
@@
 \subsection*{Challenges and future avenues for research}
+
+Pollen and seed dispersal in plants are essentially spatially-structured processes for which the outcomes of interactions with dispersal vectors is intimately linked to landscape features.
+Given this mechanistic link between the features of the vector and the environments where its displacement occurs \citep{Nathan:2008fx}, consideration of landscape is key to understand the consequences of LDD events.
+Yet these consequences hit two central aspects of plant life-histories: the demographic recruitment process \citep{Harper:1977aa}, and the genetic signatures of pollen- and seed-mediated gene flow .
 

### 2016-06-28 Manuscript/MS- What is LDD.tex
First draft finished.

@@
 
-\textbf{Key words}: ***\\
+\textbf{Key words}: dispersal, frugivory, plant-animal interactions, pollination, seed dispersal\\
 
@@
 Yet the large-bodied frugivores distinctly contribute LDD events that sustain the connectivity of metapopulation scenarios \citep{Urban:2001}.
-Yet, as shown in Table 1, SDD and LDD events can be more complex when we consider the contributions to gene flow via seed and the consequences in terms of structure and spatial distributions of the genetic pools.
+As shown in Table 1, SDD and LDD events can be more complex when we consider the contributions to gene flow via seed and the consequences in terms of structure and spatial distributions of the genetic pools.
+For example, local, within-population, dispersal events may vary enormously in terms of genetic effects and local structuring of the genetic pools depending on whether they specifically contribute $SDD_{loc}$ or instead, $LDD_{loc}$.
+Note that only the latter actually contribute erasing any form of local genetic structure by contributing to increased genetic neigborhoods.
 
-A number of classic studies have demonstrated that the activity of large furgivores may also significantly contribute to SDD events and inefficient dispersal: i.e., territorial defence, short gut retention times relative to on-tree foraging, frequent revisititation of same trees and perches, etc., resulting in substantial SDD events \citep{Pratt:1983gf,Pratt:1984ly,Snow:1984ul,Snow:1988ve,Wheelwright:1991qf}.
+A number of classic studies have demonstrated that the activity of large furgivores may also significantly contribute to SDD events and inefficient dispersal because of, i.e., territorial defence, short gut retention times relative to on-tree foraging, frequent revisitation of same trees and perches, etc., resulting in substantial SDD events \citep{Pratt:1983gf,Pratt:1984ly,Snow:1984ul,Snow:1988ve,Wheelwright:1991qf}.
 Yet these large-bodied frugivores are crucial for both $LDD_{loc}$ and $LDD_{ss}$, given that extensive movement patterns and extremely large foraging ranges may frequently contribute dissemination beyond distance thresholds defined with either spatial landscape or genetic references.
-Recent analyses of the movement ecology of large frugivores, coupled with results of their seed dispersal services emphasize that LDD are by no means exceptional, either in terms of frequency and extent \citep{Westcott:2005,Bueno:2013cg}.
+Recent analyses of the movement ecology of large frugivores, coupled with results of their seed dispersal services emphasize that LDD are by no means exceptional, either in terms of frequency and extent \citep[e.g., ][]{Westcott:2005,Bueno:2013cg,Morales:2013dg,Carlo:2013vd}.
+In addition, medium-sized birds such as thrushes (\textit{Turdus} spp.) can contribute substantial $LDD_{loc}$ events, i.e., local LDD events contributing to erase local population genetic structuring, effectively increasing the size of genetic neighborhoods.
+In the case of \textit{P.
+mahaleb} up to 55.49\% of their dispersal events are $LDD_{loc}$ events.
+These birds are efficient seed dispersers of \textit{P.
+mahaleb} and other fleshy-fruited species \citep{Snow:1988ve,Jordano:2000ft,Carlo:2013vd}, also showing significant contributions of $LDD_{ss}$ events.
 
 As defined in our framework (Table 1), LDD, and in particular $LDD_{ss}$ events are a specific case of extreme events \citep{Garcia:2017aa} consistently associated with large-sized frugivores, yet including also medium-sized and highly efficient frugivorous bird species.
-
-- Large-bodied frugivores and LDD.  Discuss lb frugivores not always contributing LDD events: Pratt, tapirs
-- Loss of large-bodied frugivores, consequences.
+Robustly characterizing the expected frequencies and extent of those extreme events would be crucial to properly assess the functional role of frugivores and the full range of influences (demographic, genetic) in plant populations.
 
@@
 Given this mechanistic link between the features of the vector and the environments where its displacement occurs \citep{Nathan:2008fx}, consideration of landscape is key to understand the consequences of LDD events.
-Yet these consequences hit two central aspects of plant life-histories: the demographic recruitment process \citep{Harper:1977aa}, and the genetic signatures of pollen- and seed-mediated gene flow .
+Yet these consequences hit two central aspects of plant life-histories: the demographic recruitment process \citep{Harper:1977aa}, and the genetic signatures of pollen- and seed-mediated gene flow in complex landscapes \citep{Sork:1999}.
+Recent evidences point out that the selective extinction of large-bodied frugivores may significantly impact plant populations dependent on frugivores both in terms  of recruitment \citep{Traveset:2012he,PerezMendez:2015hya} and genetic connectivity \citep{Perez-Mendez:2016dz}.
+Frugivore downsizing represents a lasting challenge for the collapse of seed dispersal processes where $LDD_{ss}$ events are crucial for population persistence and the cohesion of fragmented populations within metapopulation scenarios.
+
+We advocate \citep[also see ][]{Jordano:2002tr,Nathan:2003qe,Jones:2008il,Hardesty:2011jn} a combination of approaches including large-scale biotelemetry to characterize animal movement, coupled with large-scale genetic sampling of dispersed propagules, and demogenetic approaches that combine both demographic and genetic research.
+A crucial aspect would be to effectively associate the role of individual frugivore species to specific dispersal outcomes, by identifying the actual disperser contributing a dissemination event \citep{GonzalezVaro:2014ij} and simultaneously characterizing the source maternal plant \citep{Jordano:2002tr}.
+
+The actual challenges to properly characterize the typologies of LDD events outlined in Table 1 will probably persist.
+We need more efficient quantitative approaches to assess these infrequent events, that occur over enormous spatial scales and that need to be documented with sample sizes sufficient to facilitate modeling efforts and robust statistical inferences.
+These are not trivial difficulties given the urgency to assess how forest loss, defaunation, genetic purging due to logging, etc., alter plant populations.
+
+\newpage
 

### 2016-06-28 Manuscript/MS- What is LDD.tex
Minor edits.

@@
 LDD is a characteristically extreme event of propagule movement in any plant population, typically occurring with an extremely low probability but potentially reaching an extremely long distance.
-Some progress has recently been made through the fast-paced implementation of molecular tools in ecological research labs and the availability of cutting-edge technology for biotelemetry applications [REF].
+Some progress has recently been made through the fast-paced implementation of molecular tools in ecological research labs and the availability of cutting-edge technology for biotelemetry applications.
 But much of the population geneticist and ecologist communities remains unaware of the state of the art in each other and likely under-appreciates their potential to validate and enrich dispersal studies \citep{Jones:2008il}.
@@
 The exact proportional or absolute thresholds selected remain arbitrary, as no reference spatial frame is provided within the definition of LDD.
-This leaves the consideration of LDD as an extreme form of context-dependent phenomenon, strongly dependent upon the scale of the biological process studied \citep{Kinlan:2005fb}.
+This leaves the consideration of LDD as an extreme form of context-dependent phenomenon, strongly dependent upon the scale of the biological process studied \citep{Kinlan:2005fb} and of the specific organism considered.
 For example, \cite{Kinlan:2005fb} used a spatial reference frame to characterize LDD events of marine organisms, where sedentary adults and larvae differ enormously in the spatial scales of their dispersal \citep{DAloia:2013fc}.
@@
 
-We aim at providing a general framework for the quantitative analysis of LDD events so that estimates of its frequency and extent could be comparable across different study systems.
+We aim at providing a general framework for the quantitative analysis of LDD events so that estimates of their frequency and extent could be comparable across different study systems.
 We argue that both demographic and genetic elements are needed for this framework, most likely requiring a combination of field-based movement data and genetic analyses.
@@
 
-A second consideration in terms of spatial boundaries, with effects on dispersal patterns, is the genetic neighborhood area $N_b$, i.e., the spatial extent including a subset of panmictic individuals within a population \citep{Wright:1943aa,Wright:1946aa}.
-Thus, the $N_b$ area can be equal to the whole extent of the population whenever the population is unstructured and there is evidence for random mating events among all the individuals.
+A second consideration in terms of spatial boundaries, with effects on dispersal patterns, is the genetic neighborhood area $N^b_e$, i.e., the spatial extent including a subset of panmictic individuals within a population \citep{Wright:1943aa,Wright:1946aa}.
+Thus, the $N^b_e$ area can be equal to the whole extent of the population whenever the population is unstructured and there is evidence for random mating events among all the individuals.
 However, most populations and stands of long-lived trees show highly aggregated and clumped distributions \citep{Seidler:2006hx}, where relatively long distances may separate groups of individuals within the same population.
-In these cases we might expect $N_b$ area to be substantially smaller than the total population area.
+In these cases we might expect $N^b_e$ area to be substantially smaller than the total population area.
 Therefore, at least four possible scenarios exist with distinct implications in terms of consequences for dispersal (Table 1).
@@
 Table S1), $< 1km^2$, well below the extent of the local study population \citep{Garcia:2007he,Garcia:2005fu},  we cannot estimate $LDD_{neigh}$ events (Table 2), as all LDD events outside the reference population occur, by definition, outside the $N^b_e$ area.
-Larger frugivores such as corvids and the pigeon \textit{Columba palumbus} contribute most LDD events, and most immigrant seeds potentially dispersed from other populations.
+Larger frugivores such as corvids and the pigeon \textit{Columba palumbus} contribute most LDD events, and most immigrant seeds potentially dispersed from other populations (Fig.
+S2).
 Notably, strict-sense long-distance dispersal ($LDD_{ss}$) appears consistently associated with large-bodied frugivores (Table 2), most likely associated with a greater frequency of movements outside the local population (Fig.
 4).
+
+Empirically mapping of dispersal events for either pollen or seed disseminated by animals may result in a complex pattern of different combinations of dispersal events (Fig.
+S1), as animal movements are overlaid onto plant populations occupying complex landscapes, resulting in different types of SDD and LDD events.
 

### 2016-09-02 Manuscript/MS- What is LDD.tex
Minor edits to title page.
Edits to cover page. Start revised R1 version for submission
Minor edit for typos.

@@
 \renewcommand{\headrulewidth}{0pt} % customise the layout...
-\lhead{{\tiny Jordano - What is long-distance dispersal?}}\chead{}\rhead{}
+\lhead{{\tiny Jordano - What is long-distance dispersal?}}\chead{}\rhead{{\tiny JEcol-2016-0422.R1}}
 \lfoot{}\cfoot{\thepage}\rfoot{}
@@
 
-\title{Manuscript Draft\\
+\title{What is long-distance dispersal? And a taxonomy of dispersal events\\
 \vspace{2cm}
-What is long-distance dispersal? And a taxonomy of dispersal events}
+}
 
-\author{Pedro Jordano$^{\dag}$}
+\author{MS JEcol-2016-0422.R1\\\\Pedro Jordano$^{\dag}$}
 
 \date{Sevilla, \today}
+
 \maketitle
@@
 These "frontiers" set biological limits to what a LDD event is in relation to the geographic limits of the source population.
-Most plants are distributed as clumped patches, discrete stands, or relatively isolated populations, so we may distinguish between short-distance and long-distance dispersal events that end up with dissemination within or beyond, respectively, the stand or population geographic boundaries (Table 1, $SSD_{loc}$ or $LDD_{loc}$) (Figure 2).
+Most plants are distributed as clumped patches, discrete stands, or relatively isolated populations, so we may distinguish between short-distance and long-distance dispersal events that end up with dissemination within or beyond, respectively, the stand or population geographic boundaries (Table 1, $SDD_{loc}$ or $LDD_{loc}$) (Figure 2).
 
@@
 
-While both $SSD_{loc}$ and $LDD_{loc}$ can be crucial for assuring the local persistence of populations, $LDD_{neigh}$ and $LDD_{ss}$ would be extremely important contributors to the structuring of genetic pools, realized gene flow, and maintaining connectivity in metapopulation scenarios.
+While both $SDD_{loc}$ and $LDD_{loc}$ can be crucial for assuring the local persistence of populations, $LDD_{neigh}$ and $LDD_{ss}$ would be extremely important contributors to the structuring of genetic pools, realized gene flow, and maintaining connectivity in metapopulation scenarios.
 We argue that both the demographic and the genetic references are relevant for a proper definition of LDD.

### 2016-10-07 Manuscript/MS- What is LDD.tex
Edits abstract.
Edits to Abstract.

@@
 
+1.
 Dispersal is a key individual-based process influencing many life-history attributes, scaling up to population-level properties (e.g., metapopulation connectivity).
@@
 the long-distance dispersal (LDD) events, are difficult to characterize empirically and to model in realistic ways.
+\\
+2.
 But, when is it a LDD event?
@@
 Here we discuss operative definitions of LDD based on their intrinsic properties: 1) events crossing geographic boundaries among stands; and 2) events contributing to effective gene flow and propagule migration.
+\\
+3.
 Strict-sense long distance dispersal involves movement both outside the stand geographic limits and outside the genetic neighborhood area of individuals.
 Combinations of propagule movements within/outside these two spatial reference frames results in four distinct modes of LDD.
-Beyond traditional statistical approaches to characterize distributions, Extreme Value Analysis (EVA) can be used to properly and explicitly evaluate the properties of frequency and extent of LDD events.
-We discuss conditions where global change scenarios truncate dispersal processes, leading to the loss of key dispersal services in natural populations.
-Proper characterization of the LDD events helps to assess, for example, how the ongoing defaunation of large-bodied frugivores pervasively entails the loss of crucial LDD functions.
+\\
+4.
+\textit{Synthesis.} We discuss conditions where global change scenarios truncate dispersal processes, leading to the loss of key dispersal services in natural populations.
+Proper characterization of the LDD events helps to assess, for example, how the ongoing defaunation of large-bodied frugivores pervasively entails the loss of crucial LDD functions.\\
 

### 2016-10-09 Manuscript/MS- What is LDD.tex
Edited Abstract.

@@
 4.
-\textit{Synthesis.} We discuss conditions where global change scenarios truncate dispersal processes, leading to the loss of key dispersal services in natural populations.
+\textit{Synthesis.} We expect truncation of seed dispersal kernels to have multiple consequences on demography and genetics, following to the loss of key dispersal services in natural populations.
+Irrespective of neighborhood sizes, loss of LDD events may result in more structured and less cohesive genetic pools, with increased isolation-by-distance extending over broader areas.
 Proper characterization of the LDD events helps to assess, for example, how the ongoing defaunation of large-bodied frugivores pervasively entails the loss of crucial LDD functions.\\

### 2016-10-09 Manuscript/MS- What is LDD.tex
Edited rev #1 comments.

@@
 \usepackage{graphicx}
-\usepackage[utf8]{inputenc}
+%\usepackage[utf8]{inputenc}
 \usepackage[a4paper]{geometry}
@@
 
+% EDIT HERE
+While my main focus in on population-level scenarios and the role of animal vectors, recent research has shown how relevant is habitat and landscape context in determining the occurrence and directionality of LDD (e.g.
+Schurr et al.
+2008 J Ecol; Carlo et al.
+2013).
+ On the other hand, dealing with habitat effects seems obligated if aimed to discuss global change drivers (i.e.
+habitat loss and fragmentation).
 Two main conceptual approaches have been used to assess dispersal (Fig.
@@
 mahaleb} and other fleshy-fruited species \citep{Snow:1988ve,Jordano:2000ft,Carlo:2013vd}, also showing significant contributions of $LDD_{ss}$ events.
+
+Two-dimensional patterns in the \textit{P.
+mahaleb} seed rain and the individual seed shadows, accurately tracked with DNA-based genotyping methods, thus reflect the complex effects of frugivore foraging, habitat preferences and heterogeneous landscapes.
+This situation is probably generalizable to other plant--frugivore interactions (Morales and Carlo, 2006) where the combined spatial dynamics of habitat use and digestion processes determine complex seed shadows (Jordano et al., 2007; see Dennis and Westcott, [REF]; see Nathan, [REF]).
+Much of this complexity can be adequately handled by mechanistic models (Nathan et al., 2002; Morales and Carlo, 2006) incorporating very simple rules (Viswanathan et al., 2002).
+For example, earlier results (Jordano 2007) showed that the dispersal distances contributed by \textit{P.
+mahaleb} frugivores closely map the spacing patterns of fruiting trees, but only up to a certain distance ($\leq100 m$; Fig.
+10.3a in Jordano 2007).
+Beyond this, frugivores were probably responding to other major landscape elements (e.g.
+rock outcrops, forest edges, large patches of open grassland, etc) that cause the fat tail of the seed dispersal distribution, adding more frequent LDD events than expected from a Brownian random walk pattern generated by a tracking of the crops of the fruiting trees.
+For instance, the long flights performed by \textit{T.
+viscivorus} (Jordano and Schupp, 2000) frequently faced the pine forest edge, at distances $\geq100 m$ of most \textit{P.
+mahaleb} fruiting trees.
+If these medium-sized birds are selecting habitat with tall woody vegetation (e.g.
+pines $\geq6 m$ height), then they should be perceiving a much more patchy landscape, and thus requiring longer flights, than for example, small warblers seeking vegetation cover $<0.5 m$ (Fig.
+10.3b in Jordano 2007).
 

### 2016-10-09 Manuscript/MS- What is LDD.tex
Fixed Acknowldg.

@@
 
+LDD, and its variation across coexisting plant species, could also have far-reaching consequences for community assembly and forest physiognomy.
+Yet very few previous analyses addressed this point.
+Comparative information on LDD across species sharing a common environment have found strong differences in LDD potential among plants with different (e.g.
+Clark et al.
+1999 Ecology; Martínez et al.
+2008 Oecologia) or even with the same dispersal syndrome (e.g.
+García et al.
+2016 Basic Applied Ecology).
+
 The actual challenges to properly characterize the typologies of LDD events outlined in Table 1 will probably persist.
@@
 I appreciate the help and advice of Cristina García and Etienne Klein during the final stages of the manuscript.
-The study was supported by a Junta de Andalucía Excellence Grant (RNM-5731), as well as a Severo Ochoa Excellence Award from the Ministerio de Economía y Competitividad (SEV-2012-0262).
-The Agencia de Medio Ambiente, Junta de Andalucía, provided generous facilities that made possible this study in the Andalusian natural parks (Sierra de Cazorla, Alcornocales) and authorized my work there.
+The study was supported by a Junta de Andalucía Excellence Grant (RNM-5731), as well as a Severo Ochoa Excellence Award from the Ministerio de Economía y Competitividad (SEV-2012-0262) and CGL2013-47429P grant.
+The Consejería de Medio Ambiente, Junta de Andalucía, provided generous facilities that made possible this study in the Andalusian natural parks (Sierra de Cazorla, Alcornocales) and authorized my work there.
 

### 2016-10-09 Manuscript/MS- What is LDD.tex
Edited Discussion refs and Intro.

@@
 mahaleb} seed rain and the individual seed shadows, accurately tracked with DNA-based genotyping methods, thus reflect the complex effects of frugivore foraging, habitat preferences and heterogeneous landscapes.
-This situation is probably generalizable to other plant--frugivore interactions (Morales and Carlo, 2006) where the combined spatial dynamics of habitat use and digestion processes determine complex seed shadows (Jordano et al., 2007; see Dennis and Westcott, [REF]; see Nathan, [REF]).
-Much of this complexity can be adequately handled by mechanistic models (Nathan et al., 2002; Morales and Carlo, 2006) incorporating very simple rules (Viswanathan et al., 2002).
-For example, earlier results (Jordano 2007) showed that the dispersal distances contributed by \textit{P.
-mahaleb} frugivores closely map the spacing patterns of fruiting trees, but only up to a certain distance ($\leq100 m$; Fig.
-10.3a in Jordano 2007).
+This situation is probably generalizable to other plant--frugivore interactions where the combined spatial dynamics of habitat use and digestion processes determine complex seed shadows \citep{Westcott:200,Jordano:2007,Nathan:2008is}.
+Much of this complexity can be adequately handled by mechanistic models \citep{Nathan:2002a} incorporating very simple rules \citep{Guttal:2011hs}.
+For example, earlier results \citep{Jordano:2007a} showed that the dispersal distances contributed by \textit{P.
+mahaleb} frugivores closely map the spacing patterns of fruiting trees, but only up to a certain distance ($\leq$100 m) \citep[Fig.
+10.3a in][]{Jordano:2007a}.
 Beyond this, frugivores were probably responding to other major landscape elements (e.g.
@@
 For instance, the long flights performed by \textit{T.
-viscivorus} (Jordano and Schupp, 2000) frequently faced the pine forest edge, at distances $\geq100 m$ of most \textit{P.
+viscivorus} \citep{Jordano:2000ft} frequently faced the pine forest edge, at distances $\geq$100 m of most \textit{P.
 mahaleb} fruiting trees.
 If these medium-sized birds are selecting habitat with tall woody vegetation (e.g.
-pines $\geq6 m$ height), then they should be perceiving a much more patchy landscape, and thus requiring longer flights, than for example, small warblers seeking vegetation cover $<0.5 m$ (Fig.
-10.3b in Jordano 2007).
+pines $\geq$6 m height), then they should be perceiving a much more patchy landscape, and thus requiring longer flights, than for example, small warblers seeking vegetation cover $<$0.5 m \citep[Fig.
+10.3b in][]{Jordano:2007a}.
 

### 2016-10-10 Manuscript/MS- What is LDD.tex
Edits rev #2.

@@
 In the specific case of plants, dispersal has three basic components: 1) a distinct (sessile) source, the maternal plant producing the fruits or the paternal tree acting as a source of pollen; 2) a distance component between source and target locations; and 3) a vector actually performing the movement entailing the dispersal event.
-Here we discuss operative definitions of LDD based on their intrinsic properties: 1) events crossing geographic boundaries among stands; and 2) events contributing to effective gene flow and propagule migration.
+Here I discuss operative definitions of LDD based on their intrinsic properties: 1) events crossing geographic boundaries among stands; and 2) events contributing to effective gene flow and propagule migration.
 \\
@@
 4.
-\textit{Synthesis.} We expect truncation of seed dispersal kernels to have multiple consequences on demography and genetics, following to the loss of key dispersal services in natural populations.
+\textit{Synthesis.} I expect truncation of seed dispersal kernels to have multiple consequences on demography and genetics, following to the loss of key dispersal services in natural populations.
 Irrespective of neighborhood sizes, loss of LDD events may result in more structured and less cohesive genetic pools, with increased isolation-by-distance extending over broader areas.
@@
 
-We aim at providing a general framework for the quantitative analysis of LDD events so that estimates of their frequency and extent could be comparable across different study systems.
-We argue that both demographic and genetic elements are needed for this framework, most likely requiring a combination of field-based movement data and genetic analyses.
+I aim at providing a general framework for the quantitative analysis of LDD events so that estimates of their frequency and extent could be comparable across different study systems.
+I argue that both demographic and genetic elements are needed for this framework, most likely requiring a combination of field-based movement data and genetic analyses.
 These elements can be overlaid on previous definitions based on absolute and proportional characterizations of LDD.
-We start with a definition of LDD events within a spatially-explicit mechanistic framework allowing an unambiguous meaning for setting long-distance thresholds.
-We then use a case study to assess differential contributions of animal frugivores performing LDD.
+I start with a definition of LDD events within a spatially-explicit mechanistic framework allowing an unambiguous meaning for setting long-distance thresholds.
+I then use a case study to assess differential contributions of animal frugivores performing LDD.
 
 Long-distance dispersal is currently one of the most debated topics in dispersal ecology; it defines the connectedness within the network of local populations and the possibilities for range expansion and successful colonization events.
-We propose a first demogenetically-based, operational definition of what a long-distance dispersal event actually is, and review existing empirical literature on distance thresholds from population and genetic perspectives.
-We also show how molecular tools have been used to identify the respective contributions of different animal species to the LDD portion of dispersal kernels of pollen and seeds by setting empirically-derived distance thresholds.
-Finally, we highlight potential applications of molecular markers beyond the quantification of just the dispersal distances that prevails in current studies, e.g., experimental approaches to assess dispersal limitation and Janzen-Connell effects.
+I propose a first demogenetically-based, operational definition of what a long-distance dispersal event actually is, and review existing empirical literature on distance thresholds from population and genetic perspectives.
+I also show how molecular tools have been used to identify the respective contributions of different animal species to the LDD portion of dispersal kernels of pollen and seeds by setting empirically-derived distance thresholds.
+Finally, I highlight potential applications of molecular markers beyond the quantification of just the dispersal distances that prevails in current studies, e.g., experimental approaches to assess dispersal limitation and Janzen-Connell effects.
 
@@
 
-Here we propose an explicit definition of LDD and what constitutes a LDD event.
+Here I propose an explicit definition of LDD and what constitutes a LDD event.
 Previous definitions of dispersal patterns emphasized only their distance components and characterized LDD events basically in terms of geographic distance between a dispersed propagule (or an established early seedling) and its most likely maternal or paternal (in case of pollen) source.
@@
 While both $SDD_{loc}$ and $LDD_{loc}$ can be crucial for assuring the local persistence of populations, $LDD_{neigh}$ and $LDD_{ss}$ would be extremely important contributors to the structuring of genetic pools, realized gene flow, and maintaining connectivity in metapopulation scenarios.
-We argue that both the demographic and the genetic references are relevant for a proper definition of LDD.
+I argue that both the demographic and the genetic references are relevant for a proper definition of LDD.
 
@@
 On one hand, theoretical models fail to predict accurately the behavior of the tail of the dispersal functions, and thus fail to predict very basic properties of LDD.
-On the other hand, we still have very limited documentation of actual LDD events in natural populations and we still see LDD as a sporadic, rarely far-reaching process still marked with the stamp of natural history curiosity.
+On the other hand, we have very limited documentation of actual LDD events in natural populations and we still see LDD as a sporadic, rarely far-reaching process still marked with the stamp of natural history curiosity.
 
@@
 
-We advocate \citep[also see ][]{Jordano:2002tr,Nathan:2003qe,Jones:2008il,Hardesty:2011jn} a combination of approaches including large-scale biotelemetry to characterize animal movement, coupled with large-scale genetic sampling of dispersed propagules, and demogenetic approaches that combine both demographic and genetic research.
+I advocate \citep[also see ][]{Jordano:2002tr,Nathan:2003qe,Jones:2008il,Hardesty:2011jn} a combination of approaches including large-scale biotelemetry to characterize animal movement, coupled with large-scale genetic sampling of dispersed propagules, and demogenetic approaches that combine both demographic and genetic research.
 A crucial aspect would be to effectively associate the role of individual frugivore species to specific dispersal outcomes, by identifying the actual disperser contributing a dissemination event \citep{GonzalezVaro:2014ij} and simultaneously characterizing the source maternal plant \citep{Jordano:2002tr}.

### 2016-10-10 Manuscript/MS- What is LDD.tex
Updated refs to Introduction.

@@
 
-% EDIT HERE
 While my main focus in on population-level scenarios and the role of animal vectors, recent research has shown how relevant is habitat and landscape context in determining the occurrence and directionality of LDD (e.g.
@@
 Both approaches are limited logistically by the difficulties to sample the vast areas required to assess LDD events from the focal source population.
+
+LDD events have two key characteristics that make them extremely important for population dynamics, yet being very difficult to characterize: LDD events are infrequent, but with a disproportionately high influence on contemporary gene flow and structure of the genetic pools \citep[e.g., ][]{Schurr2009long,Clobert:2012vw,Travis:2013en}.
+LDDs can connect disparate populations, allowing for genetic connectivity, colonization of vacant habitat and range expansion across changing landscapes, and maintain global persistence in the face of local extinctions \citep{Trakhtenbrot:2005ij,Baguette:2006aa,Ronce:2007hc,Schloss:2012gt}.
+With their influence on the structure of genetic pools, LDD events can also drive population differentiation and speciation \citep{Ronce:2007hc}.
 

### 2016-10-10 Manuscript/MS- What is LDD.tex
Edited refs and Suppl Mat.

@@
 This conceptual isolation has been exacerbated by technical difficulties for the robust characterization of dispersal events, especially those involving movement over long-distances (long-distance dispersal, LDD).
-LDD is a characteristically extreme event of propagule movement in any plant population, typically occurring with an extremely low probability but potentially reaching an extremely long distance.
 Some progress has recently been made through the fast-paced implementation of molecular tools in ecological research labs and the availability of cutting-edge technology for biotelemetry applications.
@@
 No explicit definition of what constitutes an LDD event exists.
+LDD is a characteristically extreme event of propagule movement in any plant or animal population, typically occurring with an extremely low probability but potentially reaching an extremely long distance.
 Previous approaches \citep[e.g., ][]{Nathan:2006aa,Schurr2009long} include both absolute and proportional definitions to characterize LDD events.
-This means providing information about the absolute distances moved by a given percentile of the events and/or providing data on the proportion of events exceeding a given distance threshold \citep{Nathan:2008is}.
-The exact proportional or absolute thresholds selected remain arbitrary, as no reference spatial frame is provided within the definition of LDD.
+This means providing information about the absolute distances moved by a given percentile of the events and/or providing data on the proportion of events exceeding a given distance threshold, i.e., an operational definition \citep{Nathan:2008is}.
+The exact proportional or absolute thresholds selected remain somehow arbitrary, as no reference spatial frame is provided within the definition of LDD.
 This leaves the consideration of LDD as an extreme form of context-dependent phenomenon, strongly dependent upon the scale of the biological process studied \citep{Kinlan:2005fb} and of the specific organism considered.
 For example, \cite{Kinlan:2005fb} used a spatial reference frame to characterize LDD events of marine organisms, where sedentary adults and larvae differ enormously in the spatial scales of their dispersal \citep{DAloia:2013fc}.
-Therefore, any measure of extent and reach of LDD events requires reference to an explicit spatial frame or "local" scale \citep{Kinlan:2005fb}.
+Therefore, any measure of extent and reach of LDD events requires reference to an explicit spatial frame or "local" scale \citep{Kinlan:2005fb,Byrne:2014gu}.
 
@@
 Long-distance dispersal is currently one of the most debated topics in dispersal ecology; it defines the connectedness within the network of local populations and the possibilities for range expansion and successful colonization events.
-I propose a first demogenetically-based, operational definition of what a long-distance dispersal event actually is, and review existing empirical literature on distance thresholds from population and genetic perspectives.
+I propose a first demogenetically-based, operational definition of what a LDD event actually is, and review existing empirical literature on distance thresholds from population and genetic perspectives.
 I also show how molecular tools have been used to identify the respective contributions of different animal species to the LDD portion of dispersal kernels of pollen and seeds by setting empirically-derived distance thresholds.
@@
 Along a similar reasoning, dispersal events outside the population limits will not necessarily convey LDD (Table 1, Fig.
-2B): this is expected in cases where the genetic neighborhoods are extensive, going beyond the geographic limits of local populations, as in fig trees \citep{Nason:1998aa}, generating LDD events within the genetic neighborhood ($LDD_{neigh}$).
+2B): this is expected in cases where the genetic neighborhoods are extensive, going beyond the geographic limits of local populations, as in fig trees \citep{Nason:1998aa} with long-distance pollination, generating LDD events within the genetic neighborhood ($LDD_{neigh}$).
+Note that pollen and seeds may have contrasting movement patterns in rference to the distinct spatial scales of the population limits and of the genetic neighbourhood.
+For example, wind-dispersed species with reduced seed mobility (in terms of distance), such as oaks, can have large genetic neighbourhoods with extensive pollen dispersal \citep{Streiff99} \citep[but see, e.g.,][for fragmented stands]{Smouse:2001aa,Dutech:2005fk} so that $LDD_{neigh}$ dispersal events might frequently move beyond the physical limits of the population, patch, or stand but remain within the genetic neighbourhood.
 Finally, strict-sense LDD events would involve dissemination outside \textit{both} the population limits and the genetic neighborhood boundary ($LDD_{ss}$) (Table 1, Fig.

### 2016-10-10 Manuscript/SupplMat.tex
Minor edits.
Minor edits to title page.
Edited refs and Suppl Mat.

@@
 
-\title{Manuscript Draft\\
+\title{\textbf{Supplementary Material}\\
 \vspace{2cm}
 What is long-distance dispersal? And a taxonomy of dispersal events \\
-\textbf{Supplementary Material}}
+}
 
@@
 Summary of neighborhood area sizes and estimated neighborhood radius for tree species with different combinations of dispersal modes.
-Data from \citet{Nason:1998aa,Garcia:2005fu,Garcia:2007he} and present study.}
+Data from \citet{Nason:1998aa,Smouse:2001aa,Garcia:2005fu,Garcia:2007he,Dutech:2005fk} and present study.}
+  \vspace{0.5cm}
     \begin{tabular}{lllccc}
-     Species                & Pollinator  & Seed disperser & Density ($ha^{-1}$) & Breeding unit ($km^2$) & Radius (km) \\
+        \hline \\
+     Species                & Pollinator  & Seed disperser & Density ($ha^{-1}$) & Breeding unit ($km^2$) & Radius (km) \\\\ \hline \\
     \textit{Ficus dugandii}          & Fig wasp    & Vertebrates    & 0.004          & 631.7               & 14.2        \\
@@
     \textit{Platypodium elegans}     & Bees        & Wind           & 0.78           & 0.866               & 0.525       \\
+    \textit{Quercus lobata}          & Wind        & Vertebrates    & 3.0
+      & 0.072               & 0.719       \\ 
+    \textit{Quercus alba}            & Wind        & Vertebrates    & 92.8
+      & 0.0009              & 0.017       \\ 
     \textit{Cedrus atlantica}        & Wind        & Wind           & 61.7           & 0.151               & 0.22        \\
    \textit{Fraxinus americana}       & Wind        & Wind           & 24.7           & 0.008                & 0.05        \\
-    \textit{Pseudotsuga menziesii}   & Wind        & Wind           & 25.0           & 0.078               & 0.158       \\
+    \textit{Pseudotsuga menziesii}   & Wind        & Wind           & 25.0           & 0.078               & 0.158       \\\\
+\hline
     \end{tabular}
@@
 \newpage 
+\newpage
 

### 2016-10-10 Data/kernels.Rmd
Added distances.txt

@@
 # Get the data from my GitHub repository.
-link = "https://raw.githubusercontent.com/pedroj/dispkernels/master/distances.txt"
+link = "https://raw.githubusercontent.com/pedroj/dispkernels/master/Datasets/distances.txt"
 file = "distances.txt"

### 2016-10-10 Manuscript/MS- What is LDD.tex
Added distances.txt

@@
 2B): this is expected in cases where the genetic neighborhoods are extensive, going beyond the geographic limits of local populations, as in fig trees \citep{Nason:1998aa} with long-distance pollination, generating LDD events within the genetic neighborhood ($LDD_{neigh}$).
-Note that pollen and seeds may have contrasting movement patterns in rference to the distinct spatial scales of the population limits and of the genetic neighbourhood.
+Note that pollen and seeds may have contrasting movement patterns in reference to the distinct spatial scales of the population limits and of the genetic neighbourhood.
 For example, wind-dispersed species with reduced seed mobility (in terms of distance), such as oaks, can have large genetic neighbourhoods with extensive pollen dispersal \citep{Streiff99} \citep[but see, e.g.,][for fragmented stands]{Smouse:2001aa,Dutech:2005fk} so that $LDD_{neigh}$ dispersal events might frequently move beyond the physical limits of the population, patch, or stand but remain within the genetic neighbourhood.
@@
 I argue that both the demographic and the genetic references are relevant for a proper definition of LDD.
+These differences may have different implications in relation to the dispersal biology of species.
+For example, frugivorous vertebrates, particularly the large-sized ones, tend to disperse a high number of half-sibs propagules, from the same mother tree to the same deposition site \citep{Jordano:2000ft}, potentially increasing the frequency of $LDD_{neigh}$ events if dispersal distances are relatively short.
 

### 2016-10-10 Data/kernels.Rmd
Edited adehabitatHS code for dist.

--- 2015-02-15 Data/kernels.Rmd
@@
 date: "enero 29, 2015"
-output: html_document
-bibliography: refs.bib
+output: html_notebook
+#bibliography: refs.bib
 ---
@@
                                                # req. adehabitatHS
-dd<- as.data.frame(simm.levy(1:1000, mu = 1.5, burst = "mu = 1.5")) 
-dd<- na.omit(dd$dist)
+library(adehabitatHS)
+dd<- simm.levy(1:1000, mu = 1.5, burst = "mu = 1.5")
+dd<- na.omit(dd[[1]]$dist)
 

### 2016-10-10 Manuscript/MS- What is LDD.tex
Updated Suppl Mat and refs in main text.

--- 2016-10-10 Manuscript/MS- What is LDD.tex
@@
 I argue that both the demographic and the genetic references are relevant for a proper definition of LDD.
+Accounting for these references may have different implications in relation to the dispersal biology of species.
+For example, frugivorous vertebrates, particularly the large-sized ones, tend to disperse a high number of half-sibs propagules, from the same mother tree to the same deposition site \citep{Jordano:2007,Karubian:2010br,Scofield:2012gk}, potentially increasing the frequency of $LDD_{neigh}$ events if dispersal distances are relatively short.
 

### 2016-10-11 Manuscript/MS- What is LDD.tex
Edited Intro.

@@
 Our aim here is to review the LDD concept with a specific emphasis on dispersal of plant propagules (seeds and pollen), providing an extended definition that might be helpful in the robust quantification of LDD events.
+
+An added difficulty to bridge ecological processes of dispersal with their genetic consequences is the fact that dispersal per se does not necessarily imply realized gene flow \citep{Mallet:2001wx}.
+Yet in the genetics literature, the inescapable difficulty to robustly estimate $\sigma_x$, the the standard deviation of parent-offspring distances, as an estimate of realized gene flow distance, is analogous to the ecological scenario.
+Strongly leptokurtic distributions of dispersal distances would lead to severe underestimation of $\sigma_x$, e.g., in mark-recapture studies limited to local sites where the long tail of LDD events escapes detection beyond the edges of the study area \citep{Mallet:2001wx}.
+Thus, fat-tailed distributions of dispersal distances tend to homogenize distant populations, leading to stronger reductions in $F_{st}$ than expected from local records of dispersal events \citep{Rousset:1997bd}.
 

### 2016-10-11 Manuscript/MS- What is LDD.tex
Final edits to the Introduction.

@@
 Thus, fat-tailed distributions of dispersal distances tend to homogenize distant populations, leading to stronger reductions in $F_{st}$ than expected from local records of dispersal events \citep{Rousset:1997bd}.
+Despite such difficulties, comparative approaches \citep[e.g.,][]{Bohonak:1999aa} demonstrate that dispersal makes a measurable contribution to population genetic differentiation in most animal species in nature, and that gene flow estimates are rarely so overwhelmed by population history, departures from equilibrium, or other microevolutionary forces as to be uninformative.
+The relationship between dispersal and realized gene flow is key to understand how migration interacts with drift in driving the dynamics of genetic pools and population differentiation \citep{Garant:2007cka}.
 
@@
 LDDs can connect disparate populations, allowing for genetic connectivity, colonization of vacant habitat and range expansion across changing landscapes, and maintain global persistence in the face of local extinctions \citep{Trakhtenbrot:2005ij,Baguette:2006aa,Ronce:2007hc,Schloss:2012gt}.
-With their influence on the structure of genetic pools, LDD events can also drive population differentiation and speciation \citep{Ronce:2007hc}.
+With their influence on the structure of genetic pools, LDD events can also drive population differentiation and speciation \citep{Garant:2007cka,Ronce:2007hc}.
 

### 2016-10-11 Manuscript/MS- What is LDD.tex
Edited Figure legends.

@@
 In the specific case of plants, largely sessile organisms, dispersal has three basic components: 1) a distinct (sessile) source, the maternal plant producing the fruits or the paternal tree acting as a source of pollen; 2) a distance component between source and target locations; and 3) a vector actually performing the movement entailing the dispersal event.
-While realized dispersal also depends upon stages subsequent to dissemination (e.g., successful germination and seedling establishment) \cite{Schupp:1995}, the three previous components fully characterize the dispersal process per se.
+While realized dispersal also depends upon stages subsequent to dissemination \citep[e.g., successful germination and seedling establishment;][]{Schupp:1995}, the three previous components fully characterize the dispersal process per se.
 Therefore, plant movement differs in important natural history details from animal dispersal, yet both can be assessed within a common conceptual framework \citep[e.g., ][]{Nathan:2006aa}.
@@
 LDD, and its variation across coexisting plant species, could also have far-reaching consequences for community assembly and forest physiognomy.
-Yet very few previous analyses addressed this point.
-Comparative information on LDD across species sharing a common environment have found strong differences in LDD potential among plants with different (e.g.
-Clark et al.
-1999 Ecology; Martínez et al.
-2008 Oecologia) or even with the same dispersal syndrome (e.g.
-García et al.
-2016 Basic Applied Ecology).
+Yet very few previous analyses address this point.
+Comparative information on LDD across species sharing a common environment have found strong differences in LDD potential among plants with different \citep[e.g.,][]{Clark:1999sv,Martinez:2008jx} or even with the same dispersal syndrome \citep{Garcia:2016bd}.
 
@@
 The two approaches used in analyses of dispersal processes in plants.
-A, the “forward” approach attempts to track the dispersal events away from the known sources, e.g., by tracking the movement patterns of frugivores as they leave fruiting plants after feeding.
-B, the “backward” approach  attempts to reconstruct the most likely source of a dispersed propagule by inferring the sources given the propagule delivery pattern, the fecundity of potential sources, and the dispersal function.
+A, the “forward” approach attempts to track the dispersal events away from the known sources, e.g., by tracking the movement patterns of frugivores as they leave fruiting plants after feeding (i.e., the Lagrangian approach, \citep{Nathan:2003qe}).
+B, the “backward” approach  attempts to reconstruct the most likely source of a dispersed propagule by inferring the sources given the propagule delivery pattern, the fecundity of potential sources, and the dispersal function (i.e., the Eulerian approach).
 The main technical challenge in A is to sample enough dispersal events away from the source to be able to fully characterize the tail (long-distance dispersal, LDD, events) of the dispersal function.
-In B, the main challenge is to have a robust sampling scheme with propagule collectors (e.g., seed traps) and a good characterization of the potential sources to derive robust estimates of the actual sources with inverse-modeling techniques.\\
+In B, the main challenge is to have a robust sampling scheme with propagule collectors (e.g., seed traps) and a good characterization of the potential sources to derive robust estimates of the actual sources with inverse-modeling techniques \citep{Nathan:2003qe}.\\
 
@@
 Strict-sense long-distance dispersal events ($LDD_{ss}$) just include the LDD events that disseminate propagules out of \textit{both} the population and genetic neighborhood boundaries.
-A, the neighborhood area is included within the geographic limits of the population, with some dispersal events potentially contributing local LDD; B, the neighborhood area is much larger than the geographic limits of the population.\\
+A, the neighborhood area is included within the geographic limits of the population, with some dispersal events potentially contributing local LDD; B, the neighborhood area is much larger than the geographic limits of the population.
+Both $LDD_{ss}$ and $LDD_{loc}$ may contribute genetic novelty by moving propagules outside the $N^b_e$ area or the population limit, area or both.\\
 

### 2016-10-11 Manuscript/MS- What is LDD.tex
Edits.

@@
 In particular, LDD events remain difficult to assess, both technically- with serious methodological problems for its reliable estimation- and conceptually.
-Our aim here is to review the LDD concept with a specific emphasis on dispersal of plant propagules (seeds and pollen), providing an extended definition that might be helpful in the robust quantification of LDD events.
+My aim is to review the LDD concept with a specific emphasis on dispersal of plant propagules (seeds and pollen), providing an extended definition that might be helpful in the robust quantification of LDD events.
 
@@
 
-While my main focus in on population-level scenarios and the role of animal vectors, recent research has shown how relevant is habitat and landscape context in determining the occurrence and directionality of LDD (e.g.
-Schurr et al.
-2008 J Ecol; Carlo et al.
-2013).
+While my main focus in on population-level scenarios and the role of animal vectors, recent research has shown how relevant is habitat and landscape context in determining the occurrence and directionality of LDD \citep{Schurr:2008bd,Carlo:2013vd}.
 On the other hand, dealing with habitat effects seems obligated if aimed to discuss global change drivers (i.e.
@@
 1).
-The “forward” approach attempts to track the dispersal events away from the known sources, e.g., by tracking the movement patterns of frugivores as they leave fruiting plants after feeding (Fig.
+The “forward” (Lagrangian) approach attempts to track the dispersal events away from the known sources, e.g., by tracking the movement patterns of frugivores as they leave fruiting plants after feeding (Fig.
 1A).
 This is the main approach used in the movement ecology framework \citep{Nathan:2008fx}, with extensive application to animal movement based on the use of advanced biotelemetry.
-The “backward” approach attempts to reconstruct the most likely source of a dispersed propagule by inferring the sources given the propagule delivery pattern, the fecundity of potential sources, and the dispersal function (Fig.
+The “backward” (Eulerian) approach attempts to reconstruct the most likely source of a dispersed propagule by inferring the sources given the propagule delivery pattern, the fecundity of potential sources, and the dispersal function (Fig.
 1B), i.e., by using an inverse modeling approach.
@@
 The two approaches used in analyses of dispersal processes in plants.
-A, the “forward” approach attempts to track the dispersal events away from the known sources, e.g., by tracking the movement patterns of frugivores as they leave fruiting plants after feeding (i.e., the Lagrangian approach, \citep{Nathan:2003qe}).
+A, the “forward” approach attempts to track the dispersal events away from the known sources, e.g., by tracking the movement patterns of frugivores as they leave fruiting plants after feeding \citep[i.e., the Lagrangian approach;][]{Nathan:2003qe}.
 B, the “backward” approach  attempts to reconstruct the most likely source of a dispersed propagule by inferring the sources given the propagule delivery pattern, the fecundity of potential sources, and the dispersal function (i.e., the Eulerian approach).
 The main technical challenge in A is to sample enough dispersal events away from the source to be able to fully characterize the tail (long-distance dispersal, LDD, events) of the dispersal function.
-In B, the main challenge is to have a robust sampling scheme with propagule collectors (e.g., seed traps) and a good characterization of the potential sources to derive robust estimates of the actual sources with inverse-modeling techniques \citep{Nathan:2003qe}.\\
+In B, the main challenge is to have a robust sampling scheme with propagule collectors (e.g., seed traps) and a good characterization of the potential sources to derive robust estimates of the actual sources with inverse-modeling techniques.\\
 

### 2016-10-11 Manuscript/MS- What is LDD.tex
Final version R1 completed.

@@
 
-{\small \textbf{Manuscript information: }** Words; ** Chars; ** Pages, * Figures; * Tables.}
+{\small \textbf{Manuscript information: } 8979 Words; 37 Pages, 4 Figures; 2 Tables.}
 \end{spacing}
@@
 An added difficulty to bridge ecological processes of dispersal with their genetic consequences is the fact that dispersal per se does not necessarily imply realized gene flow \citep{Mallet:2001wx}.
-Yet in the genetics literature, the inescapable difficulty to robustly estimate $\sigma_x$, the the standard deviation of parent-offspring distances, as an estimate of realized gene flow distance, is analogous to the ecological scenario.
+Yet in the genetics literature, the inescapable difficulty to robustly estimate the the standard deviation of parent-offspring distances, $\sigma_x$ as a proxy for realized gene flow distance, is analogous to the ecological scenario.
 Strongly leptokurtic distributions of dispersal distances would lead to severe underestimation of $\sigma_x$, e.g., in mark-recapture studies limited to local sites where the long tail of LDD events escapes detection beyond the edges of the study area \citep{Mallet:2001wx}.
@@
 These "frontiers" set biological limits to what a LDD event is in relation to the geographic limits of the source population.
-Most plants are distributed as clumped patches, discrete stands, or relatively isolated populations, so we may distinguish between short-distance and long-distance dispersal events that end up with dissemination within or beyond, respectively, the stand or population geographic boundaries (Table 1, $SDD_{loc}$ or $LDD_{loc}$) (Figure 2).
+Most plants are distributed as clumped patches, discrete stands, or relatively isolated populations, so we may distinguish between short-distance and long-distance dispersal events that end up with dissemination within or beyond, respectively, the stand or population geographic boundaries (Table 1, $SDD_{loc}$ or $LDD_{loc}$) (Fig.
+2).
 

### 2016-10-15 Manuscript/MS- What is LDD.tex
Added DRYAD ref.

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Online Support Material and data accessiblity}
+
+This review does not use new raw data, but includes some re-analyses of previously published material.
+All the original data supporting the paper, R code, supplementary figures, and summaries of analytical protocols is available at the author's GitHub repository (\texttt{https://github.com/pedroj/MS\_LDD}), and also deposited in the Dryad repository \citep{Jordano:2016ab}: \\\texttt{http://datadryad.org/resource/doi:10.5061/dryad.k0v67}.
+
 \end{linenumbers}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Online Support Material and data accessiblity}
-
-This review does not use new raw data, but includes some re-analyses of previously published material.
-All the original data supporting the paper, R code, supplementary figures, and summaries of analytical protocols is available at the author's GitHub repository (\texttt{https://github.com/pedroj/MS\_LDD}), with DOI: \texttt{\#/zenodo.\#}.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
